### 2015-06-17 paper/abbreviations.tex
first version of draft

% -*- root: main.tex -*-
% @Author: Oscar Esteban
% @Date:   2015-06-16 15:45:32
% @Last Modified by:   Oscar Esteban
% @Last Modified time: 2015-06-17 16:28:24

\newacronym{dmri}{dMRI}{diffusion MRI}
\newacronym{hcp}{HCP}{Human Connectome Project}

\makeglossaries

### 2015-06-17 paper/main.tex
first version of draft

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is just an example/guide for you to refer to when submitting manuscripts to Frontiers, it is not mandatory to use Frontiers .cls files nor frontiers.tex  %
% This will only generate the Manuscript, the final article will be typeset by Frontiers after acceptance.                                                 %
%                                                                                                                                                         %
% When submitting your files, remember to upload this *tex file, the pdf generated with it, the *bib file (if bibliography is not within the *tex) and all the figures.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Version 3.1 Generated 2015/22/05 %%%
%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%

\documentclass[english]{frontiers/frontiersSCNS} % for Science, Engineering and Humanities and Social Sciences articles
%\documentclass{frontiersHLTH} % for Health articles
%\documentclass{frontiersFPHY} % for Physics and Applied Mathematics and Statistics articles

%\setcitestyle{square}
\usepackage{url,lineno,microtype}
\usepackage[colorlinks=true,linkcolor=black, citecolor=black!80, urlcolor=black!80]{hyperref}
\usepackage{doi}
\usepackage[onehalfspacing]{setspace}
\usepackage[toc,nomain,acronym,shortcuts,translate=false]{glossaries}

\def\keyFont{\fontsize{7}{9}\helveticabold }
\def\firstAuthorLast{Esteban {et~al.}} %use et al only if is more than 1 author
\def\Authors{Oscar Esteban\,$^{1,2,*}$, Emmanuel Caruyer\,$^{3}$, Alessandro Daducci\,$^{4}$, Meritxell Bach-Cuadra\,$^{4,5}$,%
M.-Jes\'us Ledesma-Carbayo\,$^{1,2}$ and Andres Santos\,$^{1,2}$}

\def\Address{%
$^{1}$Biomedical Image Technologies (BIT), ETSI Telecomunicaci\'on, Universidad Polit\'ecnica de Madrid, Madrid, Spain \\
$^{2}$Centro de Investigaci\'on Biom\'edica en Red en Bioingenier\'ia, Biomateriales y Nanomedicina (CIBER-BBN), Spain \\
$^{3}$CNRS, IRISA (UMR 6074), VisAGeS research group, Rennes, France \\
$^{4}$Signal Processing Laboratory (LTS5), \'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL), Lausanne, Switzerland \\
${^5}$Dept. of Radiology, CIBM, University Hospital Center (CHUV) and University of Lausanne (UNIL), Lausanne, Switzerland %
}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Oscar Esteban}
\def\corrAddress{Biomedical Image Technologies (BIT), ETSI Telecomunicaci\'on, Av. Complutense 30, C203, E28040 Madrid, Spain}
\def\corrEmail{phd@oscaresteban.es}

\input{abbreviations.tex}
\begin{document}

\onecolumn
\firstpage{1}

\title[Diffantom]{Diffantom: a software layer to build digital phantoms of whole-brain diffusion MRI}

\author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
\address{} %This field will be automatically populated
\correspondance{} %This field will be automatically populated

\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}

\maketitle

\linenumbers
\section{Introduction}
The assessment of tractography from \gls*{dmri} data and subsequent analyses is challenging due to
  the lack of a gold standard.
Large efforts have been devoted to the development of physical phantoms
  \citep{lin_validation_2001,campbell_flowbased_2005,perrin_validation_2005,fieremans_simulation_2008,tournier_resolving_2008}.
\cite{cote_tractometer_2013} conducted a thorough review of tractography methodologies using the
  so-called \emph{FiberCup} phantom \citep{poupon_new_2008,fillard_quantitative_2011}.
These phantoms are appropriate to evaluate the resolution of fiber crossings and accuracy of
  direction-independent scalar parameters in very simplistic geometries.
Since the complexity of whole-brain tractography is unaccountable with current materials and proposed
  methodologies to build physical phantoms, digital simulations are increasingly popular.
Early digital phantoms started with simulation of simple geometries as well
  \citep{basser_in_2000,goessl_fiber_2002,tournier_limitations_2002,leemans_mathematical_2005}
  to evaluate resolution of fiber crossings.
These tools generally implemented the multi-tensor model \citep{alexander_analysis_2001,tuch_high_2002}
  to simulate fibers crossing, fanning, kissing, etc.
\cite{close_software_2009} presented the \emph{Numerical Fibre Generator}, a software to simulate
  spherical shapes filled with digital fiber-tracts.
\cite{caruyer_phantomas_2014} proposed \emph{Phantomas} to simulate any kind of analytic geometry
  inside a sphere.
\emph{Phantomas} models diffusion by a restricted and a hindered compartment, similar to
  \citep{assaf_composite_2005}.
\cite{wilkins_fiber_2014} proposed a whole-brain simulated phantom derived from voxel-wise orientation
  of fibers averaged from real \gls*{dmri} scans and the multi-tensor model with a compartment of
  isotropic-diffusion.
Recently, \cite{neher_fiberfox_2014} proposed \emph{FiberFox}, a visualization software to develop
  complex geometries and their analytical description.
Once the geometries are obtained, the software generates the corresponding \gls*{dmri} signal with a
  methodology very close to that implemented in \emph{Phantomas}.

In this paper we present \emph{Diffantom}, a software layer to generate diffusion phantoms at whole-brain
  scale and derived from real \gls*{dmri} datasets.
Particularly, we propose the \gls*{hcp} \citep{essen_human_2012} as the source of real data.
Finally, an example dataset is openly released along with all the necessary requirements and instructions
  to generate custom phantoms.

\section{Data description}

\noindent\textbf{\textit{Data Generation\textcolon}}
The methods to generate the data in this paper are inspired by the work of \cite{wilkins_fiber_2014} and
  described in \autoref{fig:figure1}.
\emph{Diffantom} observes two major improvements with respect to \citep{wilkins_fiber_2014}.
First, since we use a dataset from the \gls*{hcp} as input, data are corrected for the most relevant distortions.
\cite{wilkins_fiber_2014} explicitly state that their original data did not subjected to certain distortion
  corrections, and thus, generated data are affected correspondingly.
Second improvement is methodological, as we use a more complete signal model to generate the phantom thanks
  to \emph{phantomas} \citep{caruyer_phantomas_2014}.

\noindent\textbf{\textit{Interpretation and use\textcolon}}
In order to release a reproducible dataset, along with an usable software to create custom data,
  we also distribute the software tools summarized in \autoref{fig:figure1}.

\section{Discussion and conclusion}

\noindent\textbf{\textit{Discussion\textcolon}}\label{sec:discussion} %
Whole-brain, realistic \gls*{dmri} phantoms are necessary in current development of connectomics.
The \emph{diffantom} is a derivative work from \citep{wilkins_fiber_2014} in terms of methodology for
  simulation with two major improvements: the correctness of the \emph{minimally preprocessed} data
  \citep{glasser_minimal_2013} released within the \gls*{hcp}, and a more realistic diffusion
  model.
A possible competitor to \emph{diffantom} is the ground-truth data generated by the organizers of the
  Tractography Challenge held in the ISMRM 2015\footnote{\url{http://www.tractometer.org/ismrm_2015_challenge/}}.
Similarly to \emph{diffantom}, one \gls*{hcp} dataset is used as source of structural information.
However, in this case 25 fiber bundles were manually segmented to obtain a summary tractogram.
Then, making use of \emph{FiberFox}, the segmentation of each bundle is mapped to an analytical description
  which is feeded into \emph{phantomas}.
While this phantom is designed for the bundle-wise evaluation of tractography (with scores such as geometrical coverage,
  valid connections, invalid connections, missed connections, etc. \citep{cote_tractometer_2013}),
  \emph{diffantom} is intended for the connectome-wise evaluation of results, yielding tractography with
  a large number of bundles.
Therefore, \emph{diffantom} is complementary as the hypotheses that can be investigated are different and
  it does not require costly manual segmentation of bundles, highly demanding in terms of physiology expertise
  and operation time.

\noindent\textbf{\textit{Conclusion\textcolon}}\label{sec:conclusion} %
\emph{Diffantom} is a software layer to be applied on \acrfull*{hcp} datasets, that is openly and freely distributed
  along with an example dataset.

\section*{Data Sharing}

\section*{Disclosure/Conflict-of-Interest Statement}

The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.

\section*{Author Contributions}
All the authors contributed to this study.
OE designed the data generation procedure, implemented the processing pipelines and generated the example dataset.
EC implemented \emph{phantomas} \citep{caruyer_phantomas_2014}, helped integrate the project with the simulation routines.
OE, EC, AD thoroughtly discussed and framed the aptness of the data in the community.
AD, MBC, MJLC, and AS interpreted the resulting datasets.
MBC, MJLC, and AS advised on all aspects of the study.

\section*{Acknowledgments}
\textit{Funding\textcolon}
This study was supported by the Spanish Ministry of Science and Innovation
  (projects TEC-2013-48251-C2-2-R and INNPACTO XIORT), Comunidad de Madrid (TOPUS) and
  European Regional Development Funds, the Center for Biomedical Imaging
  (CIBM) of the Geneva and Lausanne Universities and the EPFL, as well as the
  Leenaards and Louis Jeantet Foundations.

\bibliographystyle{frontiers/frontiersinSCNS_ENG_HUMS}
%\bibliographystyle{frontiersinHLTH&FPHY} % for Health and Physics articles
\bibliography{Remote}

%%% Upload the *bib file along with the *tex file and PDF on submission if the bibliography is not in the main *tex file

\section*{Figures}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=\linewidth]{figures/figure01}
\end{center}
\textbf{\refstepcounter{figure}\label{fig:figure1} Figure \arabic{figure}.}{ Enter the caption for your figure here.  Repeat as  necessary for each of your figures }
\end{figure}

\end{document}

### 2015-06-17 diffantom/__init__.py

@@ -1,0 +1,6 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+# @Author: oesteban
+# @Date:   2015-06-15 14:50:01
+# @Last Modified by:   oesteban
+# @Last Modified time: 2015-06-15 14:50:01

### 2015-06-18 paper/main.tex
finishing early version of paper

@@
 In order to release a reproducible dataset, along with an usable software to create custom data,
-  we also distribute the software tools summarized in \autoref{fig:figure1}.
+  we also distribute the software instrumentation presented in \autoref{fig:figure1}.
 
@@
 
+\glsresetall
+
 \section*{Figures}
@@
 \end{center}
-\textbf{\refstepcounter{figure}\label{fig:figure1} Figure \arabic{figure}.}{ Enter the caption for your figure here.  Repeat as  necessary for each of your figures }
+\textbf{\refstepcounter{figure}\label{fig:figure1} Figure \arabic{figure}. }%
+{\emph{Diffantom} is a software layer wrapping the \acrfull{hcp} datasets to generate whole-brain diffusion \gls*{dmri} data.
+First, the tissue fraction maps $\{T_1, T_2, T_3\}$ are computed using FSL FAST \citep{zhang_segmentation_2001}.
+Second, the fiber orientation maps $\{V_1, V_2, V_3\}$ and corresponding volume fraction maps $\{F_1, F_2, F_3\}$ are
+  obtained using the Multi-Shell ball-and-stick model of BEDPOSTX \citep{jbabdi_modelbased_2012}.
+The third step takes the fiber orientations $\{V_1, V_2, V_3\}$ and their weights $\{F_1, F_2, F_3\}$ to compute the
+  corresponding \glspl{fod} using spherical convolution of the single fiber response with each fiber orientation at
+  each voxel.
+The resulting \glspl{fod} map is then combined with three isotropic compartments corresponding to $\{T_1, T_2, T'_3\}$
+  before generating the signal using \emph{phantomas} \citep{caruyer_phantomas_2014}.
+The fraction map $T'_3 = T_3 - \sum F_i$ is the remaining isotropic fraction across the \gls*{wm} tissue map.
+Finally, data can be used to compute the whole-brain tractography, for instance.}
+\end{figure}
+
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[width=\linewidth]{figures/figure02}
+\end{center}
+\textbf{\refstepcounter{figure}\label{fig:figure2} Figure \arabic{figure}. }%
+{Fig 2 caption}
 \end{figure}

### 2015-06-19 paper/main.tex
add fig 2, improve draft

@@
 \linenumbers
-\section{Introduction}
+\section*{Introduction}
 The assessment of tractography from \gls*{dmri} data and subsequent analyses is challenging due to
@@
 
-\section{Data description}
+\section*{Data description}
 
@@
   described in \autoref{fig:figure1}.
-\emph{Diffantom} observes two major improvements with respect to \citep{wilkins_fiber_2014}.
+\emph{Diffantom} integrates two major improvements with respect to \citep{wilkins_fiber_2014}.
 First, since we use a dataset from the \gls*{hcp} as input, data are corrected for the most relevant distortions.
-\cite{wilkins_fiber_2014} explicitly state that their original data did not subjected to certain distortion
+\cite{wilkins_fiber_2014} explicitly state that their original data were not subjected to certain distortion
   corrections, and thus, generated data are affected correspondingly.
-Second improvement is methodological, as we use a more complete signal model to generate the phantom thanks
-  to \emph{phantomas} \citep{caruyer_phantomas_2014}.
+Second improvement is a more advanced signal model to generate the phantom using
+  \emph{phantomas} \citep{caruyer_phantomas_2014}.
+{\color{red} Insert here relaxation parameters and diffusion parameters used in simulation. Mention the gradient
+scheme used by default --150 dirs, 2 shells using \citep{caruyer_design_2013}--, and that it may be changed by
+the user. Mention noise, and default snr.}
 
@@
   we also distribute the software instrumentation presented in \autoref{fig:figure1}.
+\autoref{fig:figure2}, block A, describes the recommended use case of \emph{diffantom} in the evaluation
+  of algorithms and/or pipelines.
+Finally, \autoref{fig:figure2}, blocks B1-3, show some tractograms of fundamental bundles to illustrate
+  the features of the dataset.
+To obtain the whole-brain tractogram, we used the example dataset in a connectivity pipeline including
+  constrained spherical deconvolution and probabilistic tractography from
+  \emph{MRTrix} \citep{tournier_mrtrix_2012}.
+{\color{red} specify here more parameters (harmonics, etc)}
 
-\section{Discussion and conclusion}
+\section*{Discussion and conclusion}
 
 \noindent\textbf{\textit{Discussion\textcolon}}\label{sec:discussion} %
-Whole-brain, realistic \gls*{dmri} phantoms are necessary in current development of connectomics.
-The \emph{diffantom} is a derivative work from \citep{wilkins_fiber_2014} in terms of methodology for
+Whole-brain, realistic \gls*{dmri} phantoms are necessary in the developing field of structural
+  connectomics.
+The \emph{diffantom} is a derivative of \citep{wilkins_fiber_2014} in terms of methodology for
   simulation with two major improvements: the correctness of the \emph{minimally preprocessed} data
@@
   and operation time.
+Since the gradient scheme can be set by the user, \emph{diffantom} can be seen as a mean to convert the gradient
+  scheme of the \gls*{hcp} to any target scheme.
 
@@
 \section*{Data Sharing}
+{\color{red} Add here the pointers to data (dryad?) and software (github/zenodo doi) when available and say licences.}
 
@@
 \textbf{\refstepcounter{figure}\label{fig:figure2} Figure \arabic{figure}. }%
-{Fig 2 caption}
+{\textbf{A. Recommended use of \emph{diffantom}}.
+The dataset can be used to test algorithms and pipelines against a perturbation that is
+  introduced in \emph{diffantom}.
+Perturbations can model typical artifacts found in \gls*{dmri} datasets or pathological conditions
+  such as a tumor.
+\textbf{B1, B2, B3. Tractograms of fundamental fiber bundles}.
+}
 \end{figure}

### 2015-06-19 paper/main.tex
fix references

@@
   a large number of bundles.
-Therefore, \emph{diffantom} is complementary as the hypotheses that can be investigated are different and
-  it does not require costly manual segmentation of bundles, highly demanding in terms of physiology expertise
-  and operation time.
-Since the gradient scheme can be set by the user, \emph{diffantom} can be seen as a mean to convert the gradient
-  scheme of the \gls*{hcp} to any target scheme.
+Therefore, \emph{diffantom} and \emph{FiberFox} are complementary as the hypotheses that can be investigated are different.
+Moreover, \emph{diffantom} does not require costly manual segmentation of bundles, highly demanding in terms of physiology
+  expertise and operation time.
+This facilitates the use of \emph{diffantom} as a factory of whole-brain diffusion phantoms.
+Lastly, since the gradient scheme can be set by the user, \emph{diffantom} can be seen as a mean to translate the so-called
+  \emph{b-matrix} of the \gls*{hcp} to any target scheme.
 

### 2015-06-23 paper/main.tex
update paper

@@
 \def\Address{%
-$^{1}$Biomedical Image Technologies (BIT), ETSI Telecomunicaci\'on, Universidad Polit\'ecnica de Madrid, Madrid, Spain \\
+$^{1}$Biomedical Image Technologies (BIT), ETSI Telecom., Universidad Polit\'ecnica de Madrid, Madrid, Spain \\
 $^{2}$Centro de Investigaci\'on Biom\'edica en Red en Bioingenier\'ia, Biomateriales y Nanomedicina (CIBER-BBN), Spain \\

### 2015-06-23 diffantom/__init__.py

@@ -3,4 +3,6 @@
 # @Author: oesteban
 # @Date:   2015-06-15 14:50:01
 # @Last Modified by:   oesteban
-# @Last Modified time: 2015-06-15 14:50:01
+# @Last Modified time: 2015-06-23 12:31:14
+
+from interfaces import PhantomasSticksSim

### 2015-06-23 diffantom/interfaces.py
add code files

@@ -1,0 +1,237 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+# @Author: oesteban
+# @Date:   2015-06-23 12:29:58
+# @Last Modified by:   oesteban
+# @Last Modified time: 2015-06-23 12:39:15
+
+import os
+import os.path as op
+import warnings
+import numpy as np
+import nibabel as nb
+from math import exp
+from nipype.interfaces.base import (TraitedSpec, File, InputMultiPath,
+                                    OutputMultiPath, Undefined, traits,
+                                    isdefined, OutputMultiPath,
+                                    CommandLineInputSpec, CommandLine,
+                                    BaseInterface, BaseInterfaceInputSpec,
+                                    traits)
+from nipype.utils.filemanip import split_filename, fname_presuffix
+from utils import sigmoid_filter
+
+from nipype import logging
+iflogger = logging.getLogger('interface')
+
+
+class PhantomasSticksSimInputSpec(CommandLineInputSpec):
+    in_dirs = InputMultiPath(
+        File(exists=True), mandatory=True, argstr='--sticks_dir %s',
+        desc='list of fibers (principal directions)')
+    in_frac = InputMultiPath(
+        File(exists=True), mandatory=True, argstr='--sticks_vfs %s',
+        desc=('volume fraction of each fiber'))
+    in_vfms = InputMultiPath(
+        File(exists=True), mandatory=True, argstr='--tissue_vf %s',
+        desc=('volume fractions of isotropic compartiments'))
+
+    in_bvec = File(exists=True, argstr='-r %s',
+                   mandatory=True, desc='input bvecs file')
+    in_bval = File(exists=True, argstr='-b %s',
+                   mandatory=True, desc='input bvals file')
+
+    snr = traits.Int(100, argstr='--snr %f', usedefault=True,
+                     desc='signal-to-noise ratio (dB)')
+
+    diff_csf = traits.Float(3.0e-3, usedefault=True, argstr='--diff_csf %f',
+                            desc='csf diffusivity')
+    diff_gm = traits.Float(.7e-3, usedefault=True, argstr='--diff_gm %f',
+                           desc='gray matter diffusivity')
+    diff_wm = traits.Float(.9e-3, usedefault=True, argstr='--diff_wm %f',
+                           desc='gray matter diffusivity')
+
[…]

### 2015-06-23 diffantom/utils.py

@@ -1,0 +1,155 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+# @Author: oesteban
+# @Date:   2015-06-23 12:32:16
+# @Last Modified by:   oesteban
+# @Last Modified time: 2015-06-23 12:36:21
+
+
+def _sim_mask(in_file):
+    import nibabel as nb
+    import numpy as np
+    import os.path as op
+    import scipy.ndimage as sn
+    from scipy.ndimage.morphology import (binary_opening, binary_dilation)
+    from pyacwereg.misc import ball as gen_ball
+
+    if not isinstance(in_file, basestring):
+        in_file = in_file[2]
+
+    out_file = op.abspath('sim_mask.nii.gz')
+
+    im = nb.load(in_file)
+    data = im.get_data()
+    data[data > 1.0e-4] = 1
+    data[data < 1] = 0
+
+    ball1 = gen_ball(4, 1.9)
+    data = binary_opening(data.astype(np.uint8), structure=ball1,
+                          iterations=1).astype(np.uint8)
+    # Get largest object
+    label_im, nb_labels = sn.label(data)
+    sizes = sn.sum(data, label_im, range(nb_labels + 1))
+    larger = np.squeeze(np.argwhere(sizes == sizes.max()))
+    data[label_im != larger] = 0
+
+    # Dilate
+    # data = binary_dilation(data, structure=ball1,
+    #                        iterations=1).astype(np.uint8)
+
+    nb.Nifti1Image(
+        data, im.get_affine(), im.get_header()).to_filename(out_file)
+    return out_file
+
+
+def fix_tensors(in_file):
+    import os.path as op
+    import nibabel as nb
+    import numpy as np
+    out_file = op.abspath('tensors_fixed.nii.gz')
+    im = nb.load(in_file)
+    data = im.get_data()
+    data[..., 0] *= -1.0
+    # data[..., 2] *= -1.0
+
+    nb.Nifti1Image(
+        data, im.get_affine(), im.get_header()).to_filename(out_file)
+    return out_file
+
+
+def compute_fractions(sf_vfs, tissue_vfs, max_fa=0.80):
+    import os.path as op
+    import nibabel as nb
+    import numpy as np
+    import scipy.ndimage as sn
[…]

### 2015-06-23 diffantom/workflows.py

@@ -1,0 +1,153 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+# @Author: oesteban
+# @Date:   2015-06-23 12:32:07
+# @Last Modified by:   oesteban
+# @Last Modified time: 2015-06-23 12:37:25
+
+import os
+import os.path as op
+import numpy as np
+import nipype.pipeline.engine as pe             # pipeline engine
+
+from nipype.interfaces import utility as niu         # utility
+from nipype.interfaces import fsl                    # fsl
+from nipype.interfaces import freesurfer as fs       # freesurfer
+from nipype.interfaces.dipy import Denoise
+
+import utils as pu
+from interfaces import PhantomasSticksSim, LoadSamplingScheme, SigmoidFilter
+
+
+def simulate(name='SimDWI', icorr=True, btable=False):
+    in_fields = ['fibers', 'fractions', 'in_mask']
+    if icorr:
+        in_fields += ['jacobian']
+
+    if btable:
+        in_fields += ['bval', 'bvec']
+    else:
+        in_fields += ['scheme']
+
+    inputnode = pe.Node(niu.IdentityInterface(fields=in_fields),
+                        name='inputnode')
+    outputnode = pe.Node(niu.IdentityInterface(
+        fields=['dwi', 'bvec', 'bval', 'out_mask', 'out_fods']),
+        name='outputnode')
+
+    split = pe.Node(niu.Split(splits=[3, 3]), name='SplitFractions')
+    sch = pe.Node(LoadSamplingScheme(bvals=[1000, 3000]), name='LoadScheme')
+    simdwi = pe.Node(PhantomasSticksSim(
+        diff_gm=1.0e-3, diff_wm=0.9e-3, diff_csf=3.0e-3,
+        lambda1=2.2e-3, lambda2=.2e-3, snr=90, save_fods=True),
+        name='SimulateDWI')
+
+    wf = pe.Workflow(name=name)
+    wf.connect([
+        (inputnode, split,      [('fractions', 'inlist')]),
+        (split,     simdwi,     [('out1', 'in_frac'),
+                                 ('out2', 'in_vfms')]),
+        (inputnode, simdwi,     [('fibers', 'in_dirs')]),
+        (simdwi,    outputnode, [('out_mask', 'out_mask'),
+                                 ('out_fods', 'out_fods')])
+    ])
+
+    if icorr:
+        split = pe.Node(fsl.utils.Split(dimension='t'), name='Split_DWIs')
[…]

### 2015-06-23 paper/main.tex
add licenses

@@
 \section*{Data Sharing}
-{\color{red} Add here the pointers to data (dryad?) and software (github/zenodo doi) when available and say licences.}
+{\color{red} Add here the pointers to data (dryad?)}
+Source code of \emph{diffantom} is available at \url{https://github.com/oesteban/phantomas} under an MIT licence.
+\emph{Phantomas} is available in \url{https://github.com/ecaruyer/phantomas/} under the revised-BSD licence.
 

### 2015-06-23 diffantom/__init__.py
add new interface

@@ -2,7 +2,7 @@
 # -*- coding: utf-8 -*-
 # @Author: oesteban
 # @Date:   2015-06-15 14:50:01
-# @Last Modified by:   oesteban
-# @Last Modified time: 2015-06-23 12:31:14
+# @Last Modified by:   Oscar Esteban
+# @Last Modified time: 2015-06-23 13:05:13
 
-from interfaces import PhantomasSticksSim
+from interfaces import PhantomasSticksSim, LoadSamplingScheme

### 2015-06-23 paper/main.tex
add details missing in paper

@@
   \emph{phantomas} \citep{caruyer_phantomas_2014}.
-{\color{red} Insert here relaxation parameters and diffusion parameters used in simulation. Mention the gradient
-scheme used by default --150 dirs, 2 shells using \citep{caruyer_design_2013}--, and that it may be changed by
-the user. Mention noise, and default snr.}
+By default, diffusion data are generated using a scheme of 100 directions distributed in two shells with uniform
+  coverage \citep{caruyer_design_2013}.
+Custom schemes can be generated supplying the tables of corresponding $b$-vectors and $b$-values.
+Rician noise is also included in \emph{phantomas}, and can be set by the user.
+The default value for \gls*{snr} is pre-set to 90.0.
 
@@
   we also distribute the software instrumentation presented in \autoref{fig:figure1}.
+To ensure reproducibility and usability, the simulation pipeline is written using
+  \emph{nipype} \citep{gorgolewski_nipype_2011}.
 \autoref{fig:figure2}, block A, describes the recommended use case of \emph{diffantom} in the evaluation
@@
   the features of the dataset.
-To obtain the whole-brain tractogram, we used the example dataset in a connectivity pipeline including
-  constrained spherical deconvolution and probabilistic tractography from
+To obtain the whole-brain tractogram, the example dataset underwent a simplified connectivity pipeline including
+  \gls*{csd} and probabilistic tractography from
   \emph{MRTrix} \citep{tournier_mrtrix_2012}.
-{\color{red} specify here more parameters (harmonics, etc)}
+\Gls*{csd} was configured with up to 8 spherical harmonics, and tractography with 1.6$\times$10$^\text{6}$
+  seed points evenly distributed across a dilated mask of the \gls*{wm} tissue.
 
@@
 \section*{Data Sharing}
-{\color{red} Add here the pointers to data (dryad?)}
+The example dataset is available under the \gls*{cc0} in {\color{red} (dryad?)}.
 Source code of \emph{diffantom} is available at \url{https://github.com/oesteban/phantomas} under an MIT licence.
@@
 Second, the fiber orientation maps $\{V_1, V_2, V_3\}$ and corresponding volume fraction maps $\{F_1, F_2, F_3\}$ are
-  obtained using the Multi-Shell ball-and-stick model of BEDPOSTX \citep{jbabdi_modelbased_2012}.
+  obtained using the ball-and-stick model for multi-shell data of BEDPOSTX \citep{jbabdi_modelbased_2012}.
 The third step takes the fiber orientations $\{V_1, V_2, V_3\}$ and their weights $\{F_1, F_2, F_3\}$ to compute the

### 2015-06-25 diffantom/diffantom.py
starting workflow

@@ -1,0 +1,67 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+# @Author: Oscar Esteban
+# @Date:   2015-06-25 15:46:08
+# @Last Modified by:   Oscar Esteban
+# @Last Modified time: 2015-06-25 16:51:04
+
+"""
+============================
+The Diffantom software layer
+============================
+"""
+import os
+import os.path as op
+import sys
+
+
+__author__ = "Oscar Esteban"
+__copyright__ = "Copyright 2015, Biomedical Image Technologies (BIT), \
+                 Universidad Politécnica de Madrid"
+__credits__ = "Oscar Esteban"
+__license__ = "MIT"
+__version__ = "0.1"
+__maintainer__ = "Oscar Esteban"
+__email__ = "code@oscaresteban.es"
+__status__ = "Prototype"
+
+
+if __name__ == '__main__':
+    from argparse import ArgumentParser
+    from argparse import RawTextHelpFormatter
+    from workflows import gen_diffantom
+    import glob
+    import os
+    import os.path as op
+    from shutil import copyfileobj
+
+    parser = ArgumentParser(description='Preprocessing dMRI routine',
+                            formatter_class=RawTextHelpFormatter)
+
+    g_input = parser.add_argument_group('Input')
+    g_input.add_argument('-S', '--subjects_dir', action='store',
+                         default=os.getenv('NEURO_DATA_HOME',
+                                           op.abspath('../')),
+                         help='directory where subjects should be found')
+    g_input.add_argument('-s', '--subject', action='store',
+                         default='S*', help='subject id or pattern')
+    g_input.add_argument('-w', '--work_dir', action='store',
+                         default=os.getcwd(),
+                         help='directory to store intermediate results')
+
+    g_input.add_argument(
+        '-N', '--name', action='store', default='Diffantom',
+        help=('default workflow name, it will create a new folder'))
+
+    opts = parser.parse_args()
+
+    settings = {}
+    settings['work_dir'] = opts.work_dir
+    settings['data_dir'] = opts.subjects_dir
+    settings['subject_id'] = opts.subject
+
+    wf = gen_diffantom(opts.name, settings=settings)
+
+    wf.base_dir = settings['work_dir']
+    # wf.write_graph(format='pdf')
+    wf.run()

### 2015-06-25 diffantom/workflows.py

@@ -2,8 +2,8 @@
 # -*- coding: utf-8 -*-
 # @Author: oesteban
 # @Date:   2015-06-23 12:32:07
-# @Last Modified by:   oesteban
-# @Last Modified time: 2015-06-23 12:37:25
+# @Last Modified by:   Oscar Esteban
+# @Last Modified time: 2015-06-25 17:01:20
 
 import os
 import os.path as op
@@ -11,6 +11,7 @@ import numpy as np
 import nipype.pipeline.engine as pe             # pipeline engine
 
 from nipype.interfaces import utility as niu         # utility
+from nipype.interfaces import io as nio
 from nipype.interfaces import fsl                    # fsl
 from nipype.interfaces import freesurfer as fs       # freesurfer
 from nipype.interfaces.dipy import Denoise
@@ -19,15 +20,59 @@ import utils as pu
 from interfaces import PhantomasSticksSim, LoadSamplingScheme, SigmoidFilter
 
 
-def simulate(name='SimDWI', icorr=True, btable=False):
-    in_fields = ['fibers', 'fractions', 'in_mask']
-    if icorr:
-        in_fields += ['jacobian']
+def gen_diffantom(name='Diffantom', settings={}):
+    sgm_structures = ['L_Accu', 'R_Accu', 'L_Caud', 'R_Caud',
+                      'L_Pall', 'R_Pall', 'L_Puta', 'R_Puta',
+                      'L_Thal', 'R_Thal']
 
-    if btable:
-        in_fields += ['bval', 'bvec']
-    else:
-        in_fields += ['scheme']
+    inputnode = pe.Node(niu.IdentityInterface(
+        fields=['subject_id', 'data_dir']), name='inputnode')
+    inputnode.inputs.subject_id = settings['subject_id']
+    inputnode.inputs.data_dir = settings['data_dir']
+
+    fnames = dict(t1w='T1w_acpc_dc_restore.nii.gz',
+                  fibers='fiber*.nii.gz',
+                  vfractions='vfraction*.nii.gz',
+                  scheme='samples.txt')
+
+    ds_tpl_args = {k: [['subject_id', [v]]] for k, v in fnames.iteritems()}
+
+    ds = pe.Node(nio.DataGrabber(
+        infields=['subject_id'], sort_filelist=True, template='*',
+        outfields=ds_tpl_args.keys()), name='DataSource')
+    ds.inputs.field_template = {k: 'models/%s/%s'
[…]

### 2015-06-25 diffantom/utils.py
add surfaces of deep gray matter

@@ -2,8 +2,8 @@
 # -*- coding: utf-8 -*-
 # @Author: oesteban
 # @Date:   2015-06-23 12:32:16
-# @Last Modified by:   oesteban
-# @Last Modified time: 2015-06-23 12:36:21
+# @Last Modified by:   Oscar Esteban
+# @Last Modified time: 2015-06-25 20:49:48
 
 
 def _sim_mask(in_file):
@@ -153,3 +153,56 @@ def sigmoid_filter(data, mask=None, a=2.00, b=85.0, maxout=2000.0):
     newdata = np.zeros_like(data)
     newdata[idxs] = ddor
     return newdata
+
+
+def fixvtk(in_file, in_ref, out_file=None):
+    """
+    Translates surfaces to the corresponding origin
+    """
+    import nibabel as nb
+    import numpy as np
+    import os.path as op
+    import subprocess as sp
+
+    if out_file is None:
+        fname, ext = op.splitext(op.basename(in_file))
+        if ext == ".gz":
+            fname, _ = op.splitext(fname)
+        out_file = op.abspath("%s_fixed.vtk" % fname)
+
+    im = nb.load(in_ref)
+    aff = im.get_affine()
+    mni_orig = aff[:3, 3]
+    flip = np.ones(3, dtype=np.float32)
+    flip[mni_orig > 0.0] = -1.0
+
+    with open(in_file, 'r') as f:
+        with open(out_file, 'w+') as w:
+            npoints = 0
+            pointid = -5
+
+            w.write("# vtk DataFile Version 1.0\n"
+                    "vtk output\n"
+                    "ASCII\n"
+                    "DATASET POLYDATA\n")
+
+            for i, l in enumerate(f):
+                if (i == 4):
+                    s = l.split()
+                    npoints = int(s[1])
+                    fmt = np.dtype(s[2])
+                    w.write(l)
+                elif ((i > 4) and (pointid < npoints)):
+                    vert = np.array([float(x) for x in l.split()])
+                    newvert = vert * flip + mni_orig
+                    l = '%.5f  %.5f  %.5f\n' % tuple(newvert)
+                    w.write(l)
+                elif pointid == npoints:
+                    s = l.split()
+                    w.write("POLYGONS %d %d\n" % (int(s[1]), int(s[2])))
+                elif pointid > npoints:
+                    w.write(l)
+
+                pointid += 1
+
+    return out_file

### 2015-06-25 diffantom/workflows.py

@@ -3,7 +3,7 @@
 # @Author: oesteban
 # @Date:   2015-06-23 12:32:07
 # @Last Modified by:   Oscar Esteban
-# @Last Modified time: 2015-06-25 17:01:20
+# @Last Modified time: 2015-06-25 18:38:27
 
 import os
 import os.path as op
@@ -18,6 +18,7 @@ from nipype.interfaces.dipy import Denoise
 
 import utils as pu
 from interfaces import PhantomasSticksSim, LoadSamplingScheme, SigmoidFilter
+from pyacwereg.interfaces import Surf2Vol
 
 
 def gen_diffantom(name='Diffantom', settings={}):
@@ -50,11 +51,17 @@ def gen_diffantom(name='Diffantom', settings={}):
     fast = pe.Node(fsl.FAST(number_classes=3, img_type=1, no_bias=True,
                             probability_maps=True), name='SegmentT1')
     first = pe.Node(fsl.FIRST(
-        list_of_specific_structures=sgm_structures, brain_extracted=True),
-        name='FIRST')
+        list_of_specific_structures=sgm_structures, brain_extracted=True,
+        method='auto'), name='FIRST')
+
+    mesh2pve = pe.Node(Surf2Vol(), name='Mesh2PVE')
 
     sim_mod = preprocess_model()
     sim_ref = simulate()
+
+    fixVTK = pe.MapNode(niu.Function(
+        input_names=['in_file', 'in_ref'], output_names=['out_file'],
+        function=pu.fixvtk), name='fixVTK', iterfield=['in_file'])
 
     wf = pe.Workflow(name=name)
     wf.connect([
@@ -63,10 +70,19 @@ def gen_diffantom(name='Diffantom', settings={}):
         (ds,        bet,      [('t1w', 'in_file')]),
         (ds,        sim_mod,  [('fibers', 'inputnode.fibers'),
                                ('vfractions', 'inputnode.vfractions')]),
+        (ds,        sim_ref,  [('scheme', 'inputnode.scheme')]),
         (bet,       sim_mod,  [('mask_file', 'inputnode.in_mask')]),
         (bet,       fast,     [('out_file', 'in_files')]),
         (fast,      sim_mod,  [('partial_volume_files', 'inputnode.in_tpms')]),
-        (bet,       first,    [('out_file', 'in_file')])
+        (bet,       first,    [('out_file', 'in_file')]),
[…]

### 2015-06-25 diffantom/workflows.py
introduce mask resampling

@@ -3,7 +3,7 @@
 # @Author: oesteban
 # @Date:   2015-06-23 12:32:07
 # @Last Modified by:   Oscar Esteban
-# @Last Modified time: 2015-06-25 18:38:27
+# @Last Modified time: 2015-06-25 21:06:56
 
 import os
 import os.path as op
@@ -25,6 +25,9 @@ def gen_diffantom(name='Diffantom', settings={}):
     sgm_structures = ['L_Accu', 'R_Accu', 'L_Caud', 'R_Caud',
                       'L_Pall', 'R_Pall', 'L_Puta', 'R_Puta',
                       'L_Thal', 'R_Thal']
+
+    def _getfirst(inlist):
+        return inlist[0]
 
     inputnode = pe.Node(niu.IdentityInterface(
         fields=['subject_id', 'data_dir']), name='inputnode')
@@ -53,9 +56,8 @@ def gen_diffantom(name='Diffantom', settings={}):
     first = pe.Node(fsl.FIRST(
         list_of_specific_structures=sgm_structures, brain_extracted=True,
         method='auto'), name='FIRST')
-
+    reslice = pe.Node(fs.MRIConvert(), name='ResliceMask')
     mesh2pve = pe.Node(Surf2Vol(), name='Mesh2PVE')
-
     sim_mod = preprocess_model()
     sim_ref = simulate()
 
@@ -71,7 +73,9 @@ def gen_diffantom(name='Diffantom', settings={}):
         (ds,        sim_mod,  [('fibers', 'inputnode.fibers'),
                                ('vfractions', 'inputnode.vfractions')]),
         (ds,        sim_ref,  [('scheme', 'inputnode.scheme')]),
-        (bet,       sim_mod,  [('mask_file', 'inputnode.in_mask')]),
+        (bet,       reslice,  [('mask_file', 'in_file')]),
+        (ds,        reslice,  [(('vfractions', _getfirst), 'reslice_like')]),
+        (reslice,   sim_mod,  [('out_file', 'inputnode.in_mask')]),
         (bet,       fast,     [('out_file', 'in_files')]),
         (fast,      sim_mod,  [('partial_volume_files', 'inputnode.in_tpms')]),
         (bet,       first,    [('out_file', 'in_file')]),

### 2015-06-26 diffantom/workflows.py
fix workflow, add examples

@@ -3,7 +3,7 @@
 # @Author: oesteban
 # @Date:   2015-06-23 12:32:07
 # @Last Modified by:   Oscar Esteban
-# @Last Modified time: 2015-06-25 21:06:56
+# @Last Modified time: 2015-06-26 13:36:10
 
 import os
 import os.path as op
@@ -80,7 +80,7 @@ def gen_diffantom(name='Diffantom', settings={}):
         (fast,      sim_mod,  [('partial_volume_files', 'inputnode.in_tpms')]),
         (bet,       first,    [('out_file', 'in_file')]),
         (first,     fixVTK,   [('vtk_surfaces', 'in_file')]),
-        (ds,        fixVTK,   [('t1w', 'in_ref')]),
+        (ds,        fixVTK,   [(('vfractions', _getfirst), 'in_ref')]),
         (ds,        mesh2pve, [('t1w', 'reference')]),
         (fixVTK,    mesh2pve, [('out_file', 'surfaces')]),
         (bet,       sim_ref,  [('mask_file', 'inputnode.in_mask')]),

### 2015-07-06 diffantom/workflows.py
build matrix, track CC bundles

@@ -2,8 +2,8 @@
 # -*- coding: utf-8 -*-
 # @Author: oesteban
 # @Date:   2015-06-23 12:32:07
-# @Last Modified by:   oesteban
-# @Last Modified time: 2015-07-03 15:43:54
+# @Last Modified by:   Oscar Esteban
+# @Last Modified time: 2015-07-06 11:47:06
 
 import os
 import os.path as op
@@ -16,6 +16,7 @@ from nipype.interfaces import fsl                    # fsl
 from nipype.interfaces import freesurfer as fs       # freesurfer
 from nipype.interfaces.dipy import Denoise
 from nipype.interfaces import mrtrix3 as mrt3
+from nipype.interfaces.mrtrix import MRTrix2TrackVis as TCK2TRK
 
 import utils as pu
 from interfaces import PhantomasSticksSim, LoadSamplingScheme, SigmoidFilter
@@ -32,7 +33,8 @@ def gen_diffantom(name='Diffantom', settings={}):
     fnames = dict(t1w='T1w_acpc_dc_restore.nii.gz',
                   fibers='fiber*.nii.gz',
                   vfractions='vfraction*.nii.gz',
-                  scheme='samples.txt')
+                  scheme='samples.txt',
+                  aparc='aparc+aseg.nii.gz')
 
     ds_tpl_args = {k: [['subject_id', [v]]] for k, v in fnames.iteritems()}
 
@@ -53,7 +55,9 @@ def gen_diffantom(name='Diffantom', settings={}):
                                ('data_dir', 'base_directory')]),
         (ds,        sim_mod,  [('t1w', 'inputnode.t1w'),
                                ('fibers', 'inputnode.fibers'),
-                               ('vfractions', 'inputnode.fractions')]),
+                               ('vfractions', 'inputnode.fractions'),
+                               ('aparc', 'inputnode.parcellation')]),
+        (ds,        trk,      [('aparc', 'inputnode.aparc')]),
         (ds,        sim_ref,  [('scheme', 'inputnode.scheme')]),
         (sim_mod,   sim_ref,     [
             ('outputnode.fibers', 'inputnode.fibers'),
@@ -63,7 +67,9 @@ def gen_diffantom(name='Diffantom', settings={}):
         (sim_ref,   trk,      [
             ('outputnode.dwi', 'inputnode.in_dwi'),
[…]

### 2015-07-06 diffantom/workflows.py
add regions to see if include works

@@ -3,7 +3,7 @@
 # @Author: oesteban
 # @Date:   2015-06-23 12:32:07
 # @Last Modified by:   Oscar Esteban
-# @Last Modified time: 2015-07-06 11:47:06
+# @Last Modified time: 2015-07-06 19:09:08
 
 import os
 import os.path as op
@@ -264,7 +264,11 @@ def act_workflow(name='Tractography'):
                   name='ComputeTDI')
 
     bnd0 = track_bundle('Bundle_CC')
-    bnd0.inputs.inputnode.labels = [251]
+    bnd0.inputs.inputnode.seed_lbs = [251]
+
+    labels = [1002, 1012, 1014, 1026, 1027, 1028, 1032]
+    bnd0.inputs.inputnode.include_lbs = (np.array(labels) + 1000).tolist() + \
+        labels
 
     wf = pe.Workflow(name=name)
     wf.connect([
@@ -310,14 +314,15 @@ def act_workflow(name='Tractography'):
 
 def track_bundle(name='BundleTrack'):
     inputnode = pe.Node(niu.IdentityInterface(
-        fields=['in_fod', 'aparc', 'labels', 'parcellation',
-                'in_5tt', 'in_scheme']),
+        fields=['in_fod', 'aparc', 'seed_lbs', 'parcellation',
+                'in_5tt', 'in_scheme', 'include_lbs']),
         name='inputnode')
     outputnode = pe.Node(niu.IdentityInterface(
         fields=['out_file']),
         name='outputnode')
 
     msk = pe.Node(fs.Binarize(), name='Binarize')
+    inc = pe.Node(fs.Binarize(), name='BinarizeInclude')
 
     trk = pe.Node(mrt3.Tractography(
         nthreads=0, n_tracks=int(1e6), max_length=250.), name='Track')
@@ -325,13 +330,16 @@ def track_bundle(name='BundleTrack'):
 
     wf = pe.Workflow(name=name)
     wf.connect([
-        (inputnode, tck2trk,    [('parcellation', 'image_file')]),
+        (inputnode, tck2trk,    [('aparc', 'image_file')]),
+        (inputnode, inc,        [('aparc', 'in_file'),
+                                 ('include_lbs', 'match')]),
         (inputnode, msk,        [('aparc', 'in_file'),
-                                 ('labels', 'match')]),
+                                 ('seed_lbs', 'match')]),
         (inputnode, trk,        [('in_fod', 'in_file'),
                                  ('in_5tt', 'act_file'),
[…]

### 2015-07-08 diffantom/utils.py
new fractions F_i computation methodology

@@ -2,8 +2,8 @@
 # -*- coding: utf-8 -*-
 # @Author: oesteban
 # @Date:   2015-06-23 12:32:16
-# @Last Modified by:   Oscar Esteban
-# @Last Modified time: 2015-07-02 18:12:58
+# @Last Modified by:   oesteban
+# @Last Modified time: 2015-07-08 11:41:05
 
 
 def _sim_mask(in_file):
@@ -116,6 +116,89 @@ def compute_fractions(sf_vfs, tissue_vfs, max_fa=0.80):
 
     out_wmmsk = op.abspath('wmmsk.nii.gz')
     nb.Nifti1Image(wmmask, sfimgs[0].get_affine(), hdr).to_filename(out_wmmsk)
+
+    return out_sf, out_ts, out_wmmsk
+
+
+def fa_fractions(fa, sf_vfs, tissue_vfs):
+    import os.path as op
+    import nibabel as nb
+    import numpy as np
+    import scipy.ndimage as sn
+    from pyacwereg.misc import ball as gen_ball
+    from scipy.ndimage.morphology import binary_erosion
+
+    # Load tissue fractions
+    ntissues = len(tissue_vfs)
+    timgs = [nb.load(f) for f in tissue_vfs]
+    tvfs = np.nan_to_num(nb.concat_images(timgs).get_data())
+    tissue_totals = np.sum(tvfs, axis=3)
+    tvfs[tissue_totals > 0, ...] /= tissue_totals[tissue_totals > 0,
+                                                  np.newaxis]
+    wmvf = tvfs[..., 2].copy()
+    tvfs[..., 2] *= .05
+    wmvf -= tvfs[..., 2]
+
+    cgmvf = tvfs[..., 0].copy()
+    tvfs[..., 0] *= .75
+    cgmvf -= tvfs[..., 0]
+
+    dgmvf = tvfs[..., 1].copy()
+    tvfs[..., 1] *= .50
+    dgmvf -= tvfs[..., 1]
+
+    # Load fa
+    faimg = nb.load(fa)
+    fadata = faimg.get_data()
+
+    nfibers = 3
+    data = np.zeros(tuple(list(fadata.shape) + [nfibers]))
+
+    # Load 3rd fiber map
+    sf3vf = nb.load(sf_vfs[2]).get_data() * wmvf
+
+    resvf = dgmvf + cgmvf
+
[…]

### 2015-07-08 diffantom/workflows.py

@@ -3,7 +3,7 @@
 # @Author: oesteban
 # @Date:   2015-06-23 12:32:07
 # @Last Modified by:   Oscar Esteban
-# @Last Modified time: 2015-07-06 19:09:08
+# @Last Modified time: 2015-07-08 11:41:20
 
 import os
 import os.path as op
@@ -34,7 +34,8 @@ def gen_diffantom(name='Diffantom', settings={}):
                   fibers='fiber*.nii.gz',
                   vfractions='vfraction*.nii.gz',
                   scheme='samples.txt',
-                  aparc='aparc+aseg.nii.gz')
+                  aparc='aparc+aseg.nii.gz',
+                  fa='fa.nii.gz')
 
     ds_tpl_args = {k: [['subject_id', [v]]] for k, v in fnames.iteritems()}
 
@@ -56,7 +57,8 @@ def gen_diffantom(name='Diffantom', settings={}):
         (ds,        sim_mod,  [('t1w', 'inputnode.t1w'),
                                ('fibers', 'inputnode.fibers'),
                                ('vfractions', 'inputnode.fractions'),
-                               ('aparc', 'inputnode.parcellation')]),
+                               ('aparc', 'inputnode.parcellation'),
+                               ('fa', 'inputnode.in_fa')]),
         (ds,        trk,      [('aparc', 'inputnode.aparc')]),
         (ds,        sim_ref,  [('scheme', 'inputnode.scheme')]),
         (sim_mod,   sim_ref,     [
@@ -108,7 +110,7 @@ def simulate(name='SimDWI'):
 
 
 def preprocess_model(name='PrepareModel'):
-    in_fields = ['t1w', 'fibers', 'fractions', 'parcellation']
+    in_fields = ['t1w', 'fibers', 'fractions', 'parcellation', 'in_fa']
     sgm_structures = ['L_Accu', 'R_Accu', 'L_Caud', 'R_Caud',
                       'L_Pall', 'R_Pall', 'L_Puta', 'R_Puta',
                       'L_Thal', 'R_Thal']
@@ -183,6 +185,11 @@ def preprocess_model(name='PrepareModel'):
         os.getenv('MRTRIX3_HOME', '/home/oesteban/workspace/mrtrix3'),
         'src/dwi/tractography/connectomics/example_configs/fs_default.txt')
 
+    faden = pe.Node(Denoise(snr=90.0), name='FADenoise')
+    fapst = pe.Node(niu.Function(
[…]

### 2015-07-09 paper/main.tex
add new intro

@@
 \linenumbers
+
 \section*{Introduction}
+Fiber tracking on \gls*{dmri} data has become an important application in the in-vivo investigation
+  of the structural configuration of fiber bundles at the macro-scale.
+The final outcome of such applications, called tractography, is fundamental to gain information about
+  \gls*{wm} morphology in neurosurgical planning, post-surgery evaluations, and the study of neurological
+  diseases such as multiple sclerosis, and Alzheimer's disease \citep{chua_diffusion_2008}.
+Tractography is also crucial in the analysis of structural brain networks using graph theory.
+These analyses have been successfully applied in the definition of the unique subject-wise patterns of connectivity
+  \citep{sporns_human_2005}, the assessment of neurological diseases \citep{griffa_structural_2013}, and in the
+  study of the link of structural and functional connectivity \citep{messe_predicting_2015}.
+However, the development of the field is limited by the lack of gold standards to test and compare the
+  wide range of algorithms and processing methodologies available.
+In this work we propose \emph{diffantom}, an \emph{in-silico} dataset to assess tractography pipelines using
+  real datasets as source of whole-brain micro-structural information.
+We also provide a full software layer to produce \emph{diffantoms} from real \gls*{dmri} datasets
+  drawn from the \gls*{hcp}.
+This dataset enables the following evaluations of standard connectivity pipelines:
+\begin{itemize}
+  \item Comparison of the impact of the different diffusion sampling schemes on the local micro-structure
+  model of choice and the subsequent global tractography outcome.
+  \item Investigation of the effects of the different artifacts (noise, susceptibility-derived, Eddy-currents,
+  etc.) on the final connectivity matrix.
+  \item Determine of the \acrfull*{roc} of connectivity pipelines.
+  \item Generation of synthetic connectivity networks to test graph theory analyses.
+\end{itemize}
+
+\section*{Background}
 The assessment of tractography from \gls*{dmri} data and subsequent analyses is challenging due to
@@
   methodology very close to that implemented in \emph{Phantomas}.
-
-In this paper we present \emph{Diffantom}, a software layer to generate diffusion phantoms at whole-brain
-  scale and derived from real \gls*{dmri} datasets.
-Particularly, we propose the \gls*{hcp} \citep{essen_human_2012} as the source of real data.
-Finally, an example dataset is openly released along with all the necessary requirements and instructions
-  to generate custom phantoms.
 

### 2015-07-09 diffantom/utils.py

@@ -2,8 +2,8 @@
 # -*- coding: utf-8 -*-
 # @Author: oesteban
 # @Date:   2015-06-23 12:32:16
-# @Last Modified by:   Oscar Esteban
-# @Last Modified time: 2015-07-02 18:12:58
+# @Last Modified by:   oesteban
+# @Last Modified time: 2015-07-07 10:21:34
 
 
 def _sim_mask(in_file):

### 2015-07-10 diffantom/workflows.py
Workflow to generate whole-brain tractography. close #6

@@ -3,7 +3,7 @@
 # @Author: oesteban
 # @Date:   2015-06-23 12:32:07
 # @Last Modified by:   Oscar Esteban
-# @Last Modified time: 2015-07-08 11:41:20
+# @Last Modified time: 2015-07-10 08:32:24
 
 import os
 import os.path as op
@@ -254,7 +254,7 @@ def act_workflow(name='Tractography'):
         output_names=['out_file']), name='ComputeTrackingMask')
 
     resp = pe.Node(mrt3.ResponseSD(
-        bval_scale='yes', max_sh=8), name='EstimateResponse')
+        bval_scale='yes', max_sh=8, nthreads=2), name='EstimateResponse')
     fod = pe.Node(mrt3.EstimateFOD(
         bval_scale='yes', max_sh=8, nthreads=0), name='EstimateFODs')
     tsr = pe.Node(mrt3.FitTensor(
@@ -262,7 +262,7 @@ def act_workflow(name='Tractography'):
     met = pe.Node(mrt3.TensorMetrics(
         out_adc='adc.nii.gz', out_fa='fa.nii.gz'), name='ComputeScalars')
     trk = pe.Node(mrt3.Tractography(
-        nthreads=0, n_tracks=int(1e8), max_length=250.), name='Track')
+        nthreads=0, n_tracks=int(1e5), max_length=250.), name='Track')
 
     lc = pe.Node(mrt3.LabelConfig(), name='LabelConfig')
     lc.inputs.out_file = 'parcellation.nii.gz'
@@ -272,6 +272,7 @@ def act_workflow(name='Tractography'):
     mat = pe.Node(mrt3.BuildConnectome(nthreads=0), name='BuildMatrix')
 
     # tck2trk = pe.Node(TCK2TRK(), name='TCK2TRK')
+    tck2vtk = pe.Node(mrt3.TCK2VTK(), name='TCK2VTK')
     tdi = pe.Node(mrt3.ComputeTDI(nthreads=0, out_file='tdi.nii.gz'),
                   name='ComputeTDI')
 
@@ -294,6 +295,7 @@ def act_workflow(name='Tractography'):
         (inputnode, lc,         [('parcellation', 'in_file')]),
         (inputnode, excl,       [('aparc', 'in_file')]),
         # (inputnode, tck2trk,    [('in_dwi', 'image_file')]),
+        (inputnode, tck2vtk,    [('parcellation', 'reference')]),
         (inputnode, resp,       [('in_dwi', 'in_file'),
[…]

### 2015-07-10 diffantom/interfaces.py
a first implementation of #8

@@ -2,18 +2,19 @@
 # -*- coding: utf-8 -*-
 # @Author: oesteban
 # @Date:   2015-06-23 12:29:58
-# @Last Modified by:   oesteban
-# @Last Modified time: 2015-07-01 10:59:48
+# @Last Modified by:   Oscar Esteban
+# @Last Modified time: 2015-07-10 13:06:50
 
 import os
 import os.path as op
+import glob
 import warnings
 import numpy as np
 import nibabel as nb
 from math import exp
 from nipype.interfaces.base import (TraitedSpec, File, InputMultiPath,
                                     OutputMultiPath, Undefined, traits,
-                                    isdefined, OutputMultiPath,
+                                    isdefined,
                                     CommandLineInputSpec, CommandLine,
                                     BaseInterface, BaseInterfaceInputSpec,
                                     traits)
@@ -243,4 +244,38 @@ class LoadSamplingScheme(BaseInterface):
                               op.abspath(self.inputs.out_bval)]
         outputs['out_bvec'] = outputs['out_fsl'][0]
         outputs['out_bval'] = outputs['out_fsl'][1]
+        return outputs
+
+
+class TractQuerierInputSpec(CommandLineInputSpec):
+    in_file = File(exists=True, argstr='-t %s', mandatory=True,
+                   desc='input tracks in VTK format')
+    in_parc = File(exists=True, argstr='-a %s', mandatory=True,
+                   desc='input parcellation file')
+    in_queries = File(exists=True, argstr='-q %s', mandatory=True,
+                      desc='input query file')
+
+    out_prefix = traits.Str(
+        'query', argstr='-o %s', mandatory=True, usedefault=True,
+        desc='output prefix')
+
+
+class TractQuerierOutputSpec(TraitedSpec):
+    out_file = File(exists=True, desc='the output response file')
+
+
+class TractQuerier(CommandLine):
+
+    """
+
+    """
+
+    _cmd = 'tract_querier'
+    input_spec = TractQuerierInputSpec
+    output_spec = TractQuerierOutputSpec
+
+    def _list_outputs(self):
+        outputs = self.output_spec().get()
+        outputs['out_file'] = [
+            op.abspath(f) for f in glob.glob('%s*' % self.inputs.out_prefix)]
         return outputs

### 2015-07-10 diffantom/utils.py

@@ -2,8 +2,8 @@
 # -*- coding: utf-8 -*-
 # @Author: oesteban
 # @Date:   2015-06-23 12:32:16
-# @Last Modified by:   oesteban
-# @Last Modified time: 2015-07-08 11:41:05
+# @Last Modified by:   Oscar Esteban
+# @Last Modified time: 2015-07-10 12:45:40
 
 
 def _sim_mask(in_file):
@@ -238,7 +238,7 @@ def sigmoid_filter(data, mask=None, a=2.00, b=85.0, maxout=2000.0):
     return newdata
 
 
-def fixvtk(in_file, in_ref, out_file=None):
+def fixvtk(in_file, in_ref, flips=True, out_file=None):
     """
     Translates surfaces to the corresponding origin
     """
@@ -257,7 +257,9 @@ def fixvtk(in_file, in_ref, out_file=None):
     aff = im.get_affine()
     mni_orig = aff[:3, 3]
     flip = np.ones(3, dtype=np.float32)
-    flip[mni_orig > 0.0] = -1.0
+
+    if flips:
+        flip[mni_orig > 0.0] = -1.0
 
     with open(in_file, 'r') as f:
         with open(out_file, 'w+') as w:
@@ -288,6 +290,28 @@ def fixvtk(in_file, in_ref, out_file=None):
 
                 pointid += 1
 
+    return out_file
+
+
+def fixorigin(in_file, out_file=None):
+    """
+    Fix origin of volume for tract_querier
+    """
+    import nibabel as nb
+    import numpy as np
+    import os.path as op
+    import subprocess as sp
+
+    if out_file is None:
+        fname, ext = op.splitext(op.basename(in_file))
+        if ext == ".gz":
+            fname, _ = op.splitext(fname)
+        out_file = op.abspath("%s_fixed.nii.gz" % fname)
+
+    im = nb.load(in_file)
+    aff = im.get_affine()
+    aff[:3, 3] = 0.0
+    nb.Nifti1Image(im.get_data(), aff, im.get_header()).to_filename(out_file)
     return out_file
 
 

### 2015-07-10 diffantom/workflows.py

@@ -3,7 +3,7 @@
 # @Author: oesteban
 # @Date:   2015-06-23 12:32:07
 # @Last Modified by:   Oscar Esteban
-# @Last Modified time: 2015-07-10 08:32:24
+# @Last Modified time: 2015-07-10 13:07:36
 
 import os
 import os.path as op
@@ -19,7 +19,8 @@ from nipype.interfaces import mrtrix3 as mrt3
 from nipype.interfaces.mrtrix import MRTrix2TrackVis as TCK2TRK
 
 import utils as pu
-from interfaces import PhantomasSticksSim, LoadSamplingScheme, SigmoidFilter
+from interfaces import (PhantomasSticksSim, LoadSamplingScheme, SigmoidFilter,
+                        TractQuerier)
 from pyacwereg.interfaces import Surf2Vol
 
 
@@ -242,7 +243,8 @@ def act_workflow(name='Tractography'):
         fields=['in_dwi', 'in_scheme', 'in_5tt', 'parcellation', 'aparc']),
         name='inputnode')
     outputnode = pe.Node(niu.IdentityInterface(
-        fields=['out_file', 'out_fa', 'out_adc', 'out_tdi', 'out_map']),
+        fields=['out_file', 'out_fa', 'out_adc', 'out_tdi', 'out_map',
+                'out_cst']),
         name='outputnode')
 
     bmsk = pe.Node(niu.Function(
@@ -262,7 +264,7 @@ def act_workflow(name='Tractography'):
     met = pe.Node(mrt3.TensorMetrics(
         out_adc='adc.nii.gz', out_fa='fa.nii.gz'), name='ComputeScalars')
     trk = pe.Node(mrt3.Tractography(
-        nthreads=0, n_tracks=int(1e5), max_length=250.), name='Track')
+        nthreads=0, n_tracks=int(1e7), max_length=200.), name='Track')
 
     lc = pe.Node(mrt3.LabelConfig(), name='LabelConfig')
     lc.inputs.out_file = 'parcellation.nii.gz'
@@ -273,6 +275,7 @@ def act_workflow(name='Tractography'):
 
     # tck2trk = pe.Node(TCK2TRK(), name='TCK2TRK')
     tck2vtk = pe.Node(mrt3.TCK2VTK(), name='TCK2VTK')
+
     tdi = pe.Node(mrt3.ComputeTDI(nthreads=0, out_file='tdi.nii.gz'),
[…]

### 2015-07-13 paper/main.tex
towards new version of paper

@@
 The final outcome of such applications, called tractography, is fundamental to gain information about
-  \gls*{wm} morphology in neurosurgical planning, post-surgery evaluations, and the study of neurological
-  diseases such as multiple sclerosis, and Alzheimer's disease \citep{chua_diffusion_2008}.
+  \gls*{wm} morphology in neurosurgical planning \citep{golby_interactive_2011},
+  post-surgery evaluations \citep{toda_utility_2014},
+  and the study of neurological diseases such as multiple sclerosis, and Alzheimer's disease
+  \citep{chua_diffusion_2008}.
 Tractography is also crucial in the analysis of structural brain networks using graph theory.
@@
   wide range of algorithms and processing methodologies available.
-In this work we propose \emph{diffantom}, an \emph{in-silico} dataset to assess tractography pipelines using
-  real datasets as source of whole-brain micro-structural information.
+In this report we present \emph{diffantom}, an \emph{in-silico} dataset to assess tractography pipelines using
+  real data as source of whole-brain micro-structural information.
 We also provide a full software layer to produce \emph{diffantoms} from real \gls*{dmri} datasets
@@
 \begin{itemize}
-  \item Comparison of the impact of the different diffusion sampling schemes on the local micro-structure
+  \item Comparisons of the impact of the different diffusion sampling schemes on the local micro-structure
   model of choice and the subsequent global tractography outcome.
-  \item Investigation of the effects of the different artifacts (noise, susceptibility-derived, Eddy-currents,
-  etc.) on the final connectivity matrix.
+  \item Investigations on the effects of the different artifacts (noise, \glspl*{pve} and \gls*{csf} contamination,
+    susceptibility-derived warping, Eddy-currents-derived distortions, etc.) at both local and global levels.
   \item Determine of the \acrfull*{roc} of connectivity pipelines.
   \item Generation of synthetic connectivity networks to test graph theory analyses.
+  \item Simulation of pathological brains (i.e. tumors as in \cite{kaus_simulation_2000}).
 \end{itemize}
@@
 The default value for \gls*{snr} is pre-set to 90.0.
+\Gls*{csf} was simulated with a unique compartment with isotropic diffusivity $D_{CSF}$ of 3.0$\times$10$^{-6}$ [mm s$^{-1}$].
+Conversely, \Gls*{wm} and \gls*{gm} included one isotropic compartment with $D_{WM} =$ and $D_{GM} =$, respectively.
+Single fiber contributions had $D_{\parallel}=$ and $D_{\bot}=$.
+All these values for diffusivity can be modified by the user for custom settings.
 

### 2015-07-14 paper/abbreviations.tex
finishing early version of paper
add details missing in paper
add new intro
towards new version of paper
A major update to the paper with new fig01 and coplete fig02

@@
 % @Last Modified by:   Oscar Esteban
-% @Last Modified time: 2015-06-17 16:28:24
+% @Last Modified time: 2015-07-14 12:49:57
 
 \newacronym{dmri}{dMRI}{diffusion MRI}
-\newacronym{hcp}{HCP}{Human Connectome Project}
+\newacronym{fa}{FA}{fractional anisotropy}
+\newacronym{adc}{ADC}{anisotropic diffusion coefficient}
+% \newacronym{hcp}{HCP}{Human Connectome Project}
+\newacronym{fod}{FOD}{fiber orientation distribution}
+\newacronym{wm}{WM}{white matter}
+\newacronym{gm}{GM}{gray matter}
+\newacronym{cgm}{cGM}{cortical gray matter}
+\newacronym{dgm}{dGM}{deep gray matter}
+\newacronym{csf}{CSF}{cerebrospinal fluid}
+\newacronym{snr}{SNR}{signal-to-noise ratio}
+\newacronym{csd}{CSD}{constrained spherical deconvolution}
+\newacronym{cc0}{CC0}{Creative Commons Zero licence}
+\newacronym{roc}{ROC}{receiver operating characteristic}
+\newacronym{pve}{PVE}{partial volume effect}
+\newacronym{cst}{CST}{corticospinal tract}
+
+\newglossaryentry{hcp}
+{
+	name={Human Connectome Project},
+	text={HCP},
+	first={Human Connectome Project (HCP, \cite{essen_human_2012})},
+	long={Human Connectome Project (HCP)},
+	description={Human Connectome Project}
+}
 

### 2015-07-14 paper/main.tex
A major update to the paper with new fig01 and coplete fig02

@@
 \input{abbreviations.tex}
+
+\newcommand{\e}[1]{\ensuremath{\;\cdot\,}10\ensuremath{^{#1}}}
 \begin{document}
@@
 
-\noindent\textbf{\textit{Data Generation\textcolon}}
+\noindent\textbf{\textit{Data Generation\textcolon}\label{sec:data_generation}}
 The methods to generate the data in this paper are inspired by the work of \cite{wilkins_fiber_2014} and
@@
   \emph{phantomas} \citep{caruyer_phantomas_2014}.
-By default, diffusion data are generated using a scheme of 100 directions distributed in two shells with uniform
+
+In order to generate \emph{a diffantom}, the simulation inputs (see \autoref{fig:figure1}, A) are computed.
+On one hand, the software requires up to five volume fractions $\{T_{1\,\cdots\,5}\}$ corresponding to the isotropic
+  compartments that will be simulated.
+The isotropic fraction maps use the \emph{5TT format} established with \emph{MRTrix3} \citep{tournier_mrtrix_2012}
+  that describes the structure of the brain in the following order: \gls*{cgm}, \gls*{dgm}, \gls*{wm}, \gls*{csf},
+  and abnormal tissue.
+Since we simulated a healthy subject, the last fraction map is omitted.
+On the other hand, the anisotropic compartments are specified by up to three volume fractions $\{F_{1\,\cdots\,3}\}$
+  of three single fiber populations per voxel along with their corresponding direction maps $\{V_{1\,\cdots\,3}\}$.
+The process to obtain these inputs is described as follows:
+1) The fiber orientation maps $\{V_{1\,\cdots\,3}\}$ and their corresponding single-fiber fraction maps are
+  obtained using the ball-and-stick model for multi-shell data of BEDPOSTX \citep{jbabdi_modelbased_2012}
+  on the \gls*{dmri} data of a certain \gls*{hcp} subject.
+2) We obtain a \gls*{fa} map after fitting a tensor model with \emph{MRTrix3}.
+3) The original single-fiber fractions and the \gls*{fa} map are smoothed with a non-local means filter included
+  in \emph{dipy} \citep{garyfallis_dipy_2011}.
+4) By combining with \emph{MRTrix} the volume fraction maps computed with FSL FAST \citep{zhang_segmentation_2001} and
+  FSL FIRST \citep{patenaude_bayesian_2011}, we obtain the macro-structural 5TT map of the subject.
+5) The objects obtained in the previous steps are combined as described in the \nameref{sec:appendix} to generate the
+  outcome presented in \autoref{fig:figure1}, box A.
+
+Once a micro-structural model of the volume has been synthesized, a map of \glspl{fod} is computed using spherical
+  convolution of the single fiber response and the fiber orientation maps $\{V_{1\,\cdots\,3}\}$, weighted by the
+  corresponding single-fiber fractions $\{F_{1\,\cdots\,3}\}$.
+The single fiber response is {\color{red} (Manu, please help me on define this)}, with $D_{\parallel}=$ 2.2\e{-3}
+  and $D_{\bot}=$ 0.2\e{-3} [mm s$^{-1}$].
+The resulting \glspl{fod} map is then combined with the isotropic compartments corresponding to $\{T_{1\,\cdots\,5}\}$
+  before generating the signal using \emph{phantomas} \citep{caruyer_phantomas_2014}.
+By default, diffusion data are generated using a scheme of 100 directions distributed in one shell with uniform
   coverage \citep{caruyer_design_2013}.
-Custom schemes can be generated supplying the tables of corresponding $b$-vectors and $b$-values.
+Custom one- or multi-shell schemes can be generated supplying the tables of corresponding $b$-vectors and $b$-values.
 Rician noise is also included in \emph{phantomas}, and can be set by the user.
 The default value for \gls*{snr} is pre-set to 90.0.
-\Gls*{csf} was simulated with a unique compartment with isotropic diffusivity $D_{CSF}$ of 3.0$\times$10$^{-6}$ [mm s$^{-1}$].
-Conversely, \Gls*{wm} and \gls*{gm} included one isotropic compartment with $D_{WM} =$ and $D_{GM} =$, respectively.
-Single fiber contributions had $D_{\parallel}=$ and $D_{\bot}=$.
-All these values for diffusivity can be modified by the user for custom settings.
+\Gls*{csf} was simulated with a unique compartment with isotropic diffusivity $D_{CSF}$ of 3.0\e{-3} [mm s$^{-1}$].
+Conversely, \Gls*{wm} and \gls*{gm} included one isotropic compartment with $D_{WM} =$ 2.0\e{-4}, $D_{cGM} =$ 7.0\e{-4}
+  and $D_{dGM} =$ 9.0\e{-4}, respectively [mm s$^{-1}$].
+All these values for diffusivity (and the corresponding to the single-fiber response) can be modified by the user with
+  custom settings.
 
@@
 \noindent\textbf{\textit{Conclusion\textcolon}}\label{sec:conclusion} %
-\emph{Diffantom} is a software layer to be applied on \acrfull*{hcp} datasets, that is openly and freely distributed
-  along with an example dataset.
+\emph{Diffantom} is a digital phantom generated from a dataset belonging to the \acrlong*{hcp}, and the software layer to
+  simulate it.
+A first \emph{diffantom} is presented here to be openly and freely distributed along with the necessary software to reproduce
+  this dataset and generate new \emph{diffantoms}.
 
 \section*{Data Sharing}
-The example dataset is available under the \gls*{cc0} in {\color{red} (dryad?)}.
-Source code of \emph{diffantom} is available at \url{https://github.com/oesteban/phantomas} under an MIT licence.
-\emph{Phantomas} is available in \url{https://github.com/ecaruyer/phantomas/} under the revised-BSD licence.
+The first \emph{diffantom} is available under the \gls*{cc0} in {\color{red} (dryad?)}.
+The associated software to ``\emph{diffantomize}'' real \gls*{dmri} data is available at \url{https://github.com/oesteban/diffantom}
+  under an MIT licence.
+\emph{Phantomas} is available in \url{https://github.com/ecaruyer/phantomas} under the revised-BSD licence.
 
@@
 
+\section*{Appendix}\label{sec:appendix}
+Let $\{T'_{1\,\cdots\,5}\}$ be the set of original fractions maps obtained with the combined use of FAST and FIRST, FA the
+  \gls*{fa} map obtained from the original \gls*{dmri} data, and $\{V_{1\,\cdots\,3}, F'_{1\,\cdots\,3}\}$ the direction
+  maps $V_i$ of single-fibers with their estimated volume fractions $F'_i$ calculated with BEDPOSTX.
+The final $\{T_{1\,\cdots\,5}\}$ maps of isotropic fractions are computed as follows:
+
+  \begin{align*}
+  T_1 &= (1.0-w_{cgm}) \cdot T'_1 \\
+  T_2 &= (1.0-w_{dgm}) \cdot T'_2 \\
+  T_3 &= (1.0-w_{wm}) \cdot T'_3 \\
+  T_4 &= T'_4 \\
+  T_5 &= 0.0
+  \end{align*}
+where $w_{\{cgm, dgm, wm\}}$ are weighing factors of anisotropic diffusion for each tissue
+  (defaults are $w_{cgm} =$ 25\%, $w_{dgm} =$ 50\%, $w_{cgm} =$ 95\%).
+The final $\{F_{1\,\cdots\,3}\}$ maps are computed as follows:
+
+\begin{align*}
+F_1 &= w_{wm} \cdot T_2 \cdot \text{FA} + w_{f1} (w_{cgm} \cdot T_1 + w_{dgm} \cdot T_2) \\
+F_2 &= w_{wm} \cdot T_2 - (F_1 + F_3) + w_{f2} (w_{cgm} \cdot T_1 + w_{dgm} \cdot T_2) \\
+F_3 &= w_{wm} \cdot F'_3 + w_{f3} (w_{cgm} \cdot T_1 + w_{dgm} \cdot T_2)
+\end{align*}
+where $w_{\{f1, f2, f3\}}$ are the contributions of the \gls*{gm} compartments to each single-fiber population.
+By default: $w_{f1} = $ 48\%, $w_{f2} = $ 37\%, $w_{f3} = $ 15\%.
+Finally, the resulting maps are normalized to fulfill $\sum_i T_i + \sum_i F_i = 1.0$.
+
+\newpage
 \section*{Figures}
@@
 \begin{center}
-\includegraphics[width=\linewidth]{figures/figure01}
+\includegraphics[width=\linewidth]{figures/figure01A}
+\includegraphics[width=\linewidth]{figures/figure01B}
 \end{center}
 \textbf{\refstepcounter{figure}\label{fig:figure1} Figure \arabic{figure}. }%
-{\emph{Diffantom} is a software layer wrapping the \acrfull{hcp} datasets to generate whole-brain diffusion \gls*{dmri} data.
-First, the tissue fraction maps $\{T_1, T_2, T_3\}$ are computed using FSL FAST \citep{zhang_segmentation_2001}.
-Second, the fiber orientation maps $\{V_1, V_2, V_3\}$ and corresponding volume fraction maps $\{F_1, F_2, F_3\}$ are
-  obtained using the ball-and-stick model for multi-shell data of BEDPOSTX \citep{jbabdi_modelbased_2012}.
-The third step takes the fiber orientations $\{V_1, V_2, V_3\}$ and their weights $\{F_1, F_2, F_3\}$ to compute the
-  corresponding \glspl{fod} using spherical convolution of the single fiber response with each fiber orientation at
-  each voxel.
-The resulting \glspl{fod} map is then combined with three isotropic compartments corresponding to $\{T_1, T_2, T'_3\}$
-  before generating the signal using \emph{phantomas} \citep{caruyer_phantomas_2014}.
-The fraction map $T'_3 = T_3 - \sum F_i$ is the remaining isotropic fraction across the \gls*{wm} tissue map.
-Finally, data can be used to compute the whole-brain tractography, for instance.}
+{\textbf{Generation of \emph{diffantoms}:} First, a micro-structural model is generated as described
+  in the \nameref{sec:data_generation} subsection.
+This micro-structural model is feeded into a simulation block that generates the final \gls*{dmri} dataset
+  as described in the second block.
+}
 \end{figure}
@@
   such as a tumor.
-\textbf{B1, B2, B3. Tractograms of fundamental fiber bundles}.
+\textbf{B1. Tractogram of fibers crossing slice 56 of the \emph{diffantom}} represented over that corresponding
+  slice of the \emph{b0} volume. \textbf{B2. Segmentation of some fiber bundles}: in blue color is represented
+  the right \gls*{cst}, in red the left \gls*{cst}, and in green the forceps minor.
+  The box includes the slice 56 of the \emph{b0} and the light gray shadow is the pial surface.
 }

### 2015-07-15 diffantom/interfaces.py
finished tracks segmentation. close #7, close #8

@@ -2,8 +2,8 @@
 # -*- coding: utf-8 -*-
 # @Author: oesteban
 # @Date:   2015-06-23 12:29:58
-# @Last Modified by:   Oscar Esteban
-# @Last Modified time: 2015-07-10 13:06:50
+# @Last Modified by:   oesteban
+# @Last Modified time: 2015-07-13 11:47:13
 
 import os
 import os.path as op
@@ -261,7 +261,8 @@ class TractQuerierInputSpec(CommandLineInputSpec):
 
 
 class TractQuerierOutputSpec(TraitedSpec):
-    out_file = File(exists=True, desc='the output response file')
+    out_file = OutputMultiPath(File(exists=True),
+                               desc='the output response file')
 
 
 class TractQuerier(CommandLine):

### 2015-07-15 diffantom/workflows.py

@@ -2,8 +2,8 @@
 # -*- coding: utf-8 -*-
 # @Author: oesteban
 # @Date:   2015-06-23 12:32:07
-# @Last Modified by:   Oscar Esteban
-# @Last Modified time: 2015-07-10 13:07:36
+# @Last Modified by:   oesteban
+# @Last Modified time: 2015-07-13 14:28:28
 
 import os
 import os.path as op
@@ -36,7 +36,8 @@ def gen_diffantom(name='Diffantom', settings={}):
                   vfractions='vfraction*.nii.gz',
                   scheme='samples.txt',
                   aparc='aparc+aseg.nii.gz',
-                  fa='fa.nii.gz')
+                  fa='fa.nii.gz',
+                  wmparc='wmparc.nii.gz')
 
     ds_tpl_args = {k: [['subject_id', [v]]] for k, v in fnames.iteritems()}
 
@@ -60,7 +61,8 @@ def gen_diffantom(name='Diffantom', settings={}):
                                ('vfractions', 'inputnode.fractions'),
                                ('aparc', 'inputnode.parcellation'),
                                ('fa', 'inputnode.in_fa')]),
-        (ds,        trk,      [('aparc', 'inputnode.aparc')]),
+        (ds,        trk,      [('aparc', 'inputnode.aparc'),
+                               ('wmparc', 'inputnode.wmparc')]),
         (ds,        sim_ref,  [('scheme', 'inputnode.scheme')]),
         (sim_mod,   sim_ref,     [
             ('outputnode.fibers', 'inputnode.fibers'),
@@ -240,7 +242,8 @@ def preprocess_model(name='PrepareModel'):
 
 def act_workflow(name='Tractography'):
     inputnode = pe.Node(niu.IdentityInterface(
-        fields=['in_dwi', 'in_scheme', 'in_5tt', 'parcellation', 'aparc']),
+        fields=['in_dwi', 'in_scheme', 'in_5tt', 'parcellation', 'aparc',
+                'wmparc']),
         name='inputnode')
     outputnode = pe.Node(niu.IdentityInterface(
         fields=['out_file', 'out_fa', 'out_adc', 'out_tdi', 'out_map',
@@ -264,7 +267,7 @@ def act_workflow(name='Tractography'):
     met = pe.Node(mrt3.TensorMetrics(
         out_adc='adc.nii.gz', out_fa='fa.nii.gz'), name='ComputeScalars')
     trk = pe.Node(mrt3.Tractography(
[…]

### 2015-07-15 paper/main.tex
add comment in fig1 caption
add references, fix typos, new nomenclature (hindered...)

@@
   \citep{sporns_human_2005}, the assessment of neurological diseases \citep{griffa_structural_2013}, and in the
-  study of the link of structural and functional connectivity \citep{messe_predicting_2015}.
+  study of the link between structural and functional connectivity \citep{messe_predicting_2015}.
 However, the development of the field is limited by the lack of gold standards to test and compare the
@@
 \section*{Background}
-The assessment of tractography from \gls*{dmri} data and subsequent analyses is challenging due to
+The assessment of tractography from \gls*{dmri} data is challenging due to
   the lack of a gold standard.
@@
 \noindent\textbf{\textit{Data Generation\textcolon}\label{sec:data_generation}}
-The methods to generate the data in this paper are inspired by the work of \cite{wilkins_fiber_2014} and
+The methods to generate \emph{diffantoms} are rooted in the work of \cite{wilkins_fiber_2014}, and
   described in \autoref{fig:figure1}.
@@
 
-In order to generate \emph{a diffantom}, the simulation inputs (see \autoref{fig:figure1}, A) are computed.
-On one hand, the software requires up to five volume fractions $\{T_{1\,\cdots\,5}\}$ corresponding to the isotropic
+The simulation process begins with generating a micro-structural model from real data (\autoref{fig:figure1}).
+On one hand, the software requires up to five volume fractions $\{T_{1\,\cdots\,5}\}$ corresponding to the free-diffusion
   compartments that will be simulated.
-The isotropic fraction maps use the \emph{5TT format} established with \emph{MRTrix3} \citep{tournier_mrtrix_2012}
+These isotropic fraction maps use the \emph{5TT format} established with \emph{MRTrix3} \citep{tournier_mrtrix_2012}
   that describes the structure of the brain in the following order: \gls*{cgm}, \gls*{dgm}, \gls*{wm}, \gls*{csf},
   and abnormal tissue.
-Since we simulated a healthy subject, the last fraction map is omitted.
-On the other hand, the anisotropic compartments are specified by up to three volume fractions $\{F_{1\,\cdots\,3}\}$
+Since we simulated a healthy subject, the last fraction map is empty and can be omitted.
+On the other hand, the hindered-diffusion compartments are specified by up to three volume fractions $\{F_{1\,\cdots\,3}\}$
   of three single fiber populations per voxel along with their corresponding direction maps $\{V_{1\,\cdots\,3}\}$.
 The process to obtain these inputs is described as follows:
-1) The fiber orientation maps $\{V_{1\,\cdots\,3}\}$ and their corresponding single-fiber fraction maps are
+1) The fiber orientation maps $\{V_{1\,\cdots\,3}\}$ and their corresponding fiber-fraction maps are
   obtained using the ball-and-stick model for multi-shell data of BEDPOSTX \citep{jbabdi_modelbased_2012}
@@
 2) We obtain a \gls*{fa} map after fitting a tensor model with \emph{MRTrix3}.
-3) The original single-fiber fractions and the \gls*{fa} map are smoothed with a non-local means filter included
+3) The original fiber-fractions and the \gls*{fa} map are smoothed with a non-local means filter included
   in \emph{dipy} \citep{garyfallis_dipy_2011}.
@@
   convolution of the single fiber response and the fiber orientation maps $\{V_{1\,\cdots\,3}\}$, weighted by the
-  corresponding single-fiber fractions $\{F_{1\,\cdots\,3}\}$.
-The single fiber response is {\color{red} (Manu, please help me on define this)}, with $D_{\parallel}=$ 2.2\e{-3}
-  and $D_{\bot}=$ 0.2\e{-3} [mm s$^{-1}$].
+  corresponding fiber-fractions $\{F_{1\,\cdots\,3}\}$.
+A close-up showing how the \glspl{fod} map looks is presented in \autoref{fig:figure1}.
+The single fiber response is a Gaussian diffusion tensor with axial symmetry and eigenvalues $\lambda_1=$ 2.2\e{-3}
+  and $\lambda_{2,3}=$ 0.2\e{-3} [mm$^2$s$^{-1}$].
 The resulting \glspl{fod} map is then combined with the isotropic compartments corresponding to $\{T_{1\,\cdots\,5}\}$
@@
 Rician noise is also included in \emph{phantomas}, and can be set by the user.
-The default value for \gls*{snr} is pre-set to 90.0.
-\Gls*{csf} was simulated with a unique compartment with isotropic diffusivity $D_{CSF}$ of 3.0\e{-3} [mm s$^{-1}$].
+The default value for \gls*{snr} is preset to 90.0.
+\Gls*{csf} was simulated with a unique compartment with isotropic diffusivity $D_{CSF}$ of 3.0\e{-3} [mm$^2$s$^{-1}$].
 Conversely, \Gls*{wm} and \gls*{gm} included one isotropic compartment with $D_{WM} =$ 2.0\e{-4}, $D_{cGM} =$ 7.0\e{-4}
-  and $D_{dGM} =$ 9.0\e{-4}, respectively [mm s$^{-1}$].
+  and $D_{dGM} =$ 9.0\e{-4}, respectively [mm$^2$s$^{-1}$].
 All these values for diffusivity (and the corresponding to the single-fiber response) can be modified by the user with
@@
 \noindent\textbf{\textit{Interpretation and use\textcolon}}
-In order to release a reproducible dataset, along with an usable software to create custom data,
-  we also distribute the software instrumentation presented in \autoref{fig:figure1}.
-To ensure reproducibility and usability, the simulation pipeline is written using
-  \emph{nipype} \citep{gorgolewski_nipype_2011}.
-\autoref{fig:figure2}, block A, describes the recommended use case of \emph{diffantom} in the evaluation
-  of algorithms and/or pipelines.
-Finally, \autoref{fig:figure2}, blocks B1-3, show some tractograms of fundamental bundles to illustrate
-  the features of the dataset.
-To obtain the whole-brain tractogram, the example dataset underwent a simplified connectivity pipeline including
-  \gls*{csd} and probabilistic tractography from
+The full simulation workflow, presented in \autoref{fig:figure1}, is implemented using
+  \emph{nipype} \citep{gorgolewski_nipype_2011} to ensure reproducibility and usability.
+The recommended use case of the simulation workflow is its integration in assessment frameworks
+  (\autoref{fig:figure2}, box A) for the evaluation of algorithms and/or pipelines.
+
+To illustrate the features of the \emph{diffantom}, the example dataset underwent a simplified
+  connectivity pipeline including \gls*{csd} and probabilistic tractography from
   \emph{MRTrix} \citep{tournier_mrtrix_2012}.
@@
   seed points evenly distributed across a dilated mask of the \gls*{wm} tissue.
+\autoref{fig:figure2}, box B1 shows the result of the tractography obtained with such pipeline.
+Finally, we applied \emph{tract querier} \citep{wassermann_on_2013} on segmenting important fiber bundles such
+  as the \gls*{cst} and the forceps minor (see \autoref{fig:figure2}, box B2).
+Particularly, due to its location nearby the orbitofrontal lobe, the forceps minor is generally affected by
+  susceptibility distortions.
+The study of this phenomena \citep{esteban_simulationbased_2014} led us to design \emph{diffantom}.
 
@@
 Then, making use of \emph{FiberFox}, the segmentation of each bundle is mapped to an analytical description
-  which is feeded into \emph{phantomas}.
+  which is fed into \emph{phantomas}.
 While this phantom is designed for the bundle-wise evaluation of tractography (with scores such as geometrical coverage,
@@
 \noindent\textbf{\textit{Conclusion\textcolon}}\label{sec:conclusion} %
-\emph{Diffantom} is a digital phantom generated from a dataset belonging to the \acrlong*{hcp}, and the software layer to
-  simulate it.
-A first \emph{diffantom} is presented here to be openly and freely distributed along with the necessary software to reproduce
-  this dataset and generate new \emph{diffantoms}.
+\emph{Diffantom} is a digital phantom generated from a dataset belonging to the \acrlong*{hcp}, and the software factory to
+  reproduce it.
+The first \emph{diffantom} is presented here to be openly and freely distributed along with the necessary resources to
+  generate new \emph{diffantoms}.
+We encourage the neuroimage community to contribute with their own \emph{diffantoms} and share them openly.
 
@@
 The associated software to ``\emph{diffantomize}'' real \gls*{dmri} data is available at \url{https://github.com/oesteban/diffantom}
-  under an MIT licence.
-\emph{Phantomas} is available in \url{https://github.com/ecaruyer/phantomas} under the revised-BSD licence.
+  under an MIT license.
+\emph{Phantomas} is available in \url{https://github.com/ecaruyer/phantomas} under the revised-BSD license.
 
@@
 EC implemented \emph{phantomas} \citep{caruyer_phantomas_2014}, helped integrate the project with the simulation routines.
-OE, EC, AD thoroughtly discussed and framed the aptness of the data in the community.
+OE, EC, AD thoroughly discussed and framed the aptness of the data in the community.
 AD, MBC, MJLC, and AS interpreted the resulting datasets.
@@
   \gls*{fa} map obtained from the original \gls*{dmri} data, and $\{V_{1\,\cdots\,3}, F'_{1\,\cdots\,3}\}$ the direction
-  maps $V_i$ of single-fibers with their estimated volume fractions $F'_i$ calculated with BEDPOSTX.
+  maps $V_i$ of restricted diffusion by fibers with their estimated volume fractions $F'_i$ calculated with BEDPOSTX.
 The final $\{T_{1\,\cdots\,5}\}$ maps of isotropic fractions are computed as follows:
@@
   \end{align*}
-where $w_{\{cgm, dgm, wm\}}$ are weighing factors of anisotropic diffusion for each tissue
+where $w_{\{cgm, dgm, wm\}}$ are weighing factors of hindered diffusion for each tissue
   (defaults are $w_{cgm} =$ 25\%, $w_{dgm} =$ 50\%, $w_{cgm} =$ 95\%).
@@
 \end{align*}
-where $w_{\{f1, f2, f3\}}$ are the contributions of the \gls*{gm} compartments to each single-fiber population.
+where $w_{\{f1, f2, f3\}}$ are the contributions of the \gls*{gm} compartments to each fiber population.
 By default: $w_{f1} = $ 48\%, $w_{f2} = $ 37\%, $w_{f3} = $ 15\%.
@@
   in the \nameref{sec:data_generation} subsection.
-This micro-structural model is feeded into a simulation block that generates the final \gls*{dmri} dataset
+Please note the piece-wise linear function of the color scale to enable visibility of small volume fractions.
+This micro-structural model is fed into a simulation block that generates the final \gls*{dmri} dataset
   as described in the second block.

### 2015-07-15 paper/main.tex
fix fig02 and its caption #10

@@
 \textbf{B1. Tractogram of fibers crossing slice 56 of the \emph{diffantom}} represented over that corresponding
-  slice of the \emph{b0} volume. \textbf{B2. Segmentation of some fiber bundles}: in blue color is represented
-  the right \gls*{cst}, in red the left \gls*{cst}, and in green the forceps minor.
-  The box includes the slice 56 of the \emph{b0} and the light gray shadow is the pial surface.
+  slice of the \emph{b0} volume. \textbf{B2. Segmentation of some fiber bundles}: the right \gls*{cst} is
+  represented in blue color, the left \gls*{cst} in red, and the forceps minor in green.
+  The box includes the slice 56 of the \emph{b0} and the pial surface is represented with a light gray shadow.
 }

### 2015-07-17 diffantom/workflows.py
close #11

@@ -128,7 +128,7 @@ def preprocess_model(name='PrepareModel'):
                         name='inputnode')
     outputnode = pe.Node(niu.IdentityInterface(
         fields=['fibers', 'fractions', 'out_5tt', 'out_iso', 'out_mask',
-                'parcellation']), name='outputnode')
+                'parcellation', 't1w_brain']), name='outputnode')
 
     def _getfirst(inlist):
         return inlist[0]
@@ -229,6 +229,7 @@ def preprocess_model(name='PrepareModel'):
         (faden,     fapst,      [('out_file', 'fa')]),
         (fapst,     outputnode, [('out_sf', 'fractions'),
                                  ('out_ts', 'out_iso')]),
+        (bet,       outputnode, [('out_file', 't1w_brain')]),
         (gen5tt,    outputnode, [('out_file', 'out_5tt')]),
         (fixtsr,    outputnode, [('out_file', 'fibers')]),
         (reslice,   outputnode, [('out_file', 'out_mask')]),

### 2015-07-22 diffantom/workflows.py
file changed location in mrtrix3

@@ -186,7 +186,7 @@ def preprocess_model(name='PrepareModel'):
     fixparc = pe.Node(mrt3.ReplaceFSwithFIRST(), name='FixFSaparc')
     fixparc.inputs.in_config = op.join(
         os.getenv('MRTRIX3_HOME', '/home/oesteban/workspace/mrtrix3'),
-        'src/dwi/tractography/connectomics/example_configs/fs_default.txt')
+        'src/connectome/config/fs_default.txt')
 
     faden = pe.Node(Denoise(snr=90.0), name='FADenoise')
     fapst = pe.Node(niu.Function(

### 2015-07-23 paper/main.tex
fixes for submitting to FINF
fix typo
changes proposed by AS

@@
 
+\usepackage[mode=buildnew]{standalone}
+\usepackage{tikz}
 %\setcitestyle{square}
 \usepackage{url,lineno,microtype}
+\usepackage[toc,nomain,acronym,shortcuts,translate=false]{glossaries}
 \usepackage[colorlinks=true,linkcolor=black, citecolor=black!80, urlcolor=black!80]{hyperref}
@@
 \usepackage[onehalfspacing]{setspace}
-\usepackage[toc,nomain,acronym,shortcuts,translate=false]{glossaries}
 
@@
 \def\Authors{Oscar Esteban\,$^{1,2,*}$, Emmanuel Caruyer\,$^{3}$, Alessandro Daducci\,$^{4}$, Meritxell Bach-Cuadra\,$^{4,5}$,%
-M.-Jes\'us Ledesma-Carbayo\,$^{1,2}$ and Andres Santos\,$^{1,2}$}
+Mar\'ia-J. Ledesma-Carbayo\,$^{1,2}$ and Andres Santos\,$^{1,2}$}
 
@@
 \newcommand{\e}[1]{\ensuremath{\;\cdot\,}10\ensuremath{^{#1}}}
+\newcommand*\glslines[2][]{\glslink[#1]{#2}{\glsdisp{#2}{\acrlong{#2} --\glsentryshort{#2}--}}}
+
 \begin{document}
@@
   post-surgery evaluations \citep{toda_utility_2014},
-  and the study of neurological diseases such as multiple sclerosis, and Alzheimer's disease
+  and the study of neurological diseases such as multiple sclerosis and Alzheimer's disease
   \citep{chua_diffusion_2008}.
@@
 In this report we present \emph{diffantom}, an \emph{in-silico} dataset to assess tractography pipelines using
-  real data as source of whole-brain micro-structural information.
+  real data as source of whole-brain microstructural information.
 We also provide a full software layer to produce \emph{diffantoms} from real \gls*{dmri} datasets
   drawn from the \gls*{hcp}.
-This dataset enables the following evaluations of standard connectivity pipelines:
+The \emph{diffantom} enables the following evaluations of standard connectivity pipelines:
 \begin{itemize}
-  \item Comparisons of the impact of the different diffusion sampling schemes on the local micro-structure
+  \item Comparisons of the impact of different diffusion sampling schemes on the local micro-structure
   model of choice and the subsequent global tractography outcome.
-  \item Investigations on the effects of the different artifacts (noise, \glspl*{pve} and \gls*{csf} contamination,
+  \item Investigations on the effects of the different artifacts (noise, \acrlong{pve} and \glslines{csf} contamination,
     susceptibility-derived warping, Eddy-currents-derived distortions, etc.) at both local and global levels.
-  \item Determine of the \acrfull*{roc} of connectivity pipelines.
+  \item Characterizing the \acrfull*{roc} of connectivity pipelines.
   \item Generation of synthetic connectivity networks to test graph theory analyses.
-  \item Simulation of pathological brains (i.e. tumors as in \cite{kaus_simulation_2000}).
+  \item Simulation of pathological brains (e.g. tumors as in \cite{kaus_simulation_2000}).
 \end{itemize}
@@
   so-called \emph{FiberCup} phantom \citep{poupon_new_2008,fillard_quantitative_2011}.
-These phantoms are appropriate to evaluate the resolution of fiber crossings and accuracy of
+These phantoms are appropriate to evaluate the angular resolution in fiber crossings and accuracy of
   direction-independent scalar parameters in very simplistic geometries.
@@
 These tools generally implemented the multi-tensor model \citep{alexander_analysis_2001,tuch_high_2002}
-  to simulate fibers crossing, fanning, kissing, etc.
+  to simulate fiber crossing, fanning, kissing, etc.
 \cite{close_software_2009} presented the \emph{Numerical Fibre Generator}, a software to simulate
   spherical shapes filled with digital fiber-tracts.
-\cite{caruyer_phantomas_2014} proposed \emph{Phantomas} to simulate any kind of analytic geometry
+\cite{caruyer_Phantomas_2014} proposed \emph{Phantomas} to simulate any kind of analytic geometry
   inside a sphere.
@@
   of fibers averaged from real \gls*{dmri} scans and the multi-tensor model with a compartment of
-  isotropic-diffusion.
+  isotropic diffusion.
 Recently, \cite{neher_fiberfox_2014} proposed \emph{FiberFox}, a visualization software to develop
@@
 Second improvement is a more advanced signal model to generate the phantom using
-  \emph{phantomas} \citep{caruyer_phantomas_2014}.
+  \emph{Phantomas} \citep{caruyer_Phantomas_2014}.
 
-The simulation process begins with generating a micro-structural model from real data (\autoref{fig:figure1}).
+The simulation process begins with generating a microstructural model from real data (\autoref{fig:figure1}).
 On one hand, the software requires up to five volume fractions $\{T_{1\,\cdots\,5}\}$ corresponding to the free-diffusion
   compartments that will be simulated.
-These isotropic fraction maps use the \emph{5TT format} established with \emph{MRTrix3} \citep{tournier_mrtrix_2012}
-  that describes the structure of the brain in the following order: \gls*{cgm}, \gls*{dgm}, \gls*{wm}, \gls*{csf},
-  and abnormal tissue.
+These isotropic fraction maps use the \emph{5TT format} established with the latest version 3.0 of
+  \emph{MRTrix} \citep{tournier_mrtrix_2012}, that describes the structure of the brain in the following
+  order: \gls*{cgm}, \gls*{dgm}, \gls*{wm}, \gls*{csf}, and abnormal tissue.
 Since we simulated a healthy subject, the last fraction map is empty and can be omitted.
@@
   of three single fiber populations per voxel along with their corresponding direction maps $\{V_{1\,\cdots\,3}\}$.
-The process to obtain these inputs is described as follows:
+
+The process to obtain the microstructural model is described as follows:
 1) The fiber orientation maps $\{V_{1\,\cdots\,3}\}$ and their corresponding fiber-fraction maps are
@@
   on the \gls*{dmri} data of a certain \gls*{hcp} subject.
-2) We obtain a \gls*{fa} map after fitting a tensor model with \emph{MRTrix3}.
+2) We obtain a \gls*{fa} map after fitting a tensor model with \emph{MRTrix}.
 3) The original fiber-fractions and the \gls*{fa} map are smoothed with a non-local means filter included
@@
 5) The objects obtained in the previous steps are combined as described in the \nameref{sec:appendix} to generate the
-  outcome presented in \autoref{fig:figure1}, box A.
+  microstructural model, presented in \autoref{fig:figure1} box A.
 
@@
 The resulting \glspl{fod} map is then combined with the isotropic compartments corresponding to $\{T_{1\,\cdots\,5}\}$
-  before generating the signal using \emph{phantomas} \citep{caruyer_phantomas_2014}.
+  before generating the signal using \emph{Phantomas} \citep{caruyer_Phantomas_2014}.
 By default, diffusion data are generated using a scheme of 100 directions distributed in one shell with uniform
@@
 Custom one- or multi-shell schemes can be generated supplying the tables of corresponding $b$-vectors and $b$-values.
-Rician noise is also included in \emph{phantomas}, and can be set by the user.
+Rician noise is also included in \emph{Phantomas}, and can be set by the user.
 The default value for \gls*{snr} is preset to 90.0.
@@
 Then, making use of \emph{FiberFox}, the segmentation of each bundle is mapped to an analytical description
-  which is fed into \emph{phantomas}.
+  which is fed into \emph{Phantomas}.
 While this phantom is designed for the bundle-wise evaluation of tractography (with scores such as geometrical coverage,
@@
   under an MIT license.
-\emph{Phantomas} is available in \url{https://github.com/ecaruyer/phantomas} under the revised-BSD license.
+\emph{Phantomas} is available in \url{https://github.com/ecaruyer/Phantomas} under the revised-BSD license.
 
@@
 OE designed the data generation procedure, implemented the processing pipelines and generated the example dataset.
-EC implemented \emph{phantomas} \citep{caruyer_phantomas_2014}, helped integrate the project with the simulation routines.
+EC implemented \emph{Phantomas} \citep{caruyer_Phantomas_2014}, helped integrate the project with the simulation routines.
 OE, EC, AD thoroughly discussed and framed the aptness of the data in the community.
@@
 
+\nolinenumbers
 \bibliographystyle{frontiers/frontiersinSCNS_ENG_HUMS}
@@
 \glsresetall
-
+\linenumbers
 \section*{Appendix}\label{sec:appendix}
@@
 \begin{center}
-\includegraphics[width=\linewidth]{figures/figure01A}
-\includegraphics[width=\linewidth]{figures/figure01B}
+% \includegraphics[width=\linewidth]{figures/figure01A}
+% \includegraphics[width=\linewidth]{figures/figure01B}
+\includestandalone[width=\linewidth]{figure01}
 \end{center}
 \textbf{\refstepcounter{figure}\label{fig:figure1} Figure \arabic{figure}. }%
-{\textbf{Generation of \emph{diffantoms}:} First, a micro-structural model is generated as described
+{\textbf{Generation of \emph{diffantoms}:} First, a microstructural model is generated as described
   in the \nameref{sec:data_generation} subsection.
 Please note the piece-wise linear function of the color scale to enable visibility of small volume fractions.
-This micro-structural model is fed into a simulation block that generates the final \gls*{dmri} dataset
+This microstructural model is fed into a simulation block that generates the final \gls*{dmri} dataset
   as described in the second block.

### 2015-07-23 paper/main.tex
changes proposed by MBC #10

@@
 \section*{Introduction}
-Fiber tracking on \gls*{dmri} data has become an important application in the in-vivo investigation
+Fiber tracking on \gls*{dmri} data has become an important application for the in-vivo investigation
   of the structural configuration of fiber bundles at the macro-scale.
 The final outcome of such applications, called tractography, is fundamental to gain information about
-  \gls*{wm} morphology in neurosurgical planning \citep{golby_interactive_2011},
+  \gls*{wm} morphology in many clinical routines like neurosurgical planning \citep{golby_interactive_2011},
   post-surgery evaluations \citep{toda_utility_2014},
@@
   study of the link between structural and functional connectivity \citep{messe_predicting_2015}.
-However, the development of the field is limited by the lack of gold standards to test and compare the
-  wide range of algorithms and processing methodologies available.
+However, the development of the field is limited by the lack of a gold standard to test and compare the
+  wide range of methodologies available for processing and analyzing \gls*{dmri}.
 In this report we present \emph{diffantom}, an \emph{in-silico} dataset to assess tractography pipelines using
@@
   drawn from the \gls*{hcp}.
-The \emph{diffantom} enables the following evaluations of standard connectivity pipelines:
+The \emph{diffantom} enables the following evaluations of connectivity pipelines:
 \begin{itemize}
-  \item Comparisons of the impact of different diffusion sampling schemes on the local micro-structure
-  model of choice and the subsequent global tractography outcome.
-  \item Investigations on the effects of the different artifacts (noise, \acrlong{pve} and \glslines{csf} contamination,
+  \item Studies on the impact of different diffusion sampling schemes on the local micro-structure
+  model of choice and on the subsequent global tractography outcome.
+  \item Assessment of sensitivity and robustness to imaging artifacts (noise, \acrlong{pve} and \glslines{csf} contamination,
     susceptibility-derived warping, Eddy-currents-derived distortions, etc.) at both local and global levels.
@@
   methodologies to build physical phantoms, digital simulations are increasingly popular.
-Early digital phantoms started with simulation of simple geometries as well
+Early digital phantoms also started with simulation of simple geometries
   \citep{basser_in_2000,goessl_fiber_2002,tournier_limitations_2002,leemans_mathematical_2005}
@@
   order: \gls*{cgm}, \gls*{dgm}, \gls*{wm}, \gls*{csf}, and abnormal tissue.
-Since we simulated a healthy subject, the last fraction map is empty and can be omitted.
+Since we simulate a healthy subject, the last fraction map is empty and can be omitted.
 On the other hand, the hindered-diffusion compartments are specified by up to three volume fractions $\{F_{1\,\cdots\,3}\}$
@@
 The \emph{diffantom} is a derivative of \citep{wilkins_fiber_2014} in terms of methodology for
-  simulation with two major improvements: the correctness of the \emph{minimally preprocessed} data
+  simulation with two major advances: the correctness of the \emph{minimally preprocessed} data
   \citep{glasser_minimal_2013} released within the \gls*{hcp}, and a more realistic diffusion

### 2015-07-23 paper/main.tex
changes proposed by GW #10

@@
   post-surgery evaluations \citep{toda_utility_2014},
-  and the study of neurological diseases such as multiple sclerosis and Alzheimer's disease
-  \citep{chua_diffusion_2008}.
+  and the study of neurological diseases as in \citep{chua_diffusion_2008} addressing multiple sclerosis and
+  Alzheimer's disease.
 Tractography is also crucial in the analysis of structural brain networks using graph theory.
@@
   direction-independent scalar parameters in very simplistic geometries.
-Since the complexity of whole-brain tractography is unaccountable with current materials and proposed
-  methodologies to build physical phantoms, digital simulations are increasingly popular.
+Digital simulations are increasingly popular because the complexity of whole-brain tractography
+  can not be accounted for with current materials and proposed methodologies to build physical phantoms.
 Early digital phantoms also started with simulation of simple geometries
@@
 
-The simulation process begins with generating a microstructural model from real data (\autoref{fig:figure1}).
+The simulation process is started by generating a microstructural model from real data (\autoref{fig:figure1}).
 On one hand, the software requires up to five volume fractions $\{T_{1\,\cdots\,5}\}$ corresponding to the free-diffusion
@@
 
-The process to obtain the microstructural model is described as follows:
+The process to obtain the microstructural model can be described like follows:
 1) The fiber orientation maps $\{V_{1\,\cdots\,3}\}$ and their corresponding fiber-fraction maps are
@@
   on the \gls*{dmri} data of a certain \gls*{hcp} subject.
-2) We obtain a \gls*{fa} map after fitting a tensor model with \emph{MRTrix}.
+2) A \gls*{fa} map is obtained after fitting a tensor model with \emph{MRTrix}.
 3) The original fiber-fractions and the \gls*{fa} map are smoothed with a non-local means filter included
   in \emph{dipy} \citep{garyfallis_dipy_2011}.
-4) By combining with \emph{MRTrix} the volume fraction maps computed with FSL FAST \citep{zhang_segmentation_2001} and
-  FSL FIRST \citep{patenaude_bayesian_2011}, we obtain the macro-structural 5TT map of the subject.
+4) The macrostructural 5TT map is extracted from the \acrlong*{t1} image of the subject, using \emph{MRTrix} tools.
 5) The objects obtained in the previous steps are combined as described in the \nameref{sec:appendix} to generate the
-  microstructural model, presented in \autoref{fig:figure1} box A.
+  microstructural model, presented in \autoref{fig:figure1}-A.
 
-Once a micro-structural model of the volume has been synthesized, a map of \glspl{fod} is computed using spherical
+Once a microstructural model of the volume has been synthesized, a map of \glspl{fod} is computed using spherical
   convolution of the single fiber response and the fiber orientation maps $\{V_{1\,\cdots\,3}\}$, weighted by the
@@
 This facilitates the use of \emph{diffantom} as a factory of whole-brain diffusion phantoms.
-Lastly, since the gradient scheme can be set by the user, \emph{diffantom} can be seen as a mean to translate the so-called
+Finally, since the gradient scheme can be set by the user, \emph{diffantom} can be seen as a mean to translate the so-called
   \emph{b-matrix} of the \gls*{hcp} to any target scheme.

### 2015-07-23 paper/main.tex
fix missing references to FAST and FIRST

@@
 \section*{Appendix}\label{sec:appendix}
-Let $\{T'_{1\,\cdots\,5}\}$ be the set of original fractions maps obtained with the combined use of FAST and FIRST, FA the
-  \gls*{fa} map obtained from the original \gls*{dmri} data, and $\{V_{1\,\cdots\,3}, F'_{1\,\cdots\,3}\}$ the direction
+Let $\{T'_{1\,\cdots\,5}\}$ be the set of original fractions maps obtained with \texttt{act\_anat\_prepare\_fsl}, a
+  tool in \emph{MRTrix} that combines FAST \citep{zhang_segmentation_2001} and FIRST \citep{patenaude_bayesian_2011}
+  to generate the macrostructural 5TT map.
+FA denotes the \gls*{fa} map obtained from the original \gls*{dmri} data, and $\{V_{1\,\cdots\,3}, F'_{1\,\cdots\,3}\}$ the direction
   maps $V_i$ of restricted diffusion by fibers with their estimated volume fractions $F'_i$ calculated with BEDPOSTX.

### 2015-07-24 paper/main.tex
add ack to GW #10
some fixes, added a new approach to the fractions problem #10

@@
   spherical shapes filled with digital fiber-tracts.
-\cite{caruyer_Phantomas_2014} proposed \emph{Phantomas} to simulate any kind of analytic geometry
+\cite{caruyer_phantomas_2014} proposed \emph{Phantomas} to simulate any kind of analytic geometry
   inside a sphere.
@@
 Second improvement is a more advanced signal model to generate the phantom using
-  \emph{Phantomas} \citep{caruyer_Phantomas_2014}.
+  \emph{Phantomas} \citep{caruyer_phantomas_2014}.
 
@@
 The resulting \glspl{fod} map is then combined with the isotropic compartments corresponding to $\{T_{1\,\cdots\,5}\}$
-  before generating the signal using \emph{Phantomas} \citep{caruyer_Phantomas_2014}.
+  before generating the signal using \emph{Phantomas} \citep{caruyer_phantomas_2014}.
 By default, diffusion data are generated using a scheme of 100 directions distributed in one shell with uniform
@@
 OE designed the data generation procedure, implemented the processing pipelines and generated the example dataset.
-EC implemented \emph{Phantomas} \citep{caruyer_Phantomas_2014}, helped integrate the project with the simulation routines.
+EC implemented \emph{Phantomas} \citep{caruyer_phantomas_2014}, helped integrate the project with the simulation routines.
 OE, EC, AD thoroughly discussed and framed the aptness of the data in the community.
@@
 \section*{Acknowledgments}
+We thank Gert Wollny for his revision of this work.
 \textit{Funding\textcolon}
@@
   \begin{align*}
-  T_1 &= (1.0-w_{cgm}) \cdot T'_1 \\
-  T_2 &= (1.0-w_{dgm}) \cdot T'_2 \\
-  T_3 &= (1.0-w_{wm}) \cdot T'_3 \\
+  T_1 &= (1.0-f_{cgm}) \cdot T'_1 \\
+  T_2 &= (1.0-f_{dgm}) \cdot T'_2 \\
+  T_3 &= (1.0-f_{wm}) \cdot T'_3 \\
   T_4 &= T'_4 \\
@@
   \end{align*}
-where $w_{\{cgm, dgm, wm\}}$ are weighing factors of hindered diffusion for each tissue
-  (defaults are $w_{cgm} =$ 25\%, $w_{dgm} =$ 50\%, $w_{cgm} =$ 95\%).
+where $f_{\{cgm, dgm, wm\}}$ are the fractions of hindered diffusion for each tissue.
+\cite{sepehrband_brain_2015} found out that the fiber fraction ranges across the corpus
+  callosum from the 70$\pm$8\% in its body to an upper bound of 80$\pm$11\% in the splenium.
+Therefore, we choose $f_{wm} =$ 80\% as default fraction of hindered diffusion in the
+  \gls*{wm}.
+To our knowledge, hindered diffusion fractions have been studied only for \gls*{wm}.
+Therefore, we set $f_{cgm} =$ 25\% and $f_{dgm} =$ 50\% as they yield plausible \gls*{fa}
+  and \gls*{adc} maps, assessed visually.
 The final $\{F_{1\,\cdots\,3}\}$ maps are computed as follows:
@@
 \begin{align*}
-F_1 &= w_{wm} \cdot T_2 \cdot \text{FA} + w_{f1} (w_{cgm} \cdot T_1 + w_{dgm} \cdot T_2) \\
-F_2 &= w_{wm} \cdot T_2 - (F_1 + F_3) + w_{f2} (w_{cgm} \cdot T_1 + w_{dgm} \cdot T_2) \\
-F_3 &= w_{wm} \cdot F'_3 + w_{f3} (w_{cgm} \cdot T_1 + w_{dgm} \cdot T_2)
+F_1 &= f_{wm} \cdot T_2 \cdot \text{FA} + w_{f1} (f_{cgm} \cdot T_1 + f_{dgm} \cdot T_2) \\
+F_2 &= f_{wm} \cdot T_2 - (F_1 + F_3) + w_{f2} (f_{cgm} \cdot T_1 + f_{dgm} \cdot T_2) \\
+F_3 &= f_{wm} \cdot F'_3 + w_{f3} (f_{cgm} \cdot T_1 + f_{dgm} \cdot T_2)
 \end{align*}

### 2015-08-12 paper/main.tex
I hope this fixes #30

@@
   wide range of methodologies available for processing and analyzing \gls*{dmri}.
-In this report we present \emph{diffantom}, an \emph{in-silico} dataset to assess tractography pipelines using
-  real data as source of whole-brain microstructural information.
-We also provide a full software layer to produce \emph{diffantoms} from real \gls*{dmri} datasets
-  drawn from the \gls*{hcp}.
-The \emph{diffantom} enables the following evaluations of connectivity pipelines:
-\begin{itemize}
-  \item Studies on the impact of different diffusion sampling schemes on the local micro-structure
-  model of choice and on the subsequent global tractography outcome.
-  \item Assessment of sensitivity and robustness to imaging artifacts (noise, \acrlong{pve} and \glslines{csf} contamination,
-    susceptibility-derived warping, Eddy-currents-derived distortions, etc.) at both local and global levels.
-  \item Characterizing the \acrfull*{roc} of connectivity pipelines.
-  \item Generation of synthetic connectivity networks to test graph theory analyses.
-  \item Simulation of pathological brains (e.g. tumors as in \cite{kaus_simulation_2000}).
-\end{itemize}
 
-\section*{Background}
-The assessment of tractography from \gls*{dmri} data is challenging due to
-  the lack of a gold standard.
 Large efforts have been devoted to the development of physical phantoms
@@
 
-\section*{Data description}
-
-\noindent\textbf{\textit{Data Generation\textcolon}\label{sec:data_generation}}
-The methods to generate \emph{diffantoms} are rooted in the work of \cite{wilkins_fiber_2014}, and
-  described in \autoref{fig:figure1}.
-\emph{Diffantom} integrates two major improvements with respect to \citep{wilkins_fiber_2014}.
+\emph{Diffantom} is originally designed for the investigation of susceptibility-derived distortions, a
+  typical artifact that produces geometrical warping in certain regions of \gls*{dmri} datasets.
+In \citep{esteban_simulationbased_2014} we addressed this phenomenon and concluded that the connectivity
+  matrix of \emph{Phantomas} was not dense enough to evaluate the integration of correction methods
+  in pipelines for the connectome extraction.
+In this report we present \emph{diffantom}, an \emph{in-silico} dataset to assess pipelines for tractography
+  and connectome extraction using real data as source of whole-brain microstructural information.
+\emph{Diffantom} is inspired by the work of \cite{wilkins_fiber_2014}, with two principal novelties.
 First, since we use a dataset from the \gls*{hcp} as input, data are corrected for the most relevant distortions.
@@
   corrections, and thus, generated data are affected correspondingly.
-Second improvement is a more advanced signal model to generate the phantom using
-  \emph{Phantomas} \citep{caruyer_phantomas_2014}.
+The second improvement is a more advanced signal model to generate the phantom using the hindered and restricted
+  diffusion model of \emph{Phantomas} \citep{caruyer_phantomas_2014}.
 
-The simulation process is started by generating a microstructural model from real data (\autoref{fig:figure1}).
-On one hand, the software requires up to five volume fractions $\{T_{1\,\cdots\,5}\}$ corresponding to the free-diffusion
-  compartments that will be simulated.
-These isotropic fraction maps use the \emph{5TT format} established with the latest version 3.0 of
-  \emph{MRTrix} \citep{tournier_mrtrix_2012}, that describes the structure of the brain in the following
-  order: \gls*{cgm}, \gls*{dgm}, \gls*{wm}, \gls*{csf}, and abnormal tissue.
-Since we simulate a healthy subject, the last fraction map is empty and can be omitted.
+\section*{Data description}
+
+\noindent\textbf{\textit{Microstructural model\textcolon}\label{sec:data_model}} %
+The simulation process relies on a microstructural model derived from real data (\autoref{fig:figure1}).
+On one hand, the software requires up to five volume fractions $\{T_{1\,\cdots\,5}\}$ corresponding to the
+  free-diffusion compartments.
+These fraction maps are obtained from the macroscopic structure of tissues in the brain,
+  in the following order~%
+\footnote{Corresponding to the \emph{5TT format} established with the latest version 3.0
+  of \emph{MRTrix} \citep{tournier_mrtrix_2012}}:
+\gls*{cgm}, \gls*{dgm}, \gls*{wm}, \gls*{csf}, and abnormal tissue~%
+\footnote{Since here we simulate healthy subjects, the last fraction map is empty and can be omitted}.
 On the other hand, the hindered-diffusion compartments are specified by up to three volume fractions $\{F_{1\,\cdots\,3}\}$
@@
 
+\noindent\textbf{\textit{Diffusion signal generation\textcolon}\label{sec:data_dwi}} %
 Once a microstructural model of the volume has been synthesized, a map of \glspl{fod} is computed using spherical
@@
 
-\noindent\textbf{\textit{Interpretation and use\textcolon}}
-The full simulation workflow, presented in \autoref{fig:figure1}, is implemented using
+\noindent\textbf{\textit{Implementation and reproducibility\textcolon}\label{sec:data_workflow}} %
+We also provide a full software layer to produce \emph{diffantoms} from real \gls*{dmri} datasets
+  drawn from the \gls*{hcp}.
+This workflow, presented in \autoref{fig:figure1}, is implemented using
   \emph{nipype} \citep{gorgolewski_nipype_2011} to ensure reproducibility and usability.
-The recommended use case of the simulation workflow is its integration in assessment frameworks
-  (\autoref{fig:figure2}, box A) for the evaluation of algorithms and/or pipelines.
 
+\noindent\textbf{\textit{Interpretation and recommended uses\textcolon}\label{sec:data_use}} %
 To illustrate the features of the \emph{diffantom}, the example dataset underwent a simplified
@@
   susceptibility distortions.
-The study of this phenomena \citep{esteban_simulationbased_2014} led us to design \emph{diffantom}.
+
+Since the gradient scheme can be set by the user, \emph{diffantom} can be seen as a mean to translate the so-called
+  \emph{b-matrix} of the \gls*{hcp} to any target scheme.
+A second recommended use case of the simulation workflow is its integration in assessment frameworks
+  (\autoref{fig:figure2}, box A) for the unit testing of algorithms, the integration tests of
+  modules in workflows and the full validation of pipelines.
+The \emph{diffantom} enables the following evaluations of connectivity pipelines:
+\begin{itemize}
+  \item Studies on the impact of different diffusion sampling schemes on the local micro-structure
+  model of choice and on the subsequent global tractography outcome.
+  \item Assessment of sensitivity and robustness to imaging artifacts (noise, \acrlong{pve} and \glslines{csf} contamination,
+    susceptibility-derived warping, Eddy-currents-derived distortions, etc.) at both local and global levels.
+  \item Characterizing the \acrfull*{roc} of connectivity pipelines.
+  \item Generation of synthetic connectivity networks to test graph theory analyses.
+  \item Simulation of pathological brains (e.g. tumors as in \cite{kaus_simulation_2000}).
+\end{itemize}
 
 \section*{Discussion and conclusion}
-
 \noindent\textbf{\textit{Discussion\textcolon}}\label{sec:discussion} %
@@
 This facilitates the use of \emph{diffantom} as a factory of whole-brain diffusion phantoms.
-Finally, since the gradient scheme can be set by the user, \emph{diffantom} can be seen as a mean to translate the so-called
-  \emph{b-matrix} of the \gls*{hcp} to any target scheme.
 

### 2015-08-12 paper/main.tex
Move ismrm 2015 to intro. Fix #36. Fix #27

@@
   isotropic diffusion.
-Recently, \cite{neher_fiberfox_2014} proposed \emph{FiberFox}, a visualization software to develop
+\cite{neher_fiberfox_2014} proposed \emph{FiberFox}, a visualization software to develop
   complex geometries and their analytical description.
@@
   methodology very close to that implemented in \emph{Phantomas}.
+An interesting outcome of \emph{FiberFox} is the phantom dataset\footnote{available at
+  \url{http://www.tractometer.org/ismrm_2015_challenge/}} created for the Tractography
+  Challenge held in ISMRM 2015.
+This dataset was derived from the tractography extracted in one \gls*{hcp} dataset.
+In the tractogram, 25 fiber bundles of interest were manually segmented by experts.
+Using \emph{FiberFox}, the segmentation of each bundle was mapped to an analytical
+  description and finally simulated the signal.
 
@@
   model.
-A possible competitor to \emph{diffantom} is the ground-truth data generated by the organizers of the
-  Tractography Challenge held in the ISMRM 2015\footnote{\url{http://www.tractometer.org/ismrm_2015_challenge/}}.
-Similarly to \emph{diffantom}, one \gls*{hcp} dataset is used as source of structural information.
-However, in this case 25 fiber bundles were manually segmented to obtain a summary tractogram.
-Then, making use of \emph{FiberFox}, the segmentation of each bundle is mapped to an analytical description
-  which is fed into \emph{Phantomas}.
+A possible competitor to \emph{diffantom} is the phantom generated for the Tractography Challenge in
+  ISMRM 2015.
+Similarly to \emph{diffantom}, they organizers used a \gls*{hcp} subject as source of structural information.
 While this phantom is designed for the bundle-wise evaluation of tractography (with scores such as geometrical coverage,

### 2015-08-12 paper/main.tex
follow suggestion of EC. fix #21

@@
 \noindent\textbf{\textit{Implementation and reproducibility\textcolon}\label{sec:data_workflow}} %
-We also provide a full software layer to produce \emph{diffantoms} from real \gls*{dmri} datasets
-  drawn from the \gls*{hcp}.
+We also provide the full software stack used to generate \emph{diffantoms}, so that users can
+  regenerate similar datasets with different parameters.
 This workflow, presented in \autoref{fig:figure1}, is implemented using

### 2015-08-12 paper/main.tex
Remove software layer concept. Close #13
hindered vs restricted diffusion. Close #18

@@
 
-\title[Diffantom]{Diffantom: a software layer to build digital phantoms of whole-brain diffusion MRI}
+\title[Diffantom]{Diffantom: a software stack to build digital phantoms of whole-brain diffusion MRI}
 
@@
 \noindent\textbf{\textit{Microstructural model\textcolon}\label{sec:data_model}} %
-The simulation process relies on a microstructural model derived from real data (\autoref{fig:figure1}).
-On one hand, the software requires up to five volume fractions $\{T_{1\,\cdots\,5}\}$ corresponding to the
-  free-diffusion compartments.
-These fraction maps are obtained from the macroscopic structure of tissues in the brain,
-  in the following order~%
+The simulation process relies on a microstructural model derived from real data.
+On one hand, the software requires up to five fraction maps $\{T_{1\,\cdots\,5}\}$ of
+  free- and hindered- diffusion (see \autoref{fig:figure1}, box A).
+These compartements will be derived from the macroscopic structure of tissues within the brain,
+  specified in the following order~%
 \footnote{Corresponding to the \emph{5TT format} established with the latest version 3.0
@@
 \footnote{Since here we simulate healthy subjects, the last fraction map is empty and can be omitted}.
-On the other hand, the hindered-diffusion compartments are specified by up to three volume fractions $\{F_{1\,\cdots\,3}\}$
+On the other hand, the restricted-diffusion compartments are specified by up to three volume fractions $\{F_{1\,\cdots\,3}\}$
   of three single fiber populations per voxel along with their corresponding direction maps $\{V_{1\,\cdots\,3}\}$.
@@
   \end{align*}
-where $f_{\{cgm, dgm, wm\}}$ are the fractions of hindered diffusion for each tissue.
+where $f_{\{cgm, dgm, wm\}}$ are the fractions of restricted diffusion for each tissue.
 \cite{sepehrband_brain_2015} found out that the fiber fraction ranges across the corpus
   callosum from the 70$\pm$8\% in its body to an upper bound of 80$\pm$11\% in the splenium.
-Therefore, we choose $f_{wm} =$ 80\% as default fraction of hindered diffusion in the
+Therefore, we choose $f_{wm} =$ 80\% as default fraction of restricted diffusion in the
   \gls*{wm}.
-To our knowledge, hindered diffusion fractions have been studied only for \gls*{wm}.
+To our knowledge, restricted diffusion fractions have been studied only for \gls*{wm}.
 Therefore, we set $f_{cgm} =$ 25\% and $f_{dgm} =$ 50\% as they yield plausible \gls*{fa}
@@
 \begin{center}
-% \includegraphics[width=\linewidth]{figures/figure01A}
-% \includegraphics[width=\linewidth]{figures/figure01B}
 \includestandalone[width=\linewidth]{figure01}

### 2015-08-12 paper/main.tex
change title. fix #14. With this, we should have clarified also #12

@@
 
-\title[Diffantom]{Diffantom: a software stack to build digital phantoms of whole-brain diffusion MRI}
+\title[Diffantom]{Diffantom: a whole-brain diffusion MRI phantom derived from a real dataset}
 
@@
   methodology very close to that implemented in \emph{Phantomas}.
-An interesting outcome of \emph{FiberFox} is the phantom dataset\footnote{available at
+An interesting outcome of \emph{FiberFox} is the phantom dataset\footnote{Available at
   \url{http://www.tractometer.org/ismrm_2015_challenge/}} created for the Tractography

### 2015-08-12 paper/main.tex
fixes #15

@@
   of the structural configuration of fiber bundles at the macro-scale.
-The final outcome of such applications, called tractography, is fundamental to gain information about
-  \gls*{wm} morphology in many clinical routines like neurosurgical planning \citep{golby_interactive_2011},
-  post-surgery evaluations \citep{toda_utility_2014},
+Tractography is fundamental to gain information about \gls*{wm} morphology in many clinical routines
+  like neurosurgical planning \citep{golby_interactive_2011}, post-surgery evaluations \citep{toda_utility_2014},
   and the study of neurological diseases as in \citep{chua_diffusion_2008} addressing multiple sclerosis and
   Alzheimer's disease.
-Tractography is also crucial in the analysis of structural brain networks using graph theory.
+Tractography is also applied in research for the analysis of structural brain networks using graph theory.
 These analyses have been successfully applied in the definition of the unique subject-wise patterns of connectivity
@@
   free- and hindered- diffusion (see \autoref{fig:figure1}, box A).
-These compartements will be derived from the macroscopic structure of tissues within the brain,
+These compartments will be derived from the macroscopic structure of tissues within the brain,
   specified in the following order~%

### 2015-08-12 paper/main.tex
fix #12

@@
 \section*{Introduction}
-Fiber tracking on \gls*{dmri} data has become an important application for the in-vivo investigation
-  of the structural configuration of fiber bundles at the macro-scale.
+Fiber tracking on \gls*{dmri} data has become an important application for the \textit{in-vivo} investigation
+  of the structural configuration of fiber bundles at the macroscale.
 Tractography is fundamental to gain information about \gls*{wm} morphology in many clinical routines
@@
   Alzheimer's disease.
-Tractography is also applied in research for the analysis of structural brain networks using graph theory.
-These analyses have been successfully applied in the definition of the unique subject-wise patterns of connectivity
+The analysis of structural brain networks using graph theory can also be applied on tractography.
+For instance, in the definition of the unique subject-wise patterns of connectivity
   \citep{sporns_human_2005}, the assessment of neurological diseases \citep{griffa_structural_2013}, and in the
@@
   expertise and operation time.
-This facilitates the use of \emph{diffantom} as a factory of whole-brain diffusion phantoms.
+This facilitates the reproducibility of the dataset by using the associated software workflow to easily
+  generate custom \emph{diffantoms}.
 
 \noindent\textbf{\textit{Conclusion\textcolon}}\label{sec:conclusion} %
-\emph{Diffantom} is a digital phantom generated from a dataset belonging to the \acrlong*{hcp}, and the software factory to
-  reproduce it.
-The first \emph{diffantom} is presented here to be openly and freely distributed along with the necessary resources to
-  generate new \emph{diffantoms}.
+\emph{Diffantom} is a digital phantom generated from a dataset belonging to the \acrlong*{hcp}.
+The first \emph{diffantom} is presented here to be openly and freely distributed along with the necessary
+  software resources to generate new \emph{diffantoms}.
 We encourage the neuroimage community to contribute with their own \emph{diffantoms} and share them openly.

### 2015-08-12 paper/main.tex
Starting with #22. Close #20

@@
 
+We recommend \emph{Diffantom} as ground-truth in assessment frameworks (\autoref{fig:figure2}, box A)
+  for the unit testing of algorithms, the integration tests of modules in workflows and the full
+  validation of pipelines:
+\begin{itemize}
+  \item Investigating the impact of different diffusion sampling schemes on the local microstructure
+    model of choice and on the subsequent global tractography outcome.
   Since the gradient scheme can be set by the user, \emph{diffantom} can be seen as a mean to translate the so-called
-  \emph{b-matrix} of the \gls*{hcp} to any target scheme.
-A second recommended use case of the simulation workflow is its integration in assessment frameworks
-  (\autoref{fig:figure2}, box A) for the unit testing of algorithms, the integration tests of
-  modules in workflows and the full validation of pipelines.
-The \emph{diffantom} enables the following evaluations of connectivity pipelines:
-\begin{itemize}
-  \item Studies on the impact of different diffusion sampling schemes on the local micro-structure
-  model of choice and on the subsequent global tractography outcome.
+  \emph{b-matrix} of the source dataset to any target scheme.
   \item Assessment of sensitivity and robustness to imaging artifacts (noise, \acrlong{pve} and \glslines{csf} contamination,

### 2015-08-12 paper/main.tex
New text for applications. close #22

@@
   \item Assessment of sensitivity and robustness to imaging artifacts (noise, \acrlong{pve} and \glslines{csf} contamination,
-    susceptibility-derived warping, Eddy-currents-derived distortions, etc.) at both local and global levels.
-  \item Characterizing the \acrfull*{roc} of connectivity pipelines.
-  \item Generation of synthetic connectivity networks to test graph theory analyses.
-  \item Simulation of pathological brains (e.g. tumors as in \cite{kaus_simulation_2000}).
+    susceptibility-derived warping, Eddy-currents-derived distortions, etc.) at unit, integration and overall validation levels
+    of evaluation.
+  \item Using the \emph{Diffantom} as in box A of \autoref{fig:figure2}, if is possible to apply binary classification measures
+    to evaluate the resulting connectivity matrix.
+  Considering the connectivity matrix of the reference \emph{Diffantom} as a message to be sent through a pipeline, the resulting
+    matrix of the test \emph{Diffantom} would be the received message, and the \acrfull*{roc} of the pipeline can be
+    characterized.
+  \item Simulation of pathological brains by altering the microstructural
+    model accordingly (e.g. as tumors were simulated in \cite{kaus_simulation_2000}).
 \end{itemize}
+In order to exemplify one of these intended uses, we also release a \emph{Diffantom} including the susceptibility-derived
+  distortion in simulation.
 

### 2015-08-12 diffantom/workflows.py
change default SNR. close #19

@@ -2,8 +2,8 @@
 # -*- coding: utf-8 -*-
 # @Author: oesteban
 # @Date:   2015-06-23 12:32:07
-# @Last Modified by:   oesteban
-# @Last Modified time: 2015-07-13 14:28:28
+# @Last Modified by:   Oscar Esteban
+# @Last Modified time: 2015-08-12 15:37:54
 
 import os
 import os.path as op
@@ -90,7 +90,7 @@ def simulate(name='SimDWI'):
 
     sch = pe.Node(LoadSamplingScheme(bvals=[2000]), name='LoadScheme')
     simdwi = pe.Node(PhantomasSticksSim(
-        lambda1=2.2e-3, lambda2=.2e-3, snr=90, save_fods=True),
+        lambda1=2.2e-3, lambda2=.2e-3, snr=30, save_fods=True),
         name='SimulateDWI')
 
     wf = pe.Workflow(name=name)

### 2015-08-12 paper/main.tex
change default SNR. close #19
fixed #28

@@
 
-\newcommand{\e}[1]{\ensuremath{\;\cdot\,}10\ensuremath{^{#1}}}
+\newcommand{\e}[1]{\ensuremath{\;\cdot\,\text{10}^\text{#1}}}
+\newcommand{\vmaps}{\ensuremath{\{\mathbf{V}_i \,|\, i \in \{1,2,3\}\}}}
+\newcommand{\fmaps}{\ensuremath{\{F_i \,|\, i \in \{1,2,3\}\}}}
+\newcommand{\tmaps}{\ensuremath{\{T_j \,|\, j \in \{1,\ldots,5\}\}}}
+\newcommand{\diffunits}{\ensuremath{\text{mm}^2\text{s}^{\text{-1}}}}
+
 \newcommand*\glslines[2][]{\glslink[#1]{#2}{\glsdisp{#2}{\acrlong{#2} --\glsentryshort{#2}--}}}
@@
   in pipelines for the connectome extraction.
-In this report we present \emph{diffantom}, an \emph{in-silico} dataset to assess pipelines for tractography
+In this report we present \emph{Diffantom}, an \emph{in-silico} dataset to assess pipelines for tractography
   and connectome extraction using real data as source of whole-brain microstructural information.
@@
 First, since we use a dataset from the \gls*{hcp} as input, data are corrected for the most relevant distortions.
-\cite{wilkins_fiber_2014} explicitly state that their original data were not subjected to certain distortion
-  corrections, and thus, generated data are affected correspondingly.
+\cite{wilkins_fiber_2014} explicitly state that their original data were not corrected for certain artifacts,
+  and thus, generated data are affected correspondingly.
 The second improvement is a more advanced signal model to generate the phantom using the hindered and restricted
@@
 The simulation process relies on a microstructural model derived from real data.
-On one hand, the software requires up to five fraction maps $\{T_{1\,\cdots\,5}\}$ of
+On one hand, the software requires up to five fraction maps \tmaps{} of
   free- and hindered- diffusion (see \autoref{fig:figure1}, box A).
@@
 \gls*{cgm}, \gls*{dgm}, \gls*{wm}, \gls*{csf}, and abnormal tissue~%
-\footnote{Since here we simulate healthy subjects, the last fraction map is empty and can be omitted}.
-On the other hand, the restricted-diffusion compartments are specified by up to three volume fractions $\{F_{1\,\cdots\,3}\}$
-  of three single fiber populations per voxel along with their corresponding direction maps $\{V_{1\,\cdots\,3}\}$.
+\footnote{Since here we simulate healthy subjects, the last fraction map $F_5$ is empty and can be omitted}.
+On the other hand, the restricted-diffusion compartments are specified by up to three volume fractions \fmaps{}
+  of three single fiber populations per voxel along with their corresponding direction maps \vmaps{}.
 
 The process to obtain the microstructural model can be described like follows:
-1) The fiber orientation maps $\{V_{1\,\cdots\,3}\}$ and their corresponding fiber-fraction maps are
+1) The fiber orientation maps $\{\mathbf{V}_i\}$ and their corresponding fiber-fraction maps $\{F_i\}$ are
   obtained using the ball-and-stick model for multi-shell data of BEDPOSTX \citep{jbabdi_modelbased_2012}
@@
   in \emph{dipy} \citep{garyfallis_dipy_2011}.
-4) The macrostructural 5TT map is extracted from the \acrlong*{t1} image of the subject, using \emph{MRTrix} tools.
+4) The macrostructural 5TT map $\{T'_j\}$ is extracted from the \acrlong*{t1} image of the subject, using \emph{MRTrix} tools.
 5) The objects obtained in the previous steps are combined as described in the \nameref{sec:appendix} to generate the
@@
 Once a microstructural model of the volume has been synthesized, a map of \glspl{fod} is computed using spherical
-  convolution of the single fiber response and the fiber orientation maps $\{V_{1\,\cdots\,3}\}$, weighted by the
-  corresponding fiber-fractions $\{F_{1\,\cdots\,3}\}$.
+  convolution of the single fiber response and the fiber orientation maps $\{\mathbf{V}_i\}$, weighted by the
+  corresponding fiber-fractions $\{F_i\}$.
 A close-up showing how the \glspl{fod} map looks is presented in \autoref{fig:figure1}.
 The single fiber response is a Gaussian diffusion tensor with axial symmetry and eigenvalues $\lambda_1=$ 2.2\e{-3}
-  and $\lambda_{2,3}=$ 0.2\e{-3} [mm$^2$s$^{-1}$].
-The resulting \glspl{fod} map is then combined with the isotropic compartments corresponding to $\{T_{1\,\cdots\,5}\}$
+  \diffunits{} and $\lambda_{2,3}=$ 0.2\e{-3} \diffunits{}.
+The resulting \glspl{fod} map is then combined with the isotropic compartments corresponding to $\{T_j\}$
   before generating the signal using \emph{Phantomas} \citep{caruyer_phantomas_2014}.
@@
   coverage \citep{caruyer_design_2013}.
-Custom one- or multi-shell schemes can be generated supplying the tables of corresponding $b$-vectors and $b$-values.
-Rician noise is also included in \emph{Phantomas}, and can be set by the user.
-The default value for \gls*{snr} is preset to 90.0.
-\Gls*{csf} was simulated with a unique compartment with isotropic diffusivity $D_{CSF}$ of 3.0\e{-3} [mm$^2$s$^{-1}$].
+Custom one- or multi-shell schemes can be generated supplying the tables of corresponding vectors and $b$-values.
+Rician noise is also included in \emph{Phantomas}, and the \gls*{snr} can be set by the user.
+The default value for \gls*{snr} is preset to 30.0.
+\Gls*{csf} was simulated with a unique compartment with isotropic diffusivity $D_{CSF}$ of 3.0\e{-3} \diffunits{}.
 Conversely, \Gls*{wm} and \gls*{gm} included one isotropic compartment with $D_{WM} =$ 2.0\e{-4}, $D_{cGM} =$ 7.0\e{-4}
-  and $D_{dGM} =$ 9.0\e{-4}, respectively [mm$^2$s$^{-1}$].
+  and $D_{dGM} =$ 9.0\e{-4}, respectively [\diffunits{}].
 All these values for diffusivity (and the corresponding to the single-fiber response) can be modified by the user with
@@
 \noindent\textbf{\textit{Interpretation and recommended uses\textcolon}\label{sec:data_use}} %
-To illustrate the features of the \emph{diffantom}, the example dataset underwent a simplified
+To illustrate the features of \emph{Diffantom}, the example dataset underwent a simplified
   connectivity pipeline including \gls*{csd} and probabilistic tractography from
   \emph{MRTrix} \citep{tournier_mrtrix_2012}.
-\Gls*{csd} was configured with up to 8 spherical harmonics, and tractography with 1.6$\times$10$^\text{6}$
+\Gls*{csd} was reconstructed using 8$^\text{th}$-order spherical harmonics, and tractography with 1.6\e{6}
   seed points evenly distributed across a dilated mask of the \gls*{wm} tissue.
@@
     model of choice and on the subsequent global tractography outcome.
-  Since the gradient scheme can be set by the user, \emph{diffantom} can be seen as a mean to translate the so-called
+  Since the gradient scheme can be set by the user, \emph{Diffantom} can be seen as a mean to translate the so-called
   \emph{b-matrix} of the source dataset to any target scheme.
-  \item Assessment of sensitivity and robustness to imaging artifacts (noise, \acrlong{pve} and \glslines{csf} contamination,
+  \item Assessment of sensitivity and robustness to imaging artifacts (noise, \acrlong{pve} and \gls*{csf} contamination,
     susceptibility-derived warping, Eddy-currents-derived distortions, etc.) at unit, integration and overall validation levels
@@
   connectomics.
-The \emph{diffantom} is a derivative of \citep{wilkins_fiber_2014} in terms of methodology for
+\emph{Diffantom} is a derivative of \citep{wilkins_fiber_2014} in terms of methodology for
   simulation with two major advances: the correctness of the \emph{minimally preprocessed} data
@@
   model.
-A possible competitor to \emph{diffantom} is the phantom generated for the Tractography Challenge in
+A possible competitor to \emph{Diffantom} is the phantom generated for the Tractography Challenge in
   ISMRM 2015.
-Similarly to \emph{diffantom}, they organizers used a \gls*{hcp} subject as source of structural information.
-While this phantom is designed for the bundle-wise evaluation of tractography (with scores such as geometrical coverage,
-  valid connections, invalid connections, missed connections, etc. \citep{cote_tractometer_2013}),
-  \emph{diffantom} is intended for the connectome-wise evaluation of results, yielding tractography with
-  a large number of bundles.
-Therefore, \emph{diffantom} and \emph{FiberFox} are complementary as the hypotheses that can be investigated are different.
-Moreover, \emph{diffantom} does not require costly manual segmentation of bundles, highly demanding in terms of physiology
+Similarly to \emph{Diffantom}, they organizers used a \gls*{hcp} subject as source of structural information.
+While this phantom is designed for the bundle-wise evaluation of tractography (with the scores defined in the
+  \emph{Tractometer} \citep{cote_tractometer_2013}, such as geometrical coverage, valid connections, invalid connections,
+  missed connections, etc.), \emph{Diffantom} is intended for the connectome-wise evaluation of results,
+  yielding a tractography with a large number of bundles.
+Therefore, \emph{Diffantom} and \emph{FiberFox} are complementary as the hypotheses that can be investigated are different.
+Moreover, \emph{Diffantom} does not require costly manual segmentation of bundles, highly demanding in terms of physiology
   expertise and operation time.
@@
 \noindent\textbf{\textit{Conclusion\textcolon}}\label{sec:conclusion} %
-\emph{Diffantom} is a digital phantom generated from a dataset belonging to the \acrlong*{hcp}.
-The first \emph{diffantom} is presented here to be openly and freely distributed along with the necessary
+\emph{Diffantom} is a digital phantom generated from a dataset from the \acrlong*{hcp}.
+\emph{Diffantom} is presented here to be openly and freely distributed along with the necessary
   software resources to generate new \emph{diffantoms}.
@@
 \section*{Data Sharing}
-The first \emph{diffantom} is available under the \gls*{cc0} in {\color{red} (dryad?)}.
+The first \emph{Diffantom} is available under the \gls*{cc0} in {\color{red} (dryad?)}.
 The associated software to ``\emph{diffantomize}'' real \gls*{dmri} data is available at \url{https://github.com/oesteban/diffantom}
@@
 \section*{Appendix}\label{sec:appendix}
-Let $\{T'_{1\,\cdots\,5}\}$ be the set of original fractions maps obtained with \texttt{act\_anat\_prepare\_fsl}, a
+Let $\{T'_j\}$ be the set of original fractions maps obtained with \texttt{act\_anat\_prepare\_fsl}, a
   tool in \emph{MRTrix} that combines FAST \citep{zhang_segmentation_2001} and FIRST \citep{patenaude_bayesian_2011}
@@
   maps $V_i$ of restricted diffusion by fibers with their estimated volume fractions $F'_i$ calculated with BEDPOSTX.
-The final $\{T_{1\,\cdots\,5}\}$ maps of isotropic fractions are computed as follows:
+The final $\{T_j\}$ maps of isotropic fractions are computed as follows:
 
@@
   and \gls*{adc} maps, assessed visually.
-The final $\{F_{1\,\cdots\,3}\}$ maps are computed as follows:
+The final $\{F_i\}$ maps are computed as follows:
 
@@
 By default: $w_{f1} = $ 48\%, $w_{f2} = $ 37\%, $w_{f3} = $ 15\%.
-Finally, the resulting maps are normalized to fulfill $\sum_i T_i + \sum_i F_i = 1.0$.
+Finally, the resulting maps are normalized to fulfill $\sum_j T_j + \sum_i F_i = 1.0$.
 
@@
 \textbf{\refstepcounter{figure}\label{fig:figure2} Figure \arabic{figure}. }%
-{\textbf{A. Recommended use of \emph{diffantom}}.
+{\textbf{A. Recommended use of \emph{Diffantom}}.
 The dataset can be used to test algorithms and pipelines against a perturbation that is
-  introduced in \emph{diffantom}.
+  introduced in \emph{Diffantom}.
 Perturbations can model typical artifacts found in \gls*{dmri} datasets or pathological conditions
   such as a tumor.
-\textbf{B1. Tractogram of fibers crossing slice 56 of the \emph{diffantom}} represented over that corresponding
-  slice of the \emph{b0} volume. \textbf{B2. Segmentation of some fiber bundles}: the right \gls*{cst} is
+\textbf{B1. Tractogram of fibers crossing slice 56 of \emph{Diffantom}} as extracted with \emph{MRTrix},
+  represented over that corresponding slice of the \emph{b0} volume.
+\textbf{B2. Segmentation of some fiber bundles using \emph{tract\_querier}}: the right \gls*{cst} is
   represented in blue color, the left \gls*{cst} in red, and the forceps minor in green.
-  The box includes the slice 56 of the \emph{b0} and the pial surface is represented with a light gray shadow.
+  The box includes the slice 56 of the \emph{b0} and the pial surface is represented with transparency.
 }

### 2015-08-12 paper/main.tex
say denoising is important. close #24

@@
 2) A \gls*{fa} map is obtained after fitting a tensor model with \emph{MRTrix}.
-3) The original fiber-fractions and the \gls*{fa} map are smoothed with a non-local means filter included
+3) The original fiber-fractions and the \gls*{fa} map are denoised with a non-local means filter included
   in \emph{dipy} \citep{garyfallis_dipy_2011}.
+This step produces an important smoothing of the maps, while preserving edges in them.
+It is also beneficial in simplifying the voxel-wise diffusion model.
 4) The macrostructural 5TT map $\{T'_j\}$ is extracted from the \acrlong*{t1} image of the subject, using \emph{MRTrix} tools.

### 2015-08-12 paper/main.tex
fix #33

@@
 
-\section*{Disclosure/Conflict-of-Interest Statement}
+\section*{Disclosure}
 
@@
 \end{center}
-\textbf{\refstepcounter{figure}\label{fig:figure1} Figure \arabic{figure}. }%
-{\textbf{Generation of \emph{diffantoms}:} First, a microstructural model is generated as described
-  in the \nameref{sec:data_generation} subsection.
+\textbf{\refstepcounter{figure}\label{fig:figure1} Figure \arabic{figure}. }{%
+\textbf{A. Microstructural model of \emph{Diffantom}}.
+The phantom is simulated from an underlying microstructural model specified with the
+  following volume-fraction maps: three hindered-diffusion compartments $\{T_1, T_2, T_3\}$,
+  one free-diffusion compartment $T_4$ corresponding to the \gls*{csf},
+  three restricted-diffusion compartments $\{F_i\}$, and three vectorial maps
+  associated with the local fiber directions $\{\mathbf{V}_i\}$.
 Please note the piece-wise linear function of the color scale to enable visibility of small volume fractions.
-This microstructural model is fed into a simulation block that generates the final \gls*{dmri} dataset
-  as described in the second block.
+\textbf{B. Workflow to generate \emph{diffantoms}}.
+The pipeline to generate phantoms from any \gls*{hcp} dataset is presented in the lower box.
+Once the microstructural model shown in the upper box has been prepared as described in \nameref{sec:data_model},
+  the local orientations are computed and fed into \emph{Phantomas} to finally generate the signal.
 }

### 2015-08-12 paper/main.tex
Fixing figure 2, #37
refactoring figure 2, close #37

@@
 \begin{center}
-\includegraphics[width=\linewidth]{figures/figure02}
+\includestandalone[width=\linewidth]{figure02}
 \end{center}
@@
 {\textbf{A. Recommended use of \emph{Diffantom}}.
-The dataset can be used to test algorithms and pipelines against a perturbation that is
-  introduced in \emph{Diffantom}.
-Perturbations can model typical artifacts found in \gls*{dmri} datasets or pathological conditions
-  such as a tumor.
+The phantom is designed to be used as ground-truth information in evaluation frameworks,
+  to implement unit test of algorithms, to check integration of processing units within
+  pipelines or to validate complete workflows.
+For instance, in order to evaluate artifacts, a perturbation can be induced in the microstructural
+  model or after simulation to provide reference and test datasets.
 \textbf{B1. Tractogram of fibers crossing slice 56 of \emph{Diffantom}} as extracted with \emph{MRTrix},

### 2015-08-12 paper/main.tex
close #34

@@
 These compartments will be derived from the macroscopic structure of tissues within the brain,
-  specified in the following order~%
-\footnote{Corresponding to the \emph{5TT format} established with the latest version 3.0
-  of \emph{MRTrix} \citep{tournier_mrtrix_2012}}:
-\gls*{cgm}, \gls*{dgm}, \gls*{wm}, \gls*{csf}, and abnormal tissue~%
-\footnote{Since here we simulate healthy subjects, the last fraction map $F_5$ is empty and can be omitted}.
+  specified in the following order\footnote{Corresponding to the \emph{5TT format} established
+  with the latest version 3.0 of \emph{MRTrix} \citep{tournier_mrtrix_2012}}:
+  \gls*{cgm}, \gls*{dgm}, \gls*{wm}, \gls*{csf}, and abnormal tissue\footnote{Since here we
+  simulate healthy subjects, the last fraction map $T_5$ is empty and can be omitted}.
 On the other hand, the restricted-diffusion compartments are specified by up to three volume fractions \fmaps{}
@@
   \diffunits{} and $\lambda_{2,3}=$ 0.2\e{-3} \diffunits{}.
-The resulting \glspl{fod} map is then combined with the isotropic compartments corresponding to $\{T_j\}$
-  before generating the signal using \emph{Phantomas} \citep{caruyer_phantomas_2014}.
+The resulting \glspl{fod} map is then combined with the free- and hindered-diffusion compartments corresponding to $\{T_j\}$.
+The free-diffusion compartment corresponds to the \gls*{csf} fraction map $T_4$ and is modeled with isotropic
+  diffusivity $D_{CSF}$ of 3.0\e{-3} \diffunits{}.
+The hindered-diffusion compartments correspond to $\{T_1,T_2,T_3\}$ and are also modeled with isotropic diffusivity
+  $D_{WM} =$ 2.0\e{-4}, $D_{cGM} =$ 7.0\e{-4} and $D_{dGM} =$ 9.0\e{-4}, respectively [\diffunits{}].
+All these values for diffusivity (and the corresponding to the single-fiber response) can be modified by the user with
+  custom settings.
+The restricted- and hindered- compartments are then fed into \emph{Phantomas} \citep{caruyer_phantomas_2014}
+  and the final \gls*{dmri} signal is obtained.
 By default, diffusion data are generated using a scheme of 100 directions distributed in one shell with uniform
@@
 The default value for \gls*{snr} is preset to 30.0.
-\Gls*{csf} was simulated with a unique compartment with isotropic diffusivity $D_{CSF}$ of 3.0\e{-3} \diffunits{}.
-Conversely, \Gls*{wm} and \gls*{gm} included one isotropic compartment with $D_{WM} =$ 2.0\e{-4}, $D_{cGM} =$ 7.0\e{-4}
-  and $D_{dGM} =$ 9.0\e{-4}, respectively [\diffunits{}].
-All these values for diffusivity (and the corresponding to the single-fiber response) can be modified by the user with
-  custom settings.
 

### 2015-08-12 paper/main.tex
trying to clarify the model. close #17

@@
 
-The process to obtain the microstructural model can be described like follows:
-1) The fiber orientation maps $\{\mathbf{V}_i\}$ and their corresponding fiber-fraction maps $\{F_i\}$ are
-  obtained using the ball-and-stick model for multi-shell data of BEDPOSTX \citep{jbabdi_modelbased_2012}
-  on the \gls*{dmri} data of a certain \gls*{hcp} subject.
+The process to obtain the microstructural model from one dataset of the \gls*{hcp} can be described
+  like follows:
+1) The fiber orientation maps $\{\mathbf{V}_i\}$ and their corresponding estimations of volume fraction $\{F'_i\}$ are
+  obtained using the ball-and-stick model for multi-shell data of \emph{bedpostx} \citep{jbabdi_modelbased_2012}
+  on the \gls*{dmri} data.
+The \gls*{hcp} indicates \emph{bedpostx} to reconstruct their data \citep{glasser_minimal_2013}.
+Additionally, this software was able to exploit the multi-shell acquisitions of the \gls*{hcp} while
+  operating at whole-brain level.
 2) A \gls*{fa} map is obtained after fitting a tensor model with \emph{MRTrix}.
-3) The original fiber-fractions and the \gls*{fa} map are denoised with a non-local means filter included
+The \gls*{fa} is used to infer $F_1$ (the fraction map of the most prevalent fiber), avoiding the extremely
+  noisy estimation of $F'_1$ performed by \emph{bedpostx} in the previous step.
+3) The original fiber fractions $\{F'_i\}$ and the \gls*{fa} map are denoised with a nonlocal means filter included
   in \emph{dipy} \citep{garyfallis_dipy_2011}.
-This step produces an important smoothing of the maps, while preserving edges in them.
-It is also beneficial in simplifying the voxel-wise diffusion model.
-4) The macrostructural 5TT map $\{T'_j\}$ is extracted from the \acrlong*{t1} image of the subject, using \emph{MRTrix} tools.
-5) The objects obtained in the previous steps are combined as described in the \nameref{sec:appendix} to generate the
-  microstructural model, presented in \autoref{fig:figure1}-A.
+This step produces an important smoothing of the maps, while preserving the edges in them.
+Smoothing is also beneficial in simplifying the voxel-wise diffusion model.
+4) The macrostructural fractions $\{T'_j\}$ are extracted from the \acrlong*{t1} image of the dataset,
+  using standard \emph{FSL} segmentation tools \citep{jenkinson_fsl_2012}.
+5) The images obtained previously (\gls*{fa} map, $\{\mathbf{V}_i\}$, $\{F'_i\}$, and $\{T'_j\}$)
+  are combined as described in the \nameref{sec:appendix} to generate the final microstructural model
+  ($\{\mathbf{V}_i\}$, $\{F_i\}$, and $\{T_j\}$), presented in \autoref{fig:figure1}-A.
 

### 2015-08-14 paper/abbreviations.tex
new suggestions from MJL

@@
 % @Last Modified by:   Oscar Esteban
-% @Last Modified time: 2015-07-14 12:49:57
+% @Last Modified time: 2015-08-14 09:16:55
 
@@
 \newacronym{cst}{CST}{corticospinal tract}
+\newacronym{t1}{T1w}{T1-weighted}
 
@@
 	first={Human Connectome Project (HCP, \cite{essen_human_2012})},
-	long={Human Connectome Project (HCP)},
+	long={Human Connectome Project},
 	description={Human Connectome Project}

### 2015-08-14 paper/main.tex
new suggestions from MJL

@@
 
-\title[Diffantom]{Diffantom: a whole-brain diffusion MRI phantom derived from a real dataset}
+\title[Diffantom]{Diffantom: a whole-brain diffusion MRI phantom derived from a real dataset of the \acrlong{hcp}}
 
@@
   of the structural configuration of fiber bundles at the macroscale.
-Tractography is fundamental to gain information about \gls*{wm} morphology in many clinical routines
+Tractography is fundamental to gain information about \gls*{wm} morphology in many clinical applications
   like neurosurgical planning \citep{golby_interactive_2011}, post-surgery evaluations \citep{toda_utility_2014},
@@
   Alzheimer's disease.
-The analysis of structural brain networks using graph theory can also be applied on tractography.
-For instance, in the definition of the unique subject-wise patterns of connectivity
-  \citep{sporns_human_2005}, the assessment of neurological diseases \citep{griffa_structural_2013}, and in the
+The analysis of structural brain networks using graph theory is also applied on tractography,
+  for instance in the definition of the unique subject-wise patterns of connectivity
+  \citep{sporns_human_2005}, in the assessment of neurological diseases \citep{griffa_structural_2013}, and in the
   study of the link between structural and functional connectivity \citep{messe_predicting_2015}.
@@
 \cite{close_software_2009} presented the \emph{Numerical Fibre Generator}, a software to simulate
-  spherical shapes filled with digital fiber-tracts.
+  spherical shapes filled with digital fiber tracts.
 \cite{caruyer_phantomas_2014} proposed \emph{Phantomas} to simulate any kind of analytic geometry
@@
 Using \emph{FiberFox}, the segmentation of each bundle was mapped to an analytical
-  description and finally simulated the signal.
+  description, and finally simulated the signal.
 
@@
   in pipelines for the connectome extraction.
-In this report we present \emph{Diffantom}, an \emph{in-silico} dataset to assess pipelines for tractography
-  and connectome extraction using real data as source of whole-brain microstructural information.
+In this data report we present \emph{Diffantom}, an \emph{in-silico} dataset to assess tractography and connectivity
+  pipelines using \gls*{dmri} real data as source microstructural information.
 \emph{Diffantom} is inspired by the work of \cite{wilkins_fiber_2014}, with two principal novelties.
-First, since we use a dataset from the \gls*{hcp} as input, data are corrected for the most relevant distortions.
-\cite{wilkins_fiber_2014} explicitly state that their original data were not corrected for certain artifacts,
-  and thus, generated data are affected correspondingly.
+First, since we use a dataset from the \gls*{hcp} as input, data are already corrected for the most
+  relevant distortions.
 The second improvement is a more advanced signal model to generate the phantom using the hindered and restricted
   diffusion model of \emph{Phantomas} \citep{caruyer_phantomas_2014}.
+As a result, we provide a whole-brain digital phantom of \gls*{dmri} data with structural information derived
+  from an \gls*{hcp} dataset.
+We also openly release \emph{MkDiffantom}, the software package necessary to generate custom \emph{diffantoms}.
 
@@
 The simulation process relies on a microstructural model derived from real data.
-On one hand, the software requires up to five fraction maps \tmaps{} of
+On one hand, \emph{MkDiffantom} requires up to five fraction maps \tmaps{} of
   free- and hindered- diffusion (see \autoref{fig:figure1}, box A).
@@
 The process to obtain the microstructural model from one dataset of the \gls*{hcp} can be described
-  like follows:
+  as follows (see also \autoref{fig:figure1}, box B):
 1) The fiber orientation maps $\{\mathbf{V}_i\}$ and their corresponding estimations of volume fraction $\{F'_i\}$ are
@@
   on the \gls*{dmri} data.
-The \gls*{hcp} indicates \emph{bedpostx} to reconstruct their data \citep{glasser_minimal_2013}.
-Additionally, this software was able to exploit the multi-shell acquisitions of the \gls*{hcp} while
+The \gls*{hcp} recommends \emph{bedpostx} to reconstruct their data \citep{glasser_minimal_2013}.
+A further advantage is that \emph{bedpostx} exploits the multi-shell acquisitions of the \gls*{hcp} while
   operating at whole-brain level.
@@
   in \emph{dipy} \citep{garyfallis_dipy_2011}.
-This step produces an important smoothing of the maps, while preserving the edges in them.
+This step produces an important smoothing of the maps, while preserving the edges.
 Smoothing is also beneficial in simplifying the voxel-wise diffusion model.
@@
   corresponding fiber-fractions $\{F_i\}$.
-A close-up showing how the \glspl{fod} map looks is presented in \autoref{fig:figure1}.
+A close-up showing how the \glspl{fod} map looks is presented in \autoref{fig:figure1}-B.
 The single fiber response is a Gaussian diffusion tensor with axial symmetry and eigenvalues $\lambda_1=$ 2.2\e{-3}
@@
 \noindent\textbf{\textit{Implementation and reproducibility\textcolon}\label{sec:data_workflow}} %
-We also provide the full software stack used to generate \emph{diffantoms}, so that users can
-  regenerate similar datasets with different parameters.
+We also provide \emph{MkDiffantom}, the software package used to generate \emph{diffantoms}, so
+  that users can regenerate similar datasets with different parameters.
 This workflow, presented in \autoref{fig:figure1}, is implemented using
@@
   seed points evenly distributed across a dilated mask of the \gls*{wm} tissue.
-\autoref{fig:figure2}, box B1 shows the result of the tractography obtained with such pipeline.
+\autoref{fig:figure2}, box A1, shows the result of the tractography obtained with such pipeline.
 Finally, we applied \emph{tract querier} \citep{wassermann_on_2013} on segmenting important fiber bundles such
-  as the \gls*{cst} and the forceps minor (see \autoref{fig:figure2}, box B2).
+  as the \gls*{cst} and the forceps minor (see \autoref{fig:figure2}, box A2).
 Particularly, due to its location nearby the orbitofrontal lobe, the forceps minor is generally affected by
@@
 
-We recommend \emph{Diffantom} as ground-truth in assessment frameworks (\autoref{fig:figure2}, box A)
-  for the unit testing of algorithms, the integration tests of modules in workflows and the full
-  validation of pipelines:
+We recommend \emph{Diffantom} as ground-truth in verification and validation frameworks
+  (\autoref{fig:figure2}, box B) for testing pipelines.
+\emph{Diffantom} is applicable in the unit testing of algorithms, the integration testing of
+  modules in workflows, and the overall system testing.
+Some potential applications follow:
 \begin{itemize}
@@
   \item Assessment of sensitivity and robustness to imaging artifacts (noise, \acrlong{pve} and \gls*{csf} contamination,
-    susceptibility-derived warping, Eddy-currents-derived distortions, etc.) at unit, integration and overall validation levels
-    of evaluation.
-  \item Using the \emph{Diffantom} as in box A of \autoref{fig:figure2}, if is possible to apply binary classification measures
+    susceptibility-derived warping, Eddy-currents-derived distortions, etc.) at unit, integration and systems testing levels.
+  \item Using \emph{Diffantom} as in box B of \autoref{fig:figure2}, if is possible to apply binary classification measures
     to evaluate the resulting connectivity matrix.
-  Considering the connectivity matrix of the reference \emph{Diffantom} as a message to be sent through a pipeline, the resulting
-    matrix of the test \emph{Diffantom} would be the received message, and the \acrfull*{roc} of the pipeline can be
-    characterized.
+  Considering the connectivity matrix of the \emph{reference Diffantom} and the resulting
+    matrix of the\emph{test Diffantom}, the \acrfull*{roc} of the pipeline can be characterized.
   \item Simulation of pathological brains by altering the microstructural
@@
 \emph{Diffantom} is a derivative of \citep{wilkins_fiber_2014} in terms of methodology for
-  simulation with two major advances: the correctness of the \emph{minimally preprocessed} data
-  \citep{glasser_minimal_2013} released within the \gls*{hcp}, and a more realistic diffusion
-  model.
+  simulation with two major advances.
+First, the correctness of the \emph{minimally preprocessed} data \citep{glasser_minimal_2013}
+  released within the \gls*{hcp}.
+\cite{wilkins_fiber_2014} explicitly state that their original data were not corrected for certain artifacts,
+  and thus, generated data are affected correspondingly.
+Second, \emph{Diffantom} implements the hindered and restricted compartments model \citep{assaf_composite_2005},
+  which is a more complete model than the multi-tensor diffusion model.
+
 A possible competitor to \emph{Diffantom} is the phantom generated for the Tractography Challenge in
@@
   expertise and operation time.
-This facilitates the reproducibility of the dataset by using the associated software workflow to easily
-  generate custom \emph{diffantoms}.
+The software workflow released with this data report (\emph{MkDiffantom}) ensures the reproducibility of
+  \emph{Diffantom} and enables the generation of custom \emph{diffantoms}.
+\emph{MkDiffantom} is designed for, but not limited to, use \gls*{hcp} datasets as source of structural information.
 
 \noindent\textbf{\textit{Conclusion\textcolon}}\label{sec:conclusion} %
-\emph{Diffantom} is a digital phantom generated from a dataset from the \acrlong*{hcp}.
-\emph{Diffantom} is presented here to be openly and freely distributed along with the necessary
-  software resources to generate new \emph{diffantoms}.
+\emph{Diffantom} is a whole-brain digital phantom generated from a dataset from the \acrlong*{hcp}.
+\emph{Diffantom} is presented here to be openly and freely distributed along with \emph{MkDiffantom}
+  to generate new \emph{diffantoms}.
 We encourage the neuroimage community to contribute with their own \emph{diffantoms} and share them openly.
@@
   to generate the macrostructural 5TT map.
-FA denotes the \gls*{fa} map obtained from the original \gls*{dmri} data, and $\{V_{1\,\cdots\,3}, F'_{1\,\cdots\,3}\}$ the direction
-  maps $V_i$ of restricted diffusion by fibers with their estimated volume fractions $F'_i$ calculated with BEDPOSTX.
+FA denotes the \gls*{fa} map obtained from the original \gls*{dmri} data: the local fiber orientation maps
+  $\{\mathbf{V}_i\}$ with their estimated volume fractions $\{F'_i\}$ calculated with \emph{bedpostx}.
 The final $\{T_j\}$ maps of isotropic fractions are computed as follows:
@@
 Please note the piece-wise linear function of the color scale to enable visibility of small volume fractions.
-\textbf{B. Workflow to generate \emph{diffantoms}}.
+\textbf{B. \emph{MkDiffantom}, a workflow to generate \emph{diffantoms}}.
 The pipeline to generate phantoms from any \gls*{hcp} dataset is presented in the lower box.
 Once the microstructural model shown in the upper box has been prepared as described in \nameref{sec:data_model},
-  the local orientations are computed and fed into \emph{Phantomas} to finally generate the signal.
+  the local orientations are computed and fed into \emph{Phantomas} to finally simulate the signal.
 }
@@
 \end{center}
-\textbf{\refstepcounter{figure}\label{fig:figure2} Figure \arabic{figure}. }%
-{\textbf{A. Recommended use of \emph{Diffantom}}.
+\textbf{\refstepcounter{figure}\label{fig:figure2} Figure \arabic{figure}. }{%
+\textbf{A1. Tractogram of fibers crossing slice 56 of \emph{Diffantom}} as extracted with \emph{MRTrix},
+  represented over that corresponding slice of the \emph{b0} volume.
+\textbf{A2. Segmentation of some fiber bundles using \emph{tract\_querier}}: the right \gls*{cst} is
+  represented in blue color, the left \gls*{cst} in red, and the forceps minor in green.
+  The box includes the slice 56 of the \emph{b0} and the pial surface is represented with transparency.
+\textbf{B. Recommended use of \emph{Diffantom}}.
 The phantom is designed to be used as ground-truth information in evaluation frameworks,
@@
   model or after simulation to provide reference and test datasets.
-\textbf{B1. Tractogram of fibers crossing slice 56 of \emph{Diffantom}} as extracted with \emph{MRTrix},
-  represented over that corresponding slice of the \emph{b0} volume.
-\textbf{B2. Segmentation of some fiber bundles using \emph{tract\_querier}}: the right \gls*{cst} is
-  represented in blue color, the left \gls*{cst} in red, and the forceps minor in green.
-  The box includes the slice 56 of the \emph{b0} and the pial surface is represented with transparency.
 }

### 2015-08-26 paper/abbreviations.tex
latest suggestions from AS

@@
 % @Last Modified by:   Oscar Esteban
-% @Last Modified time: 2015-08-14 09:16:55
+% @Last Modified time: 2015-08-26 17:43:01
 
@@
 \newacronym{fod}{FOD}{fiber orientation distribution}
+\newacronym{fodf}{fODF}{fiber orientation distribution function}
 \newacronym{wm}{WM}{white matter}
@@
 
+\newglossaryentry{bedpostx}%
+{
+	name={BEDPOSTX},
+	first={BEDPOSTX (Bayesian Estimation of Diffusion Parameters Obtained using Sampling Techniques modelling crossing --X-- fibres, \cite{jbabdi_modelbased_2012})},
+	description={Bayesian Estimation of Diffusion Parameters Obtained using Sampling Techniques modelling crossing --X-- fibres}
+}
+\newacronym[first={FAST (FMRIB's Automated Segmentation Tool, \cite{zhang_segmentation_2001})}]%
+{fast}{FAST}{FMRIB's Automated Segmentation Tool}
+\newacronym[first={FIRST (FMRIB's Integrated Registration and Segmentation Tool, \cite{patenaude_bayesian_2011})}]%
+{first}{FIRST}{FMRIB's Integrated Registration and Segmentation Tool}
+
 \makeglossaries

### 2015-08-26 paper/main.tex
latest suggestions from AS

@@
 
+\newcommand{\lowb}{\textit{b0}}
 \newcommand{\e}[1]{\ensuremath{\;\cdot\,\text{10}^\text{#1}}}
@@
 
-\title[Diffantom]{Diffantom: a whole-brain diffusion MRI phantom derived from a real dataset of the \acrlong{hcp}}
+\title[Diffantom]{Diffantom: a whole-brain diffusion MRI phantoms derived from real datasets of the \acrlong{hcp}}
 
@@
 \section*{Introduction}
-Fiber tracking on \gls*{dmri} data has become an important application for the \textit{in-vivo} investigation
+Fiber tracking on \gls*{dmri} data has become an important tool for the \textit{in-vivo} investigation
   of the structural configuration of fiber bundles at the macroscale.
@@
   can not be accounted for with current materials and proposed methodologies to build physical phantoms.
-Early digital phantoms also started with simulation of simple geometries
+Early digital phantoms started with simulation of simple geometries
   \citep{basser_in_2000,goessl_fiber_2002,tournier_limitations_2002,leemans_mathematical_2005}
@@
 
-\emph{Diffantom} is originally designed for the investigation of susceptibility-derived distortions, a
-  typical artifact that produces geometrical warping in certain regions of \gls*{dmri} datasets.
-In \citep{esteban_simulationbased_2014} we addressed this phenomenon and concluded that the connectivity
-  matrix of \emph{Phantomas} was not dense enough to evaluate the integration of correction methods
-  in pipelines for the connectome extraction.
 In this data report we present \emph{Diffantom}, an \emph{in-silico} dataset to assess tractography and connectivity
@@
   from an \gls*{hcp} dataset.
-We also openly release \emph{MkDiffantom}, the software package necessary to generate custom \emph{diffantoms}.
+We also openly release the \emph{diffantomizer} workflow, the software package necessary to generate custom \emph{diffantoms}.
+\emph{Diffantom} is originally designed for the investigation of susceptibility-derived distortions, a
+  typical artifact that produces geometrical warping in certain regions of \gls*{dmri} datasets.
+In \citep{esteban_simulationbased_2014} we addressed this phenomenon and concluded that the connectivity
+  matrix of \emph{Phantomas} was not dense enough to evaluate the integration of correction methods
+  in pipelines for the connectome extraction.
 
@@
 The simulation process relies on a microstructural model derived from real data.
-On one hand, \emph{MkDiffantom} requires up to five fraction maps \tmaps{} of
+On one hand, the \emph{diffantomizer} workflow requires up to five fraction maps \tmaps{} of
   free- and hindered- diffusion (see \autoref{fig:figure1}, box A).
@@
 1) The fiber orientation maps $\{\mathbf{V}_i\}$ and their corresponding estimations of volume fraction $\{F'_i\}$ are
-  obtained using the ball-and-stick model for multi-shell data of \emph{bedpostx} \citep{jbabdi_modelbased_2012}
+  obtained using the ball-and-stick model for multi-shell data of \gls*{bedpostx}
   on the \gls*{dmri} data.
-The \gls*{hcp} recommends \emph{bedpostx} to reconstruct their data \citep{glasser_minimal_2013}.
-A further advantage is that \emph{bedpostx} exploits the multi-shell acquisitions of the \gls*{hcp} while
+The \gls*{hcp} recommends \gls*{bedpostx} to reconstruct their data \citep{glasser_minimal_2013}.
+A further advantage is that \gls*{bedpostx} exploits the multi-shell acquisitions of the \gls*{hcp} while
   operating at whole-brain level.
 2) A \gls*{fa} map is obtained after fitting a tensor model with \emph{MRTrix}.
-The \gls*{fa} is used to infer $F_1$ (the fraction map of the most prevalent fiber), avoiding the extremely
-  noisy estimation of $F'_1$ performed by \emph{bedpostx} in the previous step.
+As we shall see in the Appendix, the \gls*{fa} is used to infer $F_1$ (the fraction map of the most prevalent fiber),
+  avoiding the extremely noisy estimation of $F'_1$ performed by \gls*{bedpostx} in the previous step.
 3) The original fiber fractions $\{F'_i\}$ and the \gls*{fa} map are denoised with a nonlocal means filter included
-  in \emph{dipy} \citep{garyfallis_dipy_2011}.
+  in \emph{dipy} \citep{garyfallidis_dipy_2014}.
 This step produces an important smoothing of the maps, while preserving the edges.
@@
 \noindent\textbf{\textit{Diffusion signal generation\textcolon}\label{sec:data_dwi}} %
-Once a microstructural model of the volume has been synthesized, a map of \glspl{fod} is computed using spherical
+Once a microstructural model of the volume has been synthesized, a map of \glspl{fodf} is computed using spherical
   convolution of the single fiber response and the fiber orientation maps $\{\mathbf{V}_i\}$, weighted by the
   corresponding fiber-fractions $\{F_i\}$.
-A close-up showing how the \glspl{fod} map looks is presented in \autoref{fig:figure1}-B.
+A close-up showing how the \glspl{fodf} map looks is presented in \autoref{fig:figure1}-B.
 The single fiber response is a Gaussian diffusion tensor with axial symmetry and eigenvalues $\lambda_1=$ 2.2\e{-3}
   \diffunits{} and $\lambda_{2,3}=$ 0.2\e{-3} \diffunits{}.
-The resulting \glspl{fod} map is then combined with the free- and hindered-diffusion compartments corresponding to $\{T_j\}$.
+The resulting \glspl{fodf} map is then combined with the free- and hindered-diffusion compartments corresponding to $\{T_j\}$.
 The free-diffusion compartment corresponds to the \gls*{csf} fraction map $T_4$ and is modeled with isotropic
@@
 \noindent\textbf{\textit{Implementation and reproducibility\textcolon}\label{sec:data_workflow}} %
-We also provide \emph{MkDiffantom}, the software package used to generate \emph{diffantoms}, so
+We also provide the \emph{diffantomizer} workflow, the software package used to generate \emph{diffantoms}, so
   that users can regenerate similar datasets with different parameters.
@@
     susceptibility-derived warping, Eddy-currents-derived distortions, etc.) at unit, integration and systems testing levels.
-  \item Using \emph{Diffantom} as in box B of \autoref{fig:figure2}, if is possible to apply binary classification measures
+  \item Using \emph{Diffantom} as in box B of \autoref{fig:figure2}, it is possible to apply binary classification measures
     to evaluate the resulting connectivity matrix.
   Considering the connectivity matrix of the \emph{reference Diffantom} and the resulting
-    matrix of the\emph{test Diffantom}, the \acrfull*{roc} of the pipeline can be characterized.
+    matrix of the \emph{test Diffantom}, the \acrfull*{roc} of the pipeline can be characterized.
   \item Simulation of pathological brains by altering the microstructural
@@
   ISMRM 2015.
-Similarly to \emph{Diffantom}, they organizers used a \gls*{hcp} subject as source of structural information.
+Similarly to \emph{Diffantom}, the organizers used an \gls*{hcp} subject as source of structural information.
 While this phantom is designed for the bundle-wise evaluation of tractography (with the scores defined in the
@@
   expertise and operation time.
-The software workflow released with this data report (\emph{MkDiffantom}) ensures the reproducibility of
+The software workflow released with this data report (the \emph{diffantomizer}) ensures the reproducibility of
   \emph{Diffantom} and enables the generation of custom \emph{diffantoms}.
-\emph{MkDiffantom} is designed for, but not limited to, use \gls*{hcp} datasets as source of structural information.
+The \emph{diffantomizer} is designed for, but not limited to, use \gls*{hcp} datasets as source of structural information.
 
@@
 \emph{Diffantom} is a whole-brain digital phantom generated from a dataset from the \acrlong*{hcp}.
-\emph{Diffantom} is presented here to be openly and freely distributed along with \emph{MkDiffantom}
+\emph{Diffantom} is presented here to be openly and freely distributed along with The \emph{diffantomizer} workflow
   to generate new \emph{diffantoms}.
@@
 Let $\{T'_j\}$ be the set of original fractions maps obtained with \texttt{act\_anat\_prepare\_fsl}, a
-  tool in \emph{MRTrix} that combines FAST \citep{zhang_segmentation_2001} and FIRST \citep{patenaude_bayesian_2011}
+  tool in \emph{MRTrix} that combines \gls*{fast} and \gls*{first}
   to generate the macrostructural 5TT map.
 FA denotes the \gls*{fa} map obtained from the original \gls*{dmri} data: the local fiber orientation maps
-  $\{\mathbf{V}_i\}$ with their estimated volume fractions $\{F'_i\}$ calculated with \emph{bedpostx}.
+  $\{\mathbf{V}_i\}$ with their estimated volume fractions $\{F'_i\}$ calculated with \gls*{bedpostx}.
 The final $\{T_j\}$ maps of isotropic fractions are computed as follows:
@@
 Please note the piece-wise linear function of the color scale to enable visibility of small volume fractions.
-\textbf{B. \emph{MkDiffantom}, a workflow to generate \emph{diffantoms}}.
+\textbf{B. The \emph{diffantomizer} workflow, a workflow to generate \emph{diffantoms}}.
 The pipeline to generate phantoms from any \gls*{hcp} dataset is presented in the lower box.
@@
 \textbf{\refstepcounter{figure}\label{fig:figure2} Figure \arabic{figure}. }{%
-\textbf{A1. Tractogram of fibers crossing slice 56 of \emph{Diffantom}} as extracted with \emph{MRTrix},
-  represented over that corresponding slice of the \emph{b0} volume.
-\textbf{A2. Segmentation of some fiber bundles using \emph{tract\_querier}}: the right \gls*{cst} is
-  represented in blue color, the left \gls*{cst} in red, and the forceps minor in green.
-  The box includes the slice 56 of the \emph{b0} and the pial surface is represented with transparency.
+\textbf{A1. Example dataset.} The tractogram of fibers crossing slice 56 of \emph{Diffantom} as
+  extracted with \emph{MRTrix} is presented, represented over the corresponding slice of the
+  \lowb{} volume.
+\textbf{A2. Segmentation of some fiber bundles in the example dataset.} We used \emph{tract\_querier} to
+  segment the right \gls*{cst} represented with blue streamlines, the left \gls*{cst} (red streamlines),
+  and the forceps minor (green streamlines).
+  The box includes the slice 56 of the \lowb{} and the pial surface is represented with transparency.
 \textbf{B. Recommended use of \emph{Diffantom}}.

### 2015-08-26 paper/main.tex
fix typo in title
fixes #25

@@
 
-\title[Diffantom]{Diffantom: a whole-brain diffusion MRI phantoms derived from real datasets of the \acrlong{hcp}}
+\title[Diffantom]{Diffantom: whole-brain diffusion MRI phantoms derived from real datasets of the \acrlong{hcp}}
 
@@
 \noindent\textbf{\textit{Diffusion signal generation\textcolon}\label{sec:data_dwi}} %
-Once a microstructural model of the volume has been synthesized, a map of \glspl{fodf} is computed using spherical
-  convolution of the single fiber response and the fiber orientation maps $\{\mathbf{V}_i\}$, weighted by the
-  corresponding fiber-fractions $\{F_i\}$.
-A close-up showing how the \glspl{fodf} map looks is presented in \autoref{fig:figure1}-B.
+Once a microstructural model of the subject has been synthesized, the fiber orientation maps $\{\mathbf{V}_i\}$
+  are weighted by the fiber-fraction maps $\{F_i\}$ and projected onto a continuous representation of
+  the \glspl{fod}.
+A close-up showing how the \glspl{fod} map looks is presented in \autoref{fig:figure1}B.
 The single fiber response is a Gaussian diffusion tensor with axial symmetry and eigenvalues $\lambda_1=$ 2.2\e{-3}
   \diffunits{} and $\lambda_{2,3}=$ 0.2\e{-3} \diffunits{}.
-The resulting \glspl{fodf} map is then combined with the free- and hindered-diffusion compartments corresponding to $\{T_j\}$.
+The resulting \glspl{fod} map is then combined with the free- and hindered-diffusion compartments corresponding to $\{T_j\}$.
 The free-diffusion compartment corresponds to the \gls*{csf} fraction map $T_4$ and is modeled with isotropic

### 2015-09-09 diffantom/diffantom.py
added bedpostx, working!

@@ -84,7 +84,7 @@ if __name__ == '__main__':
         plugin_args = {'n_proc': nthreads, 'maxtasksperchild': 4}
 
     if opts.mode == 'model':
-        wf = gen_model(settings=settings)
+        wf = gen_model(opts.name, settings=settings)
     else:
         wf = gen_diffantom(opts.name, settings=settings)
 

### 2015-09-09 diffantom/workflows.py

@@ -462,9 +462,9 @@ def track_querier(name='TractQuery'):
     return wf
 
 
-def gen_model(settings={}):
+def gen_model(name='GenerateModel', settings={}):
     from nipype.workflows.dmri.fsl.dti import bedpostx_parallel
-    wf = pe.Workflow(name='GenerateModel')
+    wf = pe.Workflow(name=name)
 
     inputnode = pe.Node(niu.IdentityInterface(
         fields=['subject_id', 'data_dir']), name='inputnode')
@@ -487,7 +487,8 @@ def gen_model(settings={}):
                                 for k in ds_tpl_args.keys()}
     ds.inputs.template_args = ds_tpl_args
 
-    bpx = bedpostx_parallel(params={'n_fibres': 3, 'model': 2})
+    bpx = bedpostx_parallel(compute_all_outputs=False,
+                            params={'n_fibres': 3, 'model': 2})
     wf.connect([
         (inputnode, ds,         [('subject_id', 'subject_id'),
                                  ('data_dir', 'base_directory')]),

### 2015-09-09 diffantom/diffantom.py
add workflow for finf

@@ -2,8 +2,8 @@
 # -*- coding: utf-8 -*-
 # @Author: Oscar Esteban
 # @Date:   2015-06-25 15:46:08
-# @Last Modified by:   oesteban
-# @Last Modified time: 2015-08-06 11:13:56
+# @Last Modified by:   Oscar Esteban
+# @Last Modified time: 2015-09-09 15:18:50
 
 """
 ============================
@@ -39,7 +39,7 @@ if __name__ == '__main__':
                             formatter_class=RawTextHelpFormatter)
 
     g_input = parser.add_argument_group('Input')
-    g_input.add_argument('mode', choices=['evaluation', 'model'],
+    g_input.add_argument('mode', choices=['evaluation', 'model', 'bundles'],
                          default='evaluation')
     g_input.add_argument('-S', '--subjects_dir', action='store',
                          default=os.getenv('NEURO_DATA_HOME',
@@ -85,8 +85,10 @@ if __name__ == '__main__':
 
     if opts.mode == 'model':
         wf = gen_model(opts.name, settings=settings)
-    else:
+    elif opts.mode == 'evaluation':
         wf = gen_diffantom(opts.name, settings=settings)
+    elif opts.mode == 'bundles':
+        wf = finf_bundles(opts.name, settings=settings)
 
     wf.base_dir = settings['work_dir']
     # wf.write_graph(format='pdf')

### 2015-09-09 diffantom/workflows.py

@@ -3,7 +3,7 @@
 # @Author: oesteban
 # @Date:   2015-06-23 12:32:07
 # @Last Modified by:   Oscar Esteban
-# @Last Modified time: 2015-08-12 15:37:54
+# @Last Modified time: 2015-09-09 15:25:12
 
 import os
 import os.path as op
@@ -68,6 +68,44 @@ def gen_diffantom(name='Diffantom', settings={}):
         (sim_mod,   trk,      [
             ('outputnode.out_5tt', 'inputnode.in_5tt'),
             ('outputnode.parcellation', 'inputnode.parcellation')])
+    ])
+    return wf
+
+
+def finf_bundles(name='FINFBundles', settings={}):
+    inputnode = pe.Node(niu.IdentityInterface(
+        fields=['subject_id', 'data_dir']), name='inputnode')
+    inputnode.inputs.subject_id = settings['subject_id']
+    inputnode.inputs.data_dir = settings['data_dir']
+
+    out_fields = ['out_file', 'out_fa', 'out_adc', 'out_tdi', 'out_map',
+                  'out_cst']
+
+    outputnode = pe.Node(
+        niu.IdentityInterface(fields=out_fields), name='outputnode')
+
+    fnames = dict(in_dwi='dwi.nii.gz',
+                  in_grad='grads.txt',
+                  in_5tt='in_5tt.nii.gz',
+                  parcellation='parc.nii.gz',
+                  aparc='aparc+aseg.nii.gz',
+                  wmparc='wmparc.nii.gz')
+
+    ds_tpl_args = {k: [['subject_id', [v]]] for k, v in fnames.iteritems()}
+
+    ds = pe.Node(nio.DataGrabber(
+        infields=['subject_id'], sort_filelist=True, template='*',
+        outfields=ds_tpl_args.keys()), name='DataSource')
+    ds.inputs.field_template = {k: 'models/%s/%s'
+                                for k in ds_tpl_args.keys()}
+    ds.inputs.template_args = ds_tpl_args
+
+    trk = act_workflow()
+
+    wf = pe.Workflow(name=name)
+    wf.connect([
+        (ds, trk, [(f, 'inputnode.' + f) for f in fnames.keys()]),
+        (trk, outputnode, [('outputnode.' + f, f) for f in out_fields])
     ])
     return wf
 

### 2015-09-09 diffantom/diffantom.py
ultimate changes

@@ -3,7 +3,7 @@
 # @Author: Oscar Esteban
 # @Date:   2015-06-25 15:46:08
 # @Last Modified by:   Oscar Esteban
-# @Last Modified time: 2015-09-09 15:18:50
+# @Last Modified time: 2015-09-09 15:45:41
 
 """
 ============================
@@ -29,7 +29,7 @@ __status__ = "Prototype"
 if __name__ == '__main__':
     from argparse import ArgumentParser
     from argparse import RawTextHelpFormatter
-    from workflows import gen_diffantom, gen_model
+    from workflows import gen_diffantom, gen_model, finf_bundles
     import glob
     import os
     import os.path as op

### 2015-09-09 diffantom/workflows.py

@@ -3,7 +3,7 @@
 # @Author: oesteban
 # @Date:   2015-06-23 12:32:07
 # @Last Modified by:   Oscar Esteban
-# @Last Modified time: 2015-09-09 15:25:12
+# @Last Modified time: 2015-09-09 15:44:29
 
 import os
 import os.path as op
@@ -84,11 +84,11 @@ def finf_bundles(name='FINFBundles', settings={}):
     outputnode = pe.Node(
         niu.IdentityInterface(fields=out_fields), name='outputnode')
 
-    fnames = dict(in_dwi='dwi.nii.gz',
+    fnames = dict(in_dwi='sim_dwi.nii.gz',
                   in_grad='grads.txt',
-                  in_5tt='in_5tt.nii.gz',
-                  parcellation='parc.nii.gz',
+                  in_5tt='act5tt.nii.gz',
                   aparc='aparc+aseg.nii.gz',
+                  parcellation='aparc+aseg.nii.gz',
                   wmparc='wmparc.nii.gz')
 
     ds_tpl_args = {k: [['subject_id', [v]]] for k, v in fnames.iteritems()}
@@ -96,7 +96,7 @@ def finf_bundles(name='FINFBundles', settings={}):
     ds = pe.Node(nio.DataGrabber(
         infields=['subject_id'], sort_filelist=True, template='*',
         outfields=ds_tpl_args.keys()), name='DataSource')
-    ds.inputs.field_template = {k: 'models/%s/%s'
+    ds.inputs.field_template = {k: 'subjects/%s/%s'
                                 for k in ds_tpl_args.keys()}
     ds.inputs.template_args = ds_tpl_args
 

### 2015-09-09 diffantom/diffantom.py
fixed error in config file for mrtrix3

@@ -3,7 +3,7 @@
 # @Author: Oscar Esteban
 # @Date:   2015-06-25 15:46:08
 # @Last Modified by:   Oscar Esteban
-# @Last Modified time: 2015-09-09 15:45:41
+# @Last Modified time: 2015-09-09 15:50:20
 
 """
 ============================
@@ -62,7 +62,7 @@ if __name__ == '__main__':
 
     settings = {}
     settings['work_dir'] = opts.work_dir
-    settings['data_dir'] = opts.subjects_dir
+    settings['data_dir'] = op.abspath(opts.subjects_dir)
     settings['subject_id'] = opts.subject
 
     # Setup work_dir

### 2015-09-09 diffantom/workflows.py

@@ -3,7 +3,7 @@
 # @Author: oesteban
 # @Date:   2015-06-23 12:32:07
 # @Last Modified by:   Oscar Esteban
-# @Last Modified time: 2015-09-09 15:44:29
+# @Last Modified time: 2015-09-09 16:21:22
 
 import os
 import os.path as op
@@ -85,7 +85,7 @@ def finf_bundles(name='FINFBundles', settings={}):
         niu.IdentityInterface(fields=out_fields), name='outputnode')
 
     fnames = dict(in_dwi='sim_dwi.nii.gz',
-                  in_grad='grads.txt',
+                  in_scheme='grad.txt',
                   in_5tt='act5tt.nii.gz',
                   aparc='aparc+aseg.nii.gz',
                   parcellation='aparc+aseg.nii.gz',
@@ -104,6 +104,8 @@ def finf_bundles(name='FINFBundles', settings={}):
 
     wf = pe.Workflow(name=name)
     wf.connect([
+        (inputnode, ds,       [('subject_id', 'subject_id'),
+                               ('data_dir', 'base_directory')]),
         (ds, trk, [(f, 'inputnode.' + f) for f in fnames.keys()]),
         (trk, outputnode, [('outputnode.' + f, f) for f in out_fields])
     ])
@@ -312,6 +314,9 @@ def act_workflow(name='Tractography'):
 
     lc = pe.Node(mrt3.LabelConfig(), name='LabelConfig')
     lc.inputs.out_file = 'parcellation.nii.gz'
+    lc.inputs.in_config = op.join(
+        os.getenv('MRTRIX3_HOME', '/home/oesteban/workspace/mrtrix3'),
+        'src/connectome/config/fs_default.txt')
     lc.inputs.lut_fs = op.join(
         os.getenv('FREESURFER_HOME'), 'FreeSurferColorLUT.txt')
 

### 2015-09-09 paper/main.tex
latest changes, v2.3

@@
   \citep{basser_in_2000,goessl_fiber_2002,tournier_limitations_2002,leemans_mathematical_2005}
-  to evaluate resolution of fiber crossings.
+  to evaluate the angular resolution as well.
 These tools generally implemented the multi-tensor model \citep{alexander_analysis_2001,tuch_high_2002}
@@
 On one hand, the \emph{diffantomizer} workflow requires up to five fraction maps \tmaps{} of
-  free- and hindered- diffusion (see \autoref{fig:figure1}, box A).
+  free- and hindered- diffusion (see \autoref{fig:figure1}, panel A).
 These compartments will be derived from the macroscopic structure of tissues within the brain,
@@
 The process to obtain the microstructural model from one dataset of the \gls*{hcp} can be described
-  as follows (see also \autoref{fig:figure1}, box B):
+  as follows (see also \autoref{fig:figure1}, panel B):
 1) The fiber orientation maps $\{\mathbf{V}_i\}$ and their corresponding estimations of volume fraction $\{F'_i\}$ are
@@
   seed points evenly distributed across a dilated mask of the \gls*{wm} tissue.
-\autoref{fig:figure2}, box A1, shows the result of the tractography obtained with such pipeline.
+\autoref{fig:figure2}, panels A1 and A3, show the result of the tractography obtained with such pipeline for
+  the original \emph{Diffantom} and a distorted version.
 Finally, we applied \emph{tract querier} \citep{wassermann_on_2013} on segmenting important fiber bundles such
-  as the \gls*{cst} and the forceps minor (see \autoref{fig:figure2}, box A2).
+  as the \gls*{cst} and the forceps minor (see \autoref{fig:figure2}, panels A2, A4).
 Particularly, due to its location nearby the orbitofrontal lobe, the forceps minor is generally affected by
@@
 We recommend \emph{Diffantom} as ground-truth in verification and validation frameworks
-  (\autoref{fig:figure2}, box B) for testing pipelines.
+  (\autoref{fig:figure2}, panel B) for testing pipelines.
 \emph{Diffantom} is applicable in the unit testing of algorithms, the integration testing of
@@
     susceptibility-derived warping, Eddy-currents-derived distortions, etc.) at unit, integration and systems testing levels.
-  \item Using \emph{Diffantom} as in box B of \autoref{fig:figure2}, it is possible to apply binary classification measures
+  \item Using \emph{Diffantom} as in panel B of \autoref{fig:figure2}, it is possible to apply binary classification measures
     to evaluate the resulting connectivity matrix.
@@
 \textbf{B. The \emph{diffantomizer} workflow, a workflow to generate \emph{diffantoms}}.
-The pipeline to generate phantoms from any \gls*{hcp} dataset is presented in the lower box.
-Once the microstructural model shown in the upper box has been prepared as described in \nameref{sec:data_model},
+The pipeline to generate phantoms from any \gls*{hcp} dataset is presented in the lower panel.
+Once the microstructural model shown in the upper panel has been prepared as described in \nameref{sec:data_model},
   the local orientations are computed and fed into \emph{Phantomas} to finally simulate the signal.
@@
 \textbf{\refstepcounter{figure}\label{fig:figure2} Figure \arabic{figure}. }{%
-\textbf{A1. Example dataset.} The tractogram of fibers crossing slice 56 of \emph{Diffantom} as
-  extracted with \emph{MRTrix} is presented, represented over the corresponding slice of the
-  \lowb{} volume.
-\textbf{A2. Segmentation of some fiber bundles in the example dataset.} We used \emph{tract\_querier} to
-  segment the right \gls*{cst} represented with blue streamlines, the left \gls*{cst} (red streamlines),
-  and the forceps minor (green streamlines).
-  The box includes the slice 56 of the \lowb{} and the pial surface is represented with transparency.
+\textbf{A. Example dataset.}
+A1 and A3 show the tractogram of fibers crossing slice 56 of \emph{Diffantom} as
+  extracted with \emph{MRTrix}, represented over the corresponding slice of the
+  \lowb{} volume for the original (A1) and the distorted (A3) phantoms, with a gray
+  frame highlighting the absence of important tracks.
+Panels A2 and A4 show the segmentation of the right \gls*{cst} represented with blue
+  streamlines, the left \gls*{cst} (red streamlines), and the forceps minor (green streamlines)
+  using \emph{tract\_querier}.
+A2 and A4 include the slice 56 of the \lowb{} and the pial surface is represented
+  with transparency.
+In the distorted \emph{Diffantom} (A4) the forceps minor was not detected.
 \textbf{B. Recommended use of \emph{Diffantom}}.

### 2015-09-15 paper/main.tex
final version for bioRxiv

@@
 \documentclass[english]{frontiers/frontiersSCNS} % for Science, Engineering and Humanities and Social Sciences articles
-%\documentclass{frontiersHLTH} % for Health articles
-%\documentclass{frontiersFPHY} % for Physics and Applied Mathematics and Statistics articles
+\usepackage[mode=buildnew]{standalone}
 
-\usepackage[mode=buildnew]{standalone}
 \usepackage{tikz}
@@
   the original \emph{Diffantom} and a distorted version.
-Finally, we applied \emph{tract querier} \citep{wassermann_on_2013} on segmenting important fiber bundles such
+Finally, we applied \emph{tract querier} \citep{wassermann_on_2013} to segment some fiber bundles such
   as the \gls*{cst} and the forceps minor (see \autoref{fig:figure2}, panels A2, A4).
@@
 \emph{Diffantom} is a whole-brain digital phantom generated from a dataset from the \acrlong*{hcp}.
-\emph{Diffantom} is presented here to be openly and freely distributed along with The \emph{diffantomizer} workflow
+\emph{Diffantom} is presented here to be openly and freely distributed along with the \emph{diffantomizer} workflow
   to generate new \emph{diffantoms}.
@@
 \section*{Data Sharing}
-The first \emph{Diffantom} is available under the \gls*{cc0} in {\color{red} (dryad?)}.
-The associated software to ``\emph{diffantomize}'' real \gls*{dmri} data is available at \url{https://github.com/oesteban/diffantom}
+The first \emph{Diffantom} and its distorted version are available under the \gls*{cc0} using the Dryad Digital Repository
+  {\color{red} (reference here when published)}.
+The package is organized following the \gls*{bids} standard.
+The associated software to ``\emph{diffantomize}'' real \gls*{dmri} datasets is available at \url{https://github.com/oesteban/diffantom}
   under an MIT license.

### 2015-10-01 paper/main.tex
insert Dryad infos

@@
 \usepackage[mode=buildnew]{standalone}
-
 \usepackage{tikz}
+\usepackage[framemethod=TikZ]{mdframed}
 %\setcitestyle{square}
@@
 \maketitle
+
+\begin{mdframed}[hidealllines=true, backgroundcolor=black!60, roundcorner=10pt]
+{\color{white} Data associated to this Report is available in the Dryad Digital Repository (\doi{10.5061/dryad.4p080}).}
+\end{mdframed}
 
@@
 The first \emph{Diffantom} and its distorted version are available under the \gls*{cc0} using the Dryad Digital Repository
-  {\color{red} (reference here when published)}.
+  (\doi{10.5061/dryad.4p080}).
 The package is organized following the \gls*{bids} standard.
-The associated software to ``\emph{diffantomize}'' real \gls*{dmri} datasets is available at \url{https://github.com/oesteban/diffantom}
-  under an MIT license.
+The associated software to ``\emph{diffantomize}'' real \gls*{dmri} datasets is available at
+  \url{https://github.com/oesteban/diffantom} under an MIT license.
 \emph{Phantomas} is available in \url{https://github.com/ecaruyer/Phantomas} under the revised-BSD license.

### 2015-10-01 paper/main.tex
update reference

@@
 
+\providecommand{\diffantom}{\emph{Diffantom}}
+\providecommand{\Diffantom}{\emph{Diffantom}}
 \newcommand{\lowb}{\textit{b0}}
@@
   \citep{assaf_composite_2005}.
-\cite{wilkins_fiber_2014} proposed a whole-brain simulated phantom derived from voxel-wise orientation
+\cite{wilkins_fiber_2015} proposed a whole-brain simulated phantom derived from voxel-wise orientation
   of fibers averaged from real \gls*{dmri} scans and the multi-tensor model with a compartment of
@@
 
-In this data report we present \emph{Diffantom}, an \emph{in-silico} dataset to assess tractography and connectivity
+In this data report we present \diffantom{}, an \emph{in-silico} dataset to assess tractography and connectivity
   pipelines using \gls*{dmri} real data as source microstructural information.
-\emph{Diffantom} is inspired by the work of \cite{wilkins_fiber_2014}, with two principal novelties.
+\Diffantom{} is inspired by the work of \cite{wilkins_fiber_2015}, with two principal novelties.
 First, since we use a dataset from the \gls*{hcp} as input, data are already corrected for the most
@@
 We also openly release the \emph{diffantomizer} workflow, the software package necessary to generate custom \emph{diffantoms}.
-\emph{Diffantom} is originally designed for the investigation of susceptibility-derived distortions, a
+\Diffantom{} is originally designed for the investigation of susceptibility-derived distortions, a
   typical artifact that produces geometrical warping in certain regions of \gls*{dmri} datasets.
@@
 
-\noindent\textbf{\textit{Microstructural model\textcolon}\label{sec:data_model}} %
+\noindent\textbf{\textit{Microstructural model\textcolon}\label{sec:diffantom-data-micromodel}} %
 The simulation process relies on a microstructural model derived from real data.
 On one hand, the \emph{diffantomizer} workflow requires up to five fraction maps \tmaps{} of
-  free- and hindered- diffusion (see \autoref{fig:figure1}, panel A).
+  free- and hindered- diffusion (see \autoref{fig:diffantom-01}, panel A).
 These compartments will be derived from the macroscopic structure of tissues within the brain,
@@
 The process to obtain the microstructural model from one dataset of the \gls*{hcp} can be described
-  as follows (see also \autoref{fig:figure1}, panel B):
+  as follows (see also \autoref{fig:diffantom-01}, panel B):
 1) The fiber orientation maps $\{\mathbf{V}_i\}$ and their corresponding estimations of volume fraction $\{F'_i\}$ are
@@
   are combined as described in the \nameref{sec:appendix} to generate the final microstructural model
-  ($\{\mathbf{V}_i\}$, $\{F_i\}$, and $\{T_j\}$), presented in \autoref{fig:figure1}-A.
+  ($\{\mathbf{V}_i\}$, $\{F_i\}$, and $\{T_j\}$), presented in \autoref{fig:diffantom-01}-A.
 
-\noindent\textbf{\textit{Diffusion signal generation\textcolon}\label{sec:data_dwi}} %
+\noindent\textbf{\textit{Diffusion signal generation\textcolon}\label{sec:diffantom-data-dwi}} %
 Once a microstructural model of the subject has been synthesized, the fiber orientation maps $\{\mathbf{V}_i\}$
@@
   the \glspl{fod}.
-A close-up showing how the \glspl{fod} map looks is presented in \autoref{fig:figure1}B.
+A close-up showing how the \glspl{fod} map looks is presented in \autoref{fig:diffantom-01}B.
 The single fiber response is a Gaussian diffusion tensor with axial symmetry and eigenvalues $\lambda_1=$ 2.2\e{-3}
@@
   that users can regenerate similar datasets with different parameters.
-This workflow, presented in \autoref{fig:figure1}, is implemented using
+This workflow, presented in \autoref{fig:diffantom-01}, is implemented using
   \emph{nipype} \citep{gorgolewski_nipype_2011} to ensure reproducibility and usability.
 
-\noindent\textbf{\textit{Interpretation and recommended uses\textcolon}\label{sec:data_use}} %
-To illustrate the features of \emph{Diffantom}, the example dataset underwent a simplified
+\noindent\textbf{\textit{Interpretation and recommended uses\textcolon}\label{sec:diffantom-data-use}} %
+To illustrate the features of \diffantom{}, the example dataset underwent a simplified
   connectivity pipeline including \gls*{csd} and probabilistic tractography from
@@
   seed points evenly distributed across a dilated mask of the \gls*{wm} tissue.
-\autoref{fig:figure2}, panels A1 and A3, show the result of the tractography obtained with such pipeline for
-  the original \emph{Diffantom} and a distorted version.
+\autoref{fig:diffantom-02}, panels A1 and A3, show the result of the tractography obtained with such pipeline for
+  the original \diffantom{} and a distorted version.
 Finally, we applied \emph{tract querier} \citep{wassermann_on_2013} to segment some fiber bundles such
-  as the \gls*{cst} and the forceps minor (see \autoref{fig:figure2}, panels A2, A4).
+  as the \gls*{cst} and the forceps minor (see \autoref{fig:diffantom-02}, panels A2, A4).
 Particularly, due to its location nearby the orbitofrontal lobe, the forceps minor is generally affected by
@@
 
-We recommend \emph{Diffantom} as ground-truth in verification and validation frameworks
-  (\autoref{fig:figure2}, panel B) for testing pipelines.
-\emph{Diffantom} is applicable in the unit testing of algorithms, the integration testing of
+We recommend \diffantom{} as ground-truth in verification and validation frameworks
+  (\autoref{fig:diffantom-02}, panel B) for testing pipelines.
+\Diffantom{} is applicable in the unit testing of algorithms, the integration testing of
   modules in workflows, and the overall system testing.
@@
     model of choice and on the subsequent global tractography outcome.
-  Since the gradient scheme can be set by the user, \emph{Diffantom} can be seen as a mean to translate the so-called
+  Since the gradient scheme can be set by the user, \diffantom{} can be seen as a mean to translate the so-called
   \emph{b-matrix} of the source dataset to any target scheme.
@@
     susceptibility-derived warping, Eddy-currents-derived distortions, etc.) at unit, integration and systems testing levels.
-  \item Using \emph{Diffantom} as in panel B of \autoref{fig:figure2}, it is possible to apply binary classification measures
+  \item Using \diffantom{} as in panel B of \autoref{fig:diffantom-02}, it is possible to apply binary classification measures
     to evaluate the resulting connectivity matrix.
@@
 \end{itemize}
-In order to exemplify one of these intended uses, we also release a \emph{Diffantom} including the susceptibility-derived
+In order to exemplify one of these intended uses, we also release a \diffantom{} including the susceptibility-derived
   distortion in simulation.
@@
   connectomics.
-\emph{Diffantom} is a derivative of \citep{wilkins_fiber_2014} in terms of methodology for
+\Diffantom{} is a derivative of \citep{wilkins_fiber_2015} in terms of methodology for
   simulation with two major advances.
@@
   released within the \gls*{hcp}.
-\cite{wilkins_fiber_2014} explicitly state that their original data were not corrected for certain artifacts,
+\cite{wilkins_fiber_2015} explicitly state that their original data were not corrected for certain artifacts,
   and thus, generated data are affected correspondingly.
-Second, \emph{Diffantom} implements the hindered and restricted compartments model \citep{assaf_composite_2005},
+Second, \diffantom{} implements the hindered and restricted compartments model \citep{assaf_composite_2005},
   which is a more complete model than the multi-tensor diffusion model.
 
-A possible competitor to \emph{Diffantom} is the phantom generated for the Tractography Challenge in
+A possible competitor to \diffantom{} is the phantom generated for the Tractography Challenge in
   ISMRM 2015.
-Similarly to \emph{Diffantom}, the organizers used an \gls*{hcp} subject as source of structural information.
+Similarly to \diffantom{}, the organizers used an \gls*{hcp} subject as source of structural information.
 While this phantom is designed for the bundle-wise evaluation of tractography (with the scores defined in the
   \emph{Tractometer} \citep{cote_tractometer_2013}, such as geometrical coverage, valid connections, invalid connections,
-  missed connections, etc.), \emph{Diffantom} is intended for the connectome-wise evaluation of results,
+  missed connections, etc.), \diffantom{} is intended for the connectome-wise evaluation of results,
   yielding a tractography with a large number of bundles.
-Therefore, \emph{Diffantom} and \emph{FiberFox} are complementary as the hypotheses that can be investigated are different.
-Moreover, \emph{Diffantom} does not require costly manual segmentation of bundles, highly demanding in terms of physiology
+Therefore, \diffantom{} and \emph{FiberFox} are complementary as the hypotheses that can be investigated are different.
+Moreover, \diffantom{} does not require costly manual segmentation of bundles, highly demanding in terms of physiology
   expertise and operation time.
 The software workflow released with this data report (the \emph{diffantomizer}) ensures the reproducibility of
-  \emph{Diffantom} and enables the generation of custom \emph{diffantoms}.
+  \diffantom{} and enables the generation of custom \emph{diffantoms}.
 The \emph{diffantomizer} is designed for, but not limited to, use \gls*{hcp} datasets as source of structural information.
@@
 \noindent\textbf{\textit{Conclusion\textcolon}}\label{sec:conclusion} %
-\emph{Diffantom} is a whole-brain digital phantom generated from a dataset from the \acrlong*{hcp}.
-\emph{Diffantom} is presented here to be openly and freely distributed along with the \emph{diffantomizer} workflow
+\Diffantom{} is a whole-brain digital phantom generated from a dataset from the \acrlong*{hcp}.
+\Diffantom{} is presented here to be openly and freely distributed along with the \emph{diffantomizer} workflow
   to generate new \emph{diffantoms}.
@@
 \section*{Data Sharing}
-The first \emph{Diffantom} and its distorted version are available under the \gls*{cc0} using the Dryad Digital Repository
+The first \diffantom{} and its distorted version are available under the \gls*{cc0} using the Dryad Digital Repository
   (\doi{10.5061/dryad.4p080}).
@@
 \section*{Appendix}\label{sec:appendix}
-Let $\{T'_j\}$ be the set of original fractions maps obtained with \texttt{act\_anat\_prepare\_fsl}, a
+Let $\{T'_j\}$ be the set of original fractions maps obtained with \path{act_anat_prepare_fsl}, a
   tool in \emph{MRTrix} that combines \gls*{fast} and \gls*{first}
@@
 \end{center}
-\textbf{\refstepcounter{figure}\label{fig:figure1} Figure \arabic{figure}. }{%
-\textbf{A. Microstructural model of \emph{Diffantom}}.
+\textbf{\refstepcounter{figure}\label{fig:diffantom-01} Figure \arabic{figure}. }{%
+\textbf{A. Microstructural model of \diffantom{}}.
 The phantom is simulated from an underlying microstructural model specified with the
@@
 The pipeline to generate phantoms from any \gls*{hcp} dataset is presented in the lower panel.
-Once the microstructural model shown in the upper panel has been prepared as described in \nameref{sec:data_model},
+Once the microstructural model shown in the upper panel has been prepared as described in \nameref{sec:diffantom-data-micromodel},
   the local orientations are computed and fed into \emph{Phantomas} to finally simulate the signal.
@@
 \end{center}
-\textbf{\refstepcounter{figure}\label{fig:figure2} Figure \arabic{figure}. }{%
+\textbf{\refstepcounter{figure}\label{fig:diffantom-02} Figure \arabic{figure}. }{%
 \textbf{A. Example dataset.}
-A1 and A3 show the tractogram of fibers crossing slice 56 of \emph{Diffantom} as
+A1 and A3 show the tractogram of fibers crossing slice 56 of \diffantom{} as
   extracted with \emph{MRTrix}, represented over the corresponding slice of the
@@
   with transparency.
-In the distorted \emph{Diffantom} (A4) the forceps minor was not detected.
-\textbf{B. Recommended use of \emph{Diffantom}}.
+In the distorted \diffantom{} (A4) the forceps minor was not detected.
+\textbf{B. Recommended use of \diffantom{}}.
 The phantom is designed to be used as ground-truth information in evaluation frameworks,

### 2015-12-22 paper/abbreviations.tex
final version for bioRxiv
reply reviewers' comments

--- 2015-08-14 paper/abbreviations.tex
@@
 % @Last Modified by:   Oscar Esteban
-% @Last Modified time: 2015-08-14 09:16:55
+% @Last Modified time: 2015-12-16 16:39:23
 
@@
 
+\newglossaryentry{bids}%
+{
+	name={BIDS},
+	first={BIDS standard \cite[Brain Imaging Data Structure,][]{gorgolewski_brain_2015}},
+	long={Brain Imaging Data Structure},
+	description={Brain Imaging Data Structure}
+}
+
+\newglossaryentry{bedpostx}%
+{
+	name={BEDPOSTX},
+	first={BEDPOSTX \cite[Bayesian Estimation of Diffusion Parameters Obtained using Sampling Techniques modelling crossing --X-- fibres,][]{jbabdi_modelbased_2012}},
+	description={Bayesian Estimation of Diffusion Parameters Obtained using Sampling Techniques modelling crossing --X-- fibres}
+}
+\newacronym[first={FAST \cite[FMRIB's Automated Segmentation Tool,][]{zhang_segmentation_2001}}]%
+{fast}{FAST}{FMRIB's Automated Segmentation Tool}
+\newacronym[first={FIRST \cite[FMRIB's Integrated Registration and Segmentation Tool,][]{patenaude_bayesian_2011}}]%
+{first}{FIRST}{FMRIB's Integrated Registration and Segmentation Tool}
+
 \makeglossaries

### 2015-12-22 paper/main.tex
reply reviewers' comments

@@
 \documentclass[english]{frontiers/frontiersSCNS} % for Science, Engineering and Humanities and Social Sciences articles
+\usepackage{xifthen}
+\newboolean{review}
+\setboolean{review}{true}
+
 \usepackage[mode=buildnew]{standalone}
@@
 %\setcitestyle{square}
-\usepackage{url,lineno,microtype}
+\usepackage{url,microtype}
+\usepackage[right]{lineno}
 \usepackage[toc,nomain,acronym,shortcuts,translate=false]{glossaries}
@@
 
+\ifthenelse{\boolean{review}}
+{
+  \usepackage[colorinlistoftodos, textwidth=1.4cm]{todonotes}
+  \setlength{\marginparwidth}{1.6cm}
+  \reversemarginpar
+  \newcommand{\revcomment}[2][]{\ifthenelse{\equal{#1}{}}{}{\todo[size=\tiny]{#1}}%
+  \begingroup \color{blue!70} #2\endgroup}
+}
+{
+  \newcommand{\todo}[2][]{}
+  \newcommand{\revcomment}[2][]{#2}
+}
+
+% \usepackage{showframe}
 \def\keyFont{\fontsize{7}{9}\helveticabold }
@@
 
-\newcommand*\glslines[2][]{\glslink[#1]{#2}{\glsdisp{#2}{\acrlong{#2} --\glsentryshort{#2}--}}}
-
 \begin{document}
@@
 \maketitle
+\linenumbers
 
-\begin{mdframed}[hidealllines=true, backgroundcolor=black!60, roundcorner=10pt]
-{\color{white} Data associated to this Report is available in the Dryad Digital Repository (\doi{10.5061/dryad.4p080}).}
-\end{mdframed}
+% \begin{mdframed}[hidealllines=true, backgroundcolor=black!5, roundcorner=5pt]
+\section*{Diffantom in brief}
+\revcomment[Rev.\#2\\Q.10]{%
+\diffantom{} is a whole-brain \gls*{dmri} phantom publicly available
+  through the Dryad Digital Repository (\doi{10.5061/dryad.4p080}).
+The dataset contains two single-shell \gls*{dmri} images, along with the corresponding gradient
+  information, packed following the \gls*{bids}.
+The released dataset is designed for the evaluation of the impact of susceptibility distortions
+  and benchmarking existing correction methods.
+}%
 
-\linenumbers
+\revcomment[Rev.\#1\\Q14.1]{%
+In this Data Report we also release the software instruments involved in generating \emph{diffantoms},
+  so that researchers are able to generate new phantoms derived from different subjects,
+  and apply these data in other applications like investigating diffusion sampling
+  schemes, the assessment of \gls*{dmri} processing methods, the simulation of pathologies
+  and imaging artifacts, etc.
+In summary, \diffantom{} is intended for unit testing of novel methods, cross-comparison of established methods,
+  and integration testing of partial or complete processing flows to extract connectivity networks
+  from \gls{dmri}.}
 
@@
   \item Simulation of pathological brains by altering the microstructural
-    model accordingly (e.g. as tumors were simulated in \cite{kaus_simulation_2000}).
+    model accordingly \cite[e.g. as tumors were simulated in][]{kaus_simulation_2000}.
 \end{itemize}
@@
   distortion in simulation.
+\revcomment[Rev.\#1\\Q14.2]{%
+These two images belong to a broader dataset, automatically generated, used in a study to quantify
+  the impact of susceptibility distortions and correction methods on the connectome extraction
+  \cite[Chapter 5]{esteban_image_2015}.
+In this study, three widely-used correction methods are compared in a reference framework of
+  several \emph{Diffantoms} with realistic and controlled distortions.
+This context provides a useful resource to characterize the impact of susceptibility distortion
+  on the final connectivity network and allows the evaluation of the different correction
+  methodologies available.%
+}
 

### 2016-09-25 diffantom/__init__.py
[ENH] Add setup.py

@@ -5,4 +5,5 @@
 # @Last Modified by:   Oscar Esteban
 # @Last Modified time: 2015-06-23 13:05:13
 
-from interfaces import PhantomasSticksSim, LoadSamplingScheme
+from diffantom.interfaces import PhantomasSticksSim, LoadSamplingScheme
+from diffantom.workflows import gen_diffantom, gen_model, finf_bundles

### 2016-09-25 diffantom/info.py

@@ -1,0 +1,80 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+# emacs: -*- mode: python; py-indent-offset: 4; indent-tabs-mode: nil -*-
+# vi: set ft=python sts=4 ts=4 sw=4 et:
+"""
+Base module variables
+"""
+from __future__ import absolute_import, division, print_function, unicode_literals
+from datetime import datetime
+
+__packagename__ = 'diffantom'
+__version__ = '1.0.0'
+__author__ = 'Oscar Esteban'
+__affiliation__ = 'Psychology Department, Stanford University'
+__credits__ = ['Oscar Esteban']
+__license__ = '3-clause BSD'
+__maintainer__ = 'Oscar Esteban'
+__email__ = 'code@oscaresteban.es'
+__status__ = 'Prototype'
+__copyright__ = 'Copyright {}, {}'.format(datetime.now().year, __author__)
+
+__description__ = """\
+Diffantom: Whole-Brain Diffusion MRI Phantoms Derived from Real Datasets of the \
+Human Connectome Project"""
+__longdesc__ = """\
+Diffantom is a whole-brain diffusion MRI (dMRI) phantom publicly available through the \
+Dryad Digital Repository (doi:10.5061/dryad.4p080). The dataset contains two single-shell \
+dMRI images, along with the corresponding gradient information, packed following the BIDS \
+standard (Brain Imaging Data Structure, Gorgolewski et al., 2015). \
+The released dataset is designed for the evaluation of the impact of susceptibility \
+distortions and benchmarking existing correction methods.\
+
+This project contains the software instruments involved in generating \
+diffantoms, so that researchers are able to generate new phantoms derived \
+from different subjects, and apply these data in other applications like \
+investigating diffusion sampling schemes, the assessment of dMRI processing methods, \
+the simulation of pathologies and imaging artifacts, etc. In summary, Diffantom is \
+intended for unit testing of novel methods, cross-comparison of established methods, \
+and integration testing of partial or complete processing flows to extract connectivity \
+networks from dMRI.
+"""
+
+CLASSIFIERS = [
+    'Development Status :: 4 - Beta',
+    'Intended Audience :: Science/Research',
[…]

### 2016-09-25 diffantom/interfaces.py

@@ -19,9 +19,10 @@ from nipype.interfaces.base import (TraitedSpec, File, InputMultiPath,
                                     BaseInterface, BaseInterfaceInputSpec,
                                     traits)
 from nipype.utils.filemanip import split_filename, fname_presuffix
-from utils import sigmoid_filter
-
 from nipype import logging
+
+from diffantom.utils import sigmoid_filter
+
 iflogger = logging.getLogger('interface')
 
 

### 2016-09-25 diffantom/scripts/diffantom.py

@@ -10,30 +10,17 @@
 The Diffantom software layer
 ============================
 """
-import os
-import os.path as op
-import sys
 
 
-__author__ = "Oscar Esteban"
-__copyright__ = "Copyright 2015, Biomedical Image Technologies (BIT), \
-                 Universidad Politécnica de Madrid"
-__credits__ = "Oscar Esteban"
-__license__ = "MIT"
-__version__ = "0.1"
-__maintainer__ = "Oscar Esteban"
-__email__ = "code@oscaresteban.es"
-__status__ = "Prototype"
-
-
-if __name__ == '__main__':
+def main():
     from argparse import ArgumentParser
     from argparse import RawTextHelpFormatter
-    from workflows import gen_diffantom, gen_model, finf_bundles
     import glob
     import os
     import os.path as op
+    import sys
     from shutil import copyfileobj
+    from diffantom.workflows import gen_diffantom, gen_model, finf_bundles
 
     parser = ArgumentParser(description='Preprocessing dMRI routine',
                             formatter_class=RawTextHelpFormatter)
@@ -93,3 +80,6 @@ if __name__ == '__main__':
     wf.base_dir = settings['work_dir']
     # wf.write_graph(format='pdf')
     wf.run(plugin=plugin, plugin_args=plugin_args)
+
+if __name__ == '__main__':
+    main()

### 2016-09-25 diffantom/workflows.py

@@ -114,7 +114,7 @@ def finf_bundles(name='FINFBundles', settings={}):
 
 def simulate(name='SimDWI'):
     from nipype.interfaces.dipy import Denoise
-    from interfaces import PhantomasSticksSim, LoadSamplingScheme
+    from diffantom.interfaces import PhantomasSticksSim, LoadSamplingScheme
 
     in_fields = ['fibers', 'fractions', 'in_5tt', 'in_mask', 'scheme']
 
@@ -149,8 +149,8 @@ def simulate(name='SimDWI'):
 
 
 def preprocess_model(name='PrepareModel'):
-    from interfaces import SigmoidFilter
-    import utils as pu
+    from diffantom.interfaces import SigmoidFilter
+    import diffantom.utils as pu
 
     in_fields = ['t1w', 'fibers', 'fractions', 'parcellation', 'in_fa']
     sgm_structures = ['L_Accu', 'R_Accu', 'L_Caud', 'R_Caud',
@@ -282,7 +282,7 @@ def preprocess_model(name='PrepareModel'):
 
 def act_workflow(name='Tractography'):
     from nipype.interfaces.mrtrix import MRTrix2TrackVis as TCK2TRK
-    import utils as pu
+    import diffantom.utils as pu
 
     inputnode = pe.Node(niu.IdentityInterface(
         fields=['in_dwi', 'in_scheme', 'in_5tt', 'parcellation', 'aparc',
@@ -424,7 +424,7 @@ def act_workflow(name='Tractography'):
 
 def track_bundle(name='BundleTrack', incl_rois=True):
     from nipype.interfaces.mrtrix import MRTrix2TrackVis as TCK2TRK
-    import utils as pu
+    import diffantom.utils as pu
 
     infields = ['in_fod', 'wmparc', 'seed_lbs', 'parcellation',
                 'in_5tt', 'in_scheme', 'roi_excl']
@@ -477,8 +477,8 @@ def track_bundle(name='BundleTrack', incl_rois=True):
 
 
 def track_querier(name='TractQuery'):
-    from interfaces import TractQuerier
-    import utils as pu
+    from diffantom.interfaces import TractQuerier
+    import diffantom.utils as pu
 
     inputnode = pe.Node(niu.IdentityInterface(
         fields=['in_track', 'in_parc', 'in_qry']), name='inputnode')

### 2016-09-25 setup.py

@@ -1,0 +1,65 @@
+`#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+# emacs: -*- mode: python; py-indent-offset: 4; indent-tabs-mode: nil -*-
+# vi: set ft=python sts=4 ts=4 sw=4 et:
+#
+# @Author: oesteban - code@oscaresteban.es
+# @Date:   2014-06-05 16:51:52
+# @Last Modified by:   oesteban
+# @Last Modified time: 2016-09-06 09:35:45
+""" diffantom setup script """
+
+PACKAGE_NAME = 'diffantom'
+
+def main():
+    """ Install entry-point """
+    from os import path as op
+    from glob import glob
+    from inspect import getfile, currentframe
+    from setuptools import setup, find_packages
+    from io import open  # pylint: disable=W0622
+
+    this_path = op.dirname(op.abspath(getfile(currentframe())))
+
+    # Python 3: use a locals dictionary
+    # http://stackoverflow.com/a/1463370/6820620
+    ldict = locals()
+
+    # Get version and release info, which is all stored in diffantom/info.py
+    module_file = op.join(this_path, PACKAGE_NAME, 'info.py')
+    with open(module_file) as infofile:
+        pythoncode = [line for line in infofile.readlines() if not line.strip().startswith('#')]
+        exec('\n'.join(pythoncode), globals(), ldict)
+
+    setup(
+        name=PACKAGE_NAME,
+        version=ldict['__version__'],
+        description=ldict['__description__'],
+        long_description=ldict['__longdesc__'],
+        author=ldict['__author__'],
+        author_email=ldict['__email__'],
+        maintainer=ldict['__maintainer__'],
+        maintainer_email=ldict['__email__'],
+        license=ldict['__license__'],
+        url=ldict['URL'],
+        download_url=ldict['DOWNLOAD_URL'],
+        classifiers=ldict['CLASSIFIERS'],
+        packages=find_packages(),
+        zip_safe=False,
+        # Dependencies handling
+        setup_requires=[],
+        install_requires=ldict['REQUIRES'],
+        dependency_links=ldict['LINKS_REQUIRES'],
+        tests_require=ldict['TESTS_REQUIRES'],
+        extras_require=ldict['EXTRA_REQUIRES'],
+        package_data={PACKAGE_NAME: ['data/*.txt',]},
+        entry_points={
+            'console_scripts': [
[…]

### 2016-09-25 setup.py
fix typo

@@ -1,4 +1,4 @@
-`#!/usr/bin/env python
+#!/usr/bin/env python
 # -*- coding: utf-8 -*-
 # emacs: -*- mode: python; py-indent-offset: 4; indent-tabs-mode: nil -*-
 # vi: set ft=python sts=4 ts=4 sw=4 et:

### 2016-09-25 setup.py
fix package_data

@@ -52,7 +52,7 @@ def main():
         dependency_links=ldict['LINKS_REQUIRES'],
         tests_require=ldict['TESTS_REQUIRES'],
         extras_require=ldict['EXTRA_REQUIRES'],
-        package_data={PACKAGE_NAME: ['data/*.txt',]},
+        package_data={PACKAGE_NAME: ['data/bedpostx.json',]},
         entry_points={
             'console_scripts': [
                 'diffantom=diffantom.scripts.diffantom:main'

### 2016-09-25 diffantom/__init__.py
fix diffantom script

@@ -4,6 +4,6 @@
 # @Date:   2015-06-15 14:50:01
 # @Last Modified by:   Oscar Esteban
 # @Last Modified time: 2015-06-23 13:05:13
-
-from diffantom.interfaces import PhantomasSticksSim, LoadSamplingScheme
-from diffantom.workflows import gen_diffantom, gen_model, finf_bundles
+from __future__ import absolute_import
+from .interfaces import PhantomasSticksSim, LoadSamplingScheme
+from .workflows import gen_diffantom, gen_model, finf_bundles

### 2016-09-25 diffantom/interfaces.py

@@ -4,7 +4,7 @@
 # @Date:   2015-06-23 12:29:58
 # @Last Modified by:   oesteban
 # @Last Modified time: 2015-07-13 11:47:13
-
+from __future__ import absolute_import
 import os
 import os.path as op
 import glob
@@ -21,7 +21,7 @@ from nipype.interfaces.base import (TraitedSpec, File, InputMultiPath,
 from nipype.utils.filemanip import split_filename, fname_presuffix
 from nipype import logging
 
-from diffantom.utils import sigmoid_filter
+from .utils import sigmoid_filter
 
 iflogger = logging.getLogger('interface')
 

### 2016-09-25 diffantom/workflows.py

@@ -4,7 +4,7 @@
 # @Date:   2015-06-23 12:32:07
 # @Last Modified by:   Oscar Esteban
 # @Last Modified time: 2015-09-09 16:21:22
-
+from __future__ import absolute_import
 import os
 import os.path as op
 import numpy as np
@@ -114,7 +114,7 @@ def finf_bundles(name='FINFBundles', settings={}):
 
 def simulate(name='SimDWI'):
     from nipype.interfaces.dipy import Denoise
-    from diffantom.interfaces import PhantomasSticksSim, LoadSamplingScheme
+    from .interfaces import PhantomasSticksSim, LoadSamplingScheme
 
     in_fields = ['fibers', 'fractions', 'in_5tt', 'in_mask', 'scheme']
 
@@ -149,7 +149,7 @@ def simulate(name='SimDWI'):
 
 
 def preprocess_model(name='PrepareModel'):
-    from diffantom.interfaces import SigmoidFilter
+    from .interfaces import SigmoidFilter
     import diffantom.utils as pu
 
     in_fields = ['t1w', 'fibers', 'fractions', 'parcellation', 'in_fa']
@@ -477,7 +477,7 @@ def track_bundle(name='BundleTrack', incl_rois=True):
 
 
 def track_querier(name='TractQuery'):
-    from diffantom.interfaces import TractQuerier
+    from .interfaces import TractQuerier
     import diffantom.utils as pu
 
     inputnode = pe.Node(niu.IdentityInterface(

### 2016-09-25 setup.py

@@ -55,7 +55,7 @@ def main():
         package_data={PACKAGE_NAME: ['data/bedpostx.json',]},
         entry_points={
             'console_scripts': [
-                'diffantom=diffantom.scripts.diffantom:main'
+                'diffantom=diffantom.diffantom:main'
 
             ]
         }

### 2016-09-25 diffantom/diffantom.py
use relative-explicit imports

@@ -4,13 +4,12 @@
 # @Date:   2015-06-25 15:46:08
 # @Last Modified by:   Oscar Esteban
 # @Last Modified time: 2015-09-09 15:50:20
-
 """
 ============================
 The Diffantom software layer
 ============================
 """
-
+from __future__ import absolute_import
 
 def main():
     from argparse import ArgumentParser
@@ -20,7 +19,7 @@ def main():
     import os.path as op
     import sys
     from shutil import copyfileobj
-    from diffantom.workflows import gen_diffantom, gen_model, finf_bundles
+    from .workflows import gen_diffantom, gen_model, finf_bundles
 
     parser = ArgumentParser(description='Preprocessing dMRI routine',
                             formatter_class=RawTextHelpFormatter)

### 2016-09-25 diffantom/__init__.py
add find_packages exceptions, fix main __init__

@@ -5,5 +5,18 @@
 # @Last Modified by:   Oscar Esteban
 # @Last Modified time: 2015-06-23 13:05:13
 from __future__ import absolute_import
+from .info import (
+    __version__,
+    __author__,
+    __email__,
+    __maintainer__,
+    __copyright__,
+    __credits__,
+    __license__,
+    __status__,
+    __description__,
+    __longdesc__
+)
+
 from .interfaces import PhantomasSticksSim, LoadSamplingScheme
 from .workflows import gen_diffantom, gen_model, finf_bundles

### 2016-09-25 setup.py

@@ -44,7 +44,7 @@ def main():
         url=ldict['URL'],
         download_url=ldict['DOWNLOAD_URL'],
         classifiers=ldict['CLASSIFIERS'],
-        packages=find_packages(),
+        packages=find_packages(exclude=['build', 'docs', 'paper']),
         zip_safe=False,
         # Dependencies handling
         setup_requires=[],

### 2016-09-25 setup.py
add requirements.txt, fix setup.py, uploaded to pypi

@@ -44,7 +44,7 @@ def main():
         url=ldict['URL'],
         download_url=ldict['DOWNLOAD_URL'],
         classifiers=ldict['CLASSIFIERS'],
-        packages=find_packages(exclude=['build', 'docs', 'paper']),
+        packages=find_packages(exclude=['build', 'dist', 'docs', 'paper']),
         zip_safe=False,
         # Dependencies handling
         setup_requires=[],

### 2016-09-25 paper/appendix.tex

--- 2015-10-01 paper/main.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% This is just an example/guide for you to refer to when submitting manuscripts to Frontiers, it is not mandatory to use Frontiers .cls files nor frontiers.tex  %
-% This will only generate the Manuscript, the final article will be typeset by Frontiers after acceptance.                                                 %
-%                                                                                                                                                         %
-% When submitting your files, remember to upload this *tex file, the pdf generated with it, the *bib file (if bibliography is not within the *tex) and all the figures.
+% This is just an example/guide for you to refer to when producing your supplementary material for your Frontiers article.                                 %
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%%% Version 3.1 Generated 2015/22/05 %%%
+%%% Version 2.1 Generated 2015/05/22 %%%
 %%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
@@
 
-\documentclass[english]{frontiers/frontiersSCNS} % for Science, Engineering and Humanities and Social Sciences articles
-\usepackage[mode=buildnew]{standalone}
-\usepackage{tikz}
-\usepackage[framemethod=TikZ]{mdframed}
-%\setcitestyle{square}
-\usepackage{url,lineno,microtype}
+\documentclass{frontiers/frontiers_suppmat} % for all articles
 \usepackage[toc,nomain,acronym,shortcuts,translate=false]{glossaries}
+\usepackage{url,microtype}
 \usepackage[colorlinks=true,linkcolor=black, citecolor=black!80, urlcolor=black!80]{hyperref}
@@
 
-\def\keyFont{\fontsize{7}{9}\helveticabold }
+% Leave a blank line between paragraphs in stead of using \\
+
+\def\keyFont{\fontsize{8}{11}\helveticabold }
+
 \def\firstAuthorLast{Esteban {et~al.}} %use et al only if is more than 1 author
-\def\Authors{Oscar Esteban\,$^{1,2,*}$, Emmanuel Caruyer\,$^{3}$, Alessandro Daducci\,$^{4}$, Meritxell Bach-Cuadra\,$^{4,5}$,%
-Mar\'ia-J. Ledesma-Carbayo\,$^{1,2}$ and Andres Santos\,$^{1,2}$}
+\def\Authors{Oscar Esteban*, Emmanuel Caruyer, Alessandro Daducci, Meritxell Bach-Cuadra,%
+Mar\'ia-J. Ledesma-Carbayo and Andres Santos}
 
-\def\Address{%
-$^{1}$Biomedical Image Technologies (BIT), ETSI Telecom., Universidad Polit\'ecnica de Madrid, Madrid, Spain \\
-$^{2}$Centro de Investigaci\'on Biom\'edica en Red en Bioingenier\'ia, Biomateriales y Nanomedicina (CIBER-BBN), Spain \\
-$^{3}$CNRS, IRISA (UMR 6074), VisAGeS research group, Rennes, France \\
-$^{4}$Signal Processing Laboratory (LTS5), \'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL), Lausanne, Switzerland \\
-${^5}$Dept. of Radiology, CIBM, University Hospital Center (CHUV) and University of Lausanne (UNIL), Lausanne, Switzerland %
-}
 % The Corresponding Author should be marked with an asterisk
@@
 
-\newcommand*\glslines[2][]{\glslink[#1]{#2}{\glsdisp{#2}{\acrlong{#2} --\glsentryshort{#2}--}}}
-
 \begin{document}
-
 \onecolumn
@@
 
-\title[Diffantom]{Diffantom: whole-brain diffusion MRI phantoms derived from real datasets of the \acrlong{hcp}}
+\title[Supplementary Material]{\helvetica{Supplementary Material to: Diffantom}} %Please insert the title of your article here
 
 \author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
-\address{} %This field will be automatically populated
 \correspondance{} %This field will be automatically populated
@@
 \extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
-%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}
+%\extraAuth{Corresponding Author2: email2@uni2.edu}
 
@@
 
-\begin{mdframed}[hidealllines=true, backgroundcolor=black!60, roundcorner=10pt]
-{\color{white} Data associated to this Report is available in the Dryad Digital Repository (\doi{10.5061/dryad.4p080}).}
-\end{mdframed}
-
-\linenumbers
-
-\section*{Introduction}
-Fiber tracking on \gls*{dmri} data has become an important tool for the \textit{in-vivo} investigation
-  of the structural configuration of fiber bundles at the macroscale.
-Tractography is fundamental to gain information about \gls*{wm} morphology in many clinical applications
-  like neurosurgical planning \citep{golby_interactive_2011}, post-surgery evaluations \citep{toda_utility_2014},
-  and the study of neurological diseases as in \citep{chua_diffusion_2008} addressing multiple sclerosis and
-  Alzheimer's disease.
-The analysis of structural brain networks using graph theory is also applied on tractography,
-  for instance in the definition of the unique subject-wise patterns of connectivity
-  \citep{sporns_human_2005}, in the assessment of neurological diseases \citep{griffa_structural_2013}, and in the
-  study of the link between structural and functional connectivity \citep{messe_predicting_2015}.
-However, the development of the field is limited by the lack of a gold standard to test and compare the
-  wide range of methodologies available for processing and analyzing \gls*{dmri}.
-
-Large efforts have been devoted to the development of physical phantoms
-  \citep{lin_validation_2001,campbell_flowbased_2005,perrin_validation_2005,fieremans_simulation_2008,tournier_resolving_2008}.
-\cite{cote_tractometer_2013} conducted a thorough review of tractography methodologies using the
-  so-called \emph{FiberCup} phantom \citep{poupon_new_2008,fillard_quantitative_2011}.
-These phantoms are appropriate to evaluate the angular resolution in fiber crossings and accuracy of
-  direction-independent scalar parameters in very simplistic geometries.
-Digital simulations are increasingly popular because the complexity of whole-brain tractography
-  can not be accounted for with current materials and proposed methodologies to build physical phantoms.
-Early digital phantoms started with simulation of simple geometries
-  \citep{basser_in_2000,goessl_fiber_2002,tournier_limitations_2002,leemans_mathematical_2005}
-  to evaluate the angular resolution as well.
-These tools generally implemented the multi-tensor model \citep{alexander_analysis_2001,tuch_high_2002}
-  to simulate fiber crossing, fanning, kissing, etc.
-\cite{close_software_2009} presented the \emph{Numerical Fibre Generator}, a software to simulate
-  spherical shapes filled with digital fiber tracts.
-\cite{caruyer_phantomas_2014} proposed \emph{Phantomas} to simulate any kind of analytic geometry
-  inside a sphere.
-\emph{Phantomas} models diffusion by a restricted and a hindered compartment, similar to
-  \citep{assaf_composite_2005}.
-\cite{wilkins_fiber_2015} proposed a whole-brain simulated phantom derived from voxel-wise orientation
-  of fibers averaged from real \gls*{dmri} scans and the multi-tensor model with a compartment of
-  isotropic diffusion.
-\cite{neher_fiberfox_2014} proposed \emph{FiberFox}, a visualization software to develop
-  complex geometries and their analytical description.
-Once the geometries are obtained, the software generates the corresponding \gls*{dmri} signal with a
-  methodology very close to that implemented in \emph{Phantomas}.
-An interesting outcome of \emph{FiberFox} is the phantom dataset\footnote{Available at
-  \url{http://www.tractometer.org/ismrm_2015_challenge/}} created for the Tractography
-  Challenge held in ISMRM 2015.
-This dataset was derived from the tractography extracted in one \gls*{hcp} dataset.
-In the tractogram, 25 fiber bundles of interest were manually segmented by experts.
-Using \emph{FiberFox}, the segmentation of each bundle was mapped to an analytical
-  description, and finally simulated the signal.
-
-In this data report we present \diffantom{}, an \emph{in-silico} dataset to assess tractography and connectivity
-  pipelines using \gls*{dmri} real data as source microstructural information.
-\Diffantom{} is inspired by the work of \cite{wilkins_fiber_2015}, with two principal novelties.
-First, since we use a dataset from the \gls*{hcp} as input, data are already corrected for the most
-  relevant distortions.
-The second improvement is a more advanced signal model to generate the phantom using the hindered and restricted
-  diffusion model of \emph{Phantomas} \citep{caruyer_phantomas_2014}.
-As a result, we provide a whole-brain digital phantom of \gls*{dmri} data with structural information derived
-  from an \gls*{hcp} dataset.
-We also openly release the \emph{diffantomizer} workflow, the software package necessary to generate custom \emph{diffantoms}.
-\Diffantom{} is originally designed for the investigation of susceptibility-derived distortions, a
-  typical artifact that produces geometrical warping in certain regions of \gls*{dmri} datasets.
-In \citep{esteban_simulationbased_2014} we addressed this phenomenon and concluded that the connectivity
-  matrix of \emph{Phantomas} was not dense enough to evaluate the integration of correction methods
-  in pipelines for the connectome extraction.
-
-\section*{Data description}
-
-\noindent\textbf{\textit{Microstructural model\textcolon}\label{sec:diffantom-data-micromodel}} %
-The simulation process relies on a microstructural model derived from real data.
-On one hand, the \emph{diffantomizer} workflow requires up to five fraction maps \tmaps{} of
-  free- and hindered- diffusion (see \autoref{fig:diffantom-01}, panel A).
-These compartments will be derived from the macroscopic structure of tissues within the brain,
-  specified in the following order\footnote{Corresponding to the \emph{5TT format} established
-  with the latest version 3.0 of \emph{MRTrix} \citep{tournier_mrtrix_2012}}:
-  \gls*{cgm}, \gls*{dgm}, \gls*{wm}, \gls*{csf}, and abnormal tissue\footnote{Since here we
-  simulate healthy subjects, the last fraction map $T_5$ is empty and can be omitted}.
-On the other hand, the restricted-diffusion compartments are specified by up to three volume fractions \fmaps{}
-  of three single fiber populations per voxel along with their corresponding direction maps \vmaps{}.
-
-The process to obtain the microstructural model from one dataset of the \gls*{hcp} can be described
-  as follows (see also \autoref{fig:diffantom-01}, panel B):
-1) The fiber orientation maps $\{\mathbf{V}_i\}$ and their corresponding estimations of volume fraction $\{F'_i\}$ are
-  obtained using the ball-and-stick model for multi-shell data of \gls*{bedpostx}
-  on the \gls*{dmri} data.
-The \gls*{hcp} recommends \gls*{bedpostx} to reconstruct their data \citep{glasser_minimal_2013}.
-A further advantage is that \gls*{bedpostx} exploits the multi-shell acquisitions of the \gls*{hcp} while
-  operating at whole-brain level.
-2) A \gls*{fa} map is obtained after fitting a tensor model with \emph{MRTrix}.
-As we shall see in the Appendix, the \gls*{fa} is used to infer $F_1$ (the fraction map of the most prevalent fiber),
-  avoiding the extremely noisy estimation of $F'_1$ performed by \gls*{bedpostx} in the previous step.
-3) The original fiber fractions $\{F'_i\}$ and the \gls*{fa} map are denoised with a nonlocal means filter included
-  in \emph{dipy} \citep{garyfallidis_dipy_2014}.
-This step produces an important smoothing of the maps, while preserving the edges.
-Smoothing is also beneficial in simplifying the voxel-wise diffusion model.
-4) The macrostructural fractions $\{T'_j\}$ are extracted from the \acrlong*{t1} image of the dataset,
-  using standard \emph{FSL} segmentation tools \citep{jenkinson_fsl_2012}.
-5) The images obtained previously (\gls*{fa} map, $\{\mathbf{V}_i\}$, $\{F'_i\}$, and $\{T'_j\}$)
-  are combined as described in the \nameref{sec:appendix} to generate the final microstructural model
-  ($\{\mathbf{V}_i\}$, $\{F_i\}$, and $\{T_j\}$), presented in \autoref{fig:diffantom-01}-A.
-
-\noindent\textbf{\textit{Diffusion signal generation\textcolon}\label{sec:diffantom-data-dwi}} %
-Once a microstructural model of the subject has been synthesized, the fiber orientation maps $\{\mathbf{V}_i\}$
-  are weighted by the fiber-fraction maps $\{F_i\}$ and projected onto a continuous representation of
-  the \glspl{fod}.
-A close-up showing how the \glspl{fod} map looks is presented in \autoref{fig:diffantom-01}B.
-The single fiber response is a Gaussian diffusion tensor with axial symmetry and eigenvalues $\lambda_1=$ 2.2\e{-3}
-  \diffunits{} and $\lambda_{2,3}=$ 0.2\e{-3} \diffunits{}.
-The resulting \glspl{fod} map is then combined with the free- and hindered-diffusion compartments corresponding to $\{T_j\}$.
-The free-diffusion compartment corresponds to the \gls*{csf} fraction map $T_4$ and is modeled with isotropic
-  diffusivity $D_{CSF}$ of 3.0\e{-3} \diffunits{}.
-The hindered-diffusion compartments correspond to $\{T_1,T_2,T_3\}$ and are also modeled with isotropic diffusivity
-  $D_{WM} =$ 2.0\e{-4}, $D_{cGM} =$ 7.0\e{-4} and $D_{dGM} =$ 9.0\e{-4}, respectively [\diffunits{}].
-All these values for diffusivity (and the corresponding to the single-fiber response) can be modified by the user with
-  custom settings.
-The restricted- and hindered- compartments are then fed into \emph{Phantomas} \citep{caruyer_phantomas_2014}
-  and the final \gls*{dmri} signal is obtained.
-By default, diffusion data are generated using a scheme of 100 directions distributed in one shell with uniform
-  coverage \citep{caruyer_design_2013}.
-Custom one- or multi-shell schemes can be generated supplying the tables of corresponding vectors and $b$-values.
-Rician noise is also included in \emph{Phantomas}, and the \gls*{snr} can be set by the user.
-The default value for \gls*{snr} is preset to 30.0.
-
-\noindent\textbf{\textit{Implementation and reproducibility\textcolon}\label{sec:data_workflow}} %
-We also provide the \emph{diffantomizer} workflow, the software package used to generate \emph{diffantoms}, so
-  that users can regenerate similar datasets with different parameters.
-This workflow, presented in \autoref{fig:diffantom-01}, is implemented using
-  \emph{nipype} \citep{gorgolewski_nipype_2011} to ensure reproducibility and usability.
-
-\noindent\textbf{\textit{Interpretation and recommended uses\textcolon}\label{sec:diffantom-data-use}} %
-To illustrate the features of \diffantom{}, the example dataset underwent a simplified
-  connectivity pipeline including \gls*{csd} and probabilistic tractography from
-  \emph{MRTrix} \citep{tournier_mrtrix_2012}.
-\Gls*{csd} was reconstructed using 8$^\text{th}$-order spherical harmonics, and tractography with 1.6\e{6}
-  seed points evenly distributed across a dilated mask of the \gls*{wm} tissue.
-\autoref{fig:diffantom-02}, panels A1 and A3, show the result of the tractography obtained with such pipeline for
-  the original \diffantom{} and a distorted version.
-Finally, we applied \emph{tract querier} \citep{wassermann_on_2013} to segment some fiber bundles such
-  as the \gls*{cst} and the forceps minor (see \autoref{fig:diffantom-02}, panels A2, A4).
-Particularly, due to its location nearby the orbitofrontal lobe, the forceps minor is generally affected by
-  susceptibility distortions.
-
-We recommend \diffantom{} as ground-truth in verification and validation frameworks
-  (\autoref{fig:diffantom-02}, panel B) for testing pipelines.
-\Diffantom{} is applicable in the unit testing of algorithms, the integration testing of
-  modules in workflows, and the overall system testing.
-Some potential applications follow:
-\begin{itemize}
-  \item Investigating the impact of different diffusion sampling schemes on the local microstructure
-    model of choice and on the subsequent global tractography outcome.
-  Since the gradient scheme can be set by the user, \diffantom{} can be seen as a mean to translate the so-called
-  \emph{b-matrix} of the source dataset to any target scheme.
-  \item Assessment of sensitivity and robustness to imaging artifacts (noise, \acrlong{pve} and \gls*{csf} contamination,
-    susceptibility-derived warping, Eddy-currents-derived distortions, etc.) at unit, integration and systems testing levels.
-  \item Using \diffantom{} as in panel B of \autoref{fig:diffantom-02}, it is possible to apply binary classification measures
-    to evaluate the resulting connectivity matrix.
-  Considering the connectivity matrix of the \emph{reference Diffantom} and the resulting
-    matrix of the \emph{test Diffantom}, the \acrfull*{roc} of the pipeline can be characterized.
-  \item Simulation of pathological brains by altering the microstructural
-    model accordingly (e.g. as tumors were simulated in \cite{kaus_simulation_2000}).
-\end{itemize}
-In order to exemplify one of these intended uses, we also release a \diffantom{} including the susceptibility-derived
-  distortion in simulation.
-
-\section*{Discussion and conclusion}
-\noindent\textbf{\textit{Discussion\textcolon}}\label{sec:discussion} %
-Whole-brain, realistic \gls*{dmri} phantoms are necessary in the developing field of structural
-  connectomics.
-\Diffantom{} is a derivative of \citep{wilkins_fiber_2015} in terms of methodology for
-  simulation with two major advances.
-First, the correctness of the \emph{minimally preprocessed} data \citep{glasser_minimal_2013}
-  released within the \gls*{hcp}.
-\cite{wilkins_fiber_2015} explicitly state that their original data were not corrected for certain artifacts,
-  and thus, generated data are affected correspondingly.
-Second, \diffantom{} implements the hindered and restricted compartments model \citep{assaf_composite_2005},
-  which is a more complete model than the multi-tensor diffusion model.
-
-A possible competitor to \diffantom{} is the phantom generated for the Tractography Challenge in
-  ISMRM 2015.
-Similarly to \diffantom{}, the organizers used an \gls*{hcp} subject as source of structural information.
-While this phantom is designed for the bundle-wise evaluation of tractography (with the scores defined in the
-  \emph{Tractometer} \citep{cote_tractometer_2013}, such as geometrical coverage, valid connections, invalid connections,
-  missed connections, etc.), \diffantom{} is intended for the connectome-wise evaluation of results,
-  yielding a tractography with a large number of bundles.
-Therefore, \diffantom{} and \emph{FiberFox} are complementary as the hypotheses that can be investigated are different.
-Moreover, \diffantom{} does not require costly manual segmentation of bundles, highly demanding in terms of physiology
-  expertise and operation time.
-The software workflow released with this data report (the \emph{diffantomizer}) ensures the reproducibility of
-  \diffantom{} and enables the generation of custom \emph{diffantoms}.
-The \emph{diffantomizer} is designed for, but not limited to, use \gls*{hcp} datasets as source of structural information.
-
-\noindent\textbf{\textit{Conclusion\textcolon}}\label{sec:conclusion} %
-\Diffantom{} is a whole-brain digital phantom generated from a dataset from the \acrlong*{hcp}.
-\Diffantom{} is presented here to be openly and freely distributed along with the \emph{diffantomizer} workflow
-  to generate new \emph{diffantoms}.
-We encourage the neuroimage community to contribute with their own \emph{diffantoms} and share them openly.
-
-\section*{Data Sharing}
-The first \diffantom{} and its distorted version are available under the \gls*{cc0} using the Dryad Digital Repository
-  (\doi{10.5061/dryad.4p080}).
-The package is organized following the \gls*{bids} standard.
-The associated software to ``\emph{diffantomize}'' real \gls*{dmri} datasets is available at
-  \url{https://github.com/oesteban/diffantom} under an MIT license.
-\emph{Phantomas} is available in \url{https://github.com/ecaruyer/Phantomas} under the revised-BSD license.
-
-\section*{Disclosure}
-
-The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.
-
-\section*{Author Contributions}
-All the authors contributed to this study.
-OE designed the data generation procedure, implemented the processing pipelines and generated the example dataset.
-EC implemented \emph{Phantomas} \citep{caruyer_phantomas_2014}, helped integrate the project with the simulation routines.
-OE, EC, AD thoroughly discussed and framed the aptness of the data in the community.
-AD, MBC, MJLC, and AS interpreted the resulting datasets.
-MBC, MJLC, and AS advised on all aspects of the study.
-
-\section*{Acknowledgments}
-We thank Gert Wollny for his revision of this work.
-\textit{Funding\textcolon}
-This study was supported by the Spanish Ministry of Science and Innovation
-  (projects TEC-2013-48251-C2-2-R and INNPACTO XIORT), Comunidad de Madrid (TOPUS) and
-  European Regional Development Funds, the Center for Biomedical Imaging
-  (CIBM) of the Geneva and Lausanne Universities and the EPFL, as well as the
-  Leenaards and Louis Jeantet Foundations.
-
-\nolinenumbers
-\bibliographystyle{frontiers/frontiersinSCNS_ENG_HUMS}
-%\bibliographystyle{frontiersinHLTH&FPHY} % for Health and Physics articles
-\bibliography{Remote}
-
-%%% Upload the *bib file along with the *tex file and PDF on submission if the bibliography is not in the main *tex file
-
-\glsresetall
-\linenumbers
 \section*{Appendix}\label{sec:appendix}
@@
 The final $\{T_j\}$ maps of isotropic fractions are computed as follows:
-
   \begin{align*}
@@
 The final $\{F_i\}$ maps are computed as follows:
-
 \begin{align*}
@@
 \newpage
-\section*{Figures}
-
-\begin{figure}[h!]
-\begin{center}
-\includestandalone[width=\linewidth]{figure01}
-\end{center}
-\textbf{\refstepcounter{figure}\label{fig:diffantom-01} Figure \arabic{figure}. }{%
-\textbf{A. Microstructural model of \diffantom{}}.
-The phantom is simulated from an underlying microstructural model specified with the
-  following volume-fraction maps: three hindered-diffusion compartments $\{T_1, T_2, T_3\}$,
-  one free-diffusion compartment $T_4$ corresponding to the \gls*{csf},
-  three restricted-diffusion compartments $\{F_i\}$, and three vectorial maps
-  associated with the local fiber directions $\{\mathbf{V}_i\}$.
-Please note the piece-wise linear function of the color scale to enable visibility of small volume fractions.
-\textbf{B. The \emph{diffantomizer} workflow, a workflow to generate \emph{diffantoms}}.
-The pipeline to generate phantoms from any \gls*{hcp} dataset is presented in the lower panel.
-Once the microstructural model shown in the upper panel has been prepared as described in \nameref{sec:diffantom-data-micromodel},
-  the local orientations are computed and fed into \emph{Phantomas} to finally simulate the signal.
-}
-\end{figure}
-
-\begin{figure}[h!]
-\begin{center}
-\includestandalone[width=\linewidth]{figure02}
-\end{center}
-\textbf{\refstepcounter{figure}\label{fig:diffantom-02} Figure \arabic{figure}. }{%
-\textbf{A. Example dataset.}
-A1 and A3 show the tractogram of fibers crossing slice 56 of \diffantom{} as
-  extracted with \emph{MRTrix}, represented over the corresponding slice of the
-  \lowb{} volume for the original (A1) and the distorted (A3) phantoms, with a gray
-  frame highlighting the absence of important tracks.
-Panels A2 and A4 show the segmentation of the right \gls*{cst} represented with blue
-  streamlines, the left \gls*{cst} (red streamlines), and the forceps minor (green streamlines)
-  using \emph{tract\_querier}.
-A2 and A4 include the slice 56 of the \lowb{} and the pial surface is represented
-  with transparency.
-In the distorted \diffantom{} (A4) the forceps minor was not detected.
-\textbf{B. Recommended use of \diffantom{}}.
-The phantom is designed to be used as ground-truth information in evaluation frameworks,
-  to implement unit test of algorithms, to check integration of processing units within
-  pipelines or to validate complete workflows.
-For instance, in order to evaluate artifacts, a perturbation can be induced in the microstructural
-  model or after simulation to provide reference and test datasets.
-}
-\end{figure}
+\bibliographystyle{frontiers/frontiersinSCNS_ENG_HUMS}
+%\bibliographystyle{frontiersinHLTH&FPHY} % for Health and Physics articles
+\bibliography{Remote}
 
 \end{document}
+
+% \documentclass[english]{frontiers/frontiers_suppmat} % for Science, Engineering and Humanities and Social Sciences articles
+% \usepackage{xifthen}
+% \newboolean{review}
+% \setboolean{review}{true}
+
+% \usepackage[mode=buildnew]{standalone}
+% \usepackage{tikz}
+% \usepackage[framemethod=TikZ]{mdframed}
+% %\setcitestyle{square}
+% \usepackage{url,microtype}
+% \usepackage[right]{lineno}
+% \usepackage[toc,nomain,acronym,shortcuts,translate=false]{glossaries}
+% \usepackage[colorlinks=true,linkcolor=black, citecolor=black!80, urlcolor=black!80]{hyperref}
+% \usepackage{doi}
+% \usepackage[onehalfspacing]{setspace}
+
+% \def\keyFont{\fontsize{7}{9}\helveticabold }
+
+% \begin{document}
+% \firstpage{1}
+% \title[Diffantom]{Diffantom: whole-brain diffusion MRI phantoms derived from real datasets of the \acrlong{hcp}}
+% \maketitle
+
+% \end{document}

### 2016-09-26 diffantom/info.py
Update info.py
Fix license

@@ -13,7 +13,7 @@ __version__ = '1.0.0'
 __author__ = 'Oscar Esteban'
 __affiliation__ = 'Psychology Department, Stanford University'
 __credits__ = ['Oscar Esteban']
-__license__ = '3-clause BSD'
+__license__ = 'MIT License'
 __maintainer__ = 'Oscar Esteban'
 __email__ = 'code@oscaresteban.es'
 __status__ = 'Prototype'
@@ -44,7 +44,7 @@ CLASSIFIERS = [
     'Development Status :: 4 - Beta',
     'Intended Audience :: Science/Research',
     'Topic :: Scientific/Engineering :: Image Recognition',
-    'License :: OSI Approved :: BSD License',
+    'License :: OSI Approved :: MIT License',
     'Programming Language :: Python :: 2.7',
 ]
 

### 2016-09-26 diffantom/info.py
add sticks2dwis

@@ -9,7 +9,7 @@ from __future__ import absolute_import, division, print_function, unicode_litera
 from datetime import datetime
 
 __packagename__ = 'diffantom'
-__version__ = '1.0.0'
+__version__ = '1.0.1'
 __author__ = 'Oscar Esteban'
 __affiliation__ = 'Psychology Department, Stanford University'
 __credits__ = ['Oscar Esteban']

### 2016-09-26 diffantom/scripts/__init__.py

@@ -1,0 +1,6 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+# @Author: oesteban
+# @Date:   2015-06-15 14:50:01
+# @Last Modified by:   Oscar Esteban
+# @Last Modified time: 2015-06-23 13:05:13

### 2016-09-26 diffantom/scripts/sticks2dwis.py

@@ -1,0 +1,332 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+# emacs: -*- mode: python; py-indent-offset: 4; indent-tabs-mode: nil -*-
+# vi: set ft=python sts=4 ts=4 sw=4 et:
+#
+# @Author: oesteban - code@oscaresteban.es
+# @Date:   2014-06-05 16:51:52
+# @Last Modified by:   oesteban
+# @Last Modified time: 2016-09-06 09:35:45
+"""
+Example script to be used with the Diffantom
+--------------------------------------------
+
+This is the script used by Diffantom [Esteban2016]_
+
+  .. [Esteban2016] Esteban O. et al., Diffantom: whole-brain diffusion MRI phantoms
+       derived from real datasets of the Human Connectome Project, Front. Neuroinf. 10(4),
+       2016. doi:`10.3389/fninf.2016.00004 <http://dx.doi.org/10.3389/fninf.2016.00004>`_.
+
+"""
+
+from __future__ import division
+import argparse
+import os
+import sys
+
+import numpy as np
+import nibabel as nb
+
+import logging
+from phantomas.geometry.fiber import FiberSource
+from phantomas.geometry.models import Fiber, IsotropicRegion
+from phantomas.mr_simul.partial_volume \
+    import (compute_affine_matrix,
+            compute_spherical_region_masks,
+            compute_fiber_masks,
+            compute_volume_fractions)
+from phantomas.mr_simul.image_formation \
+    import (relaxation_time_images,
+            mr_signal,
+            rician_noise,
+            get_tissue_physical_parameters)
+from phantomas.mr_simul.fod \
+    import (compute_directions,
+            compute_fod,
+            compute_fod_sh)
+from phantomas.utils import shm
+from phantomas.mr_simul.synthetic import GaussianModel
+
+LOGGER = logging.getLogger('diffantom')
+LOGGER.setLevel(logging.INFO)
+
+# Function to check if file exists
+def is_valid_file(parser, arg):
+    if not os.path.exists(arg):
+        parser.error("Cannot open file %s." % arg)
+    else:
+        return open(arg, 'r')
+
+def main():
+    """
+    Simulates diffusion-weighted images from the description of a phantom.
+    """
[…]

### 2016-09-26 setup.py

@@ -55,7 +55,8 @@ def main():
         package_data={PACKAGE_NAME: ['data/bedpostx.json',]},
         entry_points={
             'console_scripts': [
-                'diffantom=diffantom.diffantom:main'
+                'diffantom=diffantom.scripts.diffantom:main',
+                'sticks2dwis=diffantom.scripts.sticks2dwis:main'
 
             ]
         }

### 2016-09-26 diffantom/scripts/diffantom.py
fix error: incorrect import in diffantom script

@@ -2,8 +2,8 @@
 # -*- coding: utf-8 -*-
 # @Author: Oscar Esteban
 # @Date:   2015-06-25 15:46:08
-# @Last Modified by:   Oscar Esteban
-# @Last Modified time: 2015-09-09 15:50:20
+# @Last Modified by:   oesteban
+# @Last Modified time: 2016-09-26 09:44:07
 """
 ============================
 The Diffantom software layer
@@ -19,7 +19,7 @@ def main():
     import os.path as op
     import sys
     from shutil import copyfileobj
-    from .workflows import gen_diffantom, gen_model, finf_bundles
+    from diffantom.workflows import gen_diffantom, gen_model, finf_bundles
 
     parser = ArgumentParser(description='Preprocessing dMRI routine',
                             formatter_class=RawTextHelpFormatter)

### 2016-09-26 diffantom/info.py
version bump

@@ -9,7 +9,7 @@ from __future__ import absolute_import, division, print_function, unicode_litera
 from datetime import datetime
 
 __packagename__ = 'diffantom'
-__version__ = '1.0.1'
+__version__ = '1.0.2a0'
 __author__ = 'Oscar Esteban'
 __affiliation__ = 'Psychology Department, Stanford University'
 __credits__ = ['Oscar Esteban']
